### 2019-03-11 paper.tex
Added online phase/ only thing missing is the improvemetn with different beaver triples.

% This template has been tested with LLNCS DOCUMENT CLASS -- version 2.20 (10-Mar-2018)

% !TeX spellcheck = en-US
% !TeX encoding = utf8
% !TeX program = pdflatex
% !BIB program = bibtex
% -*- coding:utf-8 mod:LaTeX -*-

% "a4paper" enables:
%  - easy print out on DIN A4 paper size
%
% One can configure a4 vs. letter in the LaTeX installation. So it is configuration dependend, what the paper size will be.
% This option  present, because the current word template offered by Springer is DIN A4.
% We accept that DIN A4 cause WTFs at persons not used to A4 in USA.

% "runningheads" enables:
%  - page number on page 2 onwards
%  - title/authors on even/odd pages
% This is good for other readers to enable proper archiving among other papers and pointing to
% content. Even if the title page states the title, when printed and stored in a folder, when
% blindly opening the folder, one could hit not the title page, but an arbitrary page. Therefore,
% it is good to have title printed on the pages, too.
%
% It is enabled by default as the springer template as of 2018/03/10 uses this as default

% German documents: pass ngerman as class option
% \documentclass[ngerman,runningheads,a4paper]{llncs}[2018/03/10]
% English documents: pass english as class option
\documentclass[english,runningheads,a4paper]{llncs}[2018/03/10]

%% If you need packages for other papers,
%% START COPYING HERE

% Set English as language and allow to write hyphenated"=words
%
% In case you write German, switch the parameters, so that the command becomes
%\usepackage[english,main=ngerman]{babel}
%
% Even though `american`, `english` and `USenglish` are synonyms for babel package (according to https://tex.stackexchange.com/questions/12775/babel-english-american-usenglish), the llncs document class is prepared to avoid the overriding of certain names (such as "Abstract." -> "Abstract" or "Fig." -> "Figure") when using `english`, but not when using the other 2.
% english has to go last to set it as default language
\usepackage[ngerman,main=english]{babel}
%
% Hint by http://tex.stackexchange.com/a/321066/9075 -> enable "= as dashes
\addto\extrasenglish{\languageshorthands{ngerman}\useshorthands{"}}
%
% Fix by https://tex.stackexchange.com/a/441701/9075
\usepackage{regexpatch}
\makeatletter
\edef\switcht@albion{%
  \relax\unexpanded\expandafter{\switcht@albion}%
}
\xpatchcmd*{\switcht@albion}{ \def}{\def}{}{}
\xpatchcmd{\switcht@albion}{\relax}{}{}{}
\edef\switcht@deutsch{%
  \relax\unexpanded\expandafter{\switcht@deutsch}%
}
\xpatchcmd*{\switcht@deutsch}{ \def}{\def}{}{}
\xpatchcmd{\switcht@deutsch}{\relax}{}{}{}
\edef\switcht@francais{%
  \relax\unexpanded\expandafter{\switcht@francais}%
}
\xpatchcmd*{\switcht@francais}{ \def}{\def}{}{}
\xpatchcmd{\switcht@francais}{\relax}{}{}{}
\makeatother

\usepackage{url}
\usepackage{ifluatex}
\ifluatex
  \usepackage{fontspec}
  \usepackage[english]{selnolig}
\fi

\iftrue % use default-font
  \ifluatex
    % use the better (sharper, ...) Latin Modern variant of Computer Modern
    \setmainfont{Latin Modern Roman}
    \setsansfont{Latin Modern Sans}
    \setmonofont{Latin Modern Mono} % "variable=false"
    %\setmonofont{Latin Modern Mono Prop} % "variable=true"
  \else
    % better font, similar to the default springer font
    % cfr-lm is preferred over lmodern. Reasoning at http://tex.stackexchange.com/a/247543/9075
    \usepackage[%
      rm={oldstyle=false,proportional=true},%
      sf={oldstyle=false,proportional=true},%
      tt={oldstyle=false,proportional=true,variable=false},%
      qt=false%
    ]{cfr-lm}
  \fi
\else
  % In case more space is needed, it is accepted to use Times New Roman
  \ifluatex
    \setmainfont{TeX Gyre Termes}
    \setsansfont[Scale=.9]{TeX Gyre Heros}
    % newtxtt looks good with times, but no equivalent for lualatex found,
    % therefore tried to replace with inconsolata.
    % However, inconsolata does not look good in the context of LNCS ...
    %\setmonofont[StylisticSet={1,3},Scale=.9]{inconsolata}
    % ... thus, we use the good old Latin Modern Mono font for source code.
    \setmonofont{Latin Modern Mono} % "variable=false"
    %\setmonofont{Latin Modern Mono Prop} % "variable=true"
  \else
    % overwrite cmodern with the Times variant
    \usepackage{newtxtext}
    \usepackage{newtxmath}
    \usepackage[zerostyle=b,scaled=.9]{newtxtt}
  \fi
\fi

\ifluatex
\else
  % fontenc and inputenc are not required when using lualatex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc} %support umlauts in the input
\fi

\usepackage{graphicx}

% backticks (`) are rendered as such in verbatim environment. See https://tex.stackexchange.com/a/341057/9075 for details.
\usepackage{upquote}

% Nicer tables (\toprule, \midrule, \bottomrule - see example)
\usepackage{booktabs}

%extended enumerate, such as \begin{compactenum}
\usepackage{paralist}

%put figures inside a text
%\usepackage{picins}
%use
%\piccaptioninside
%\piccaption{...}
%\parpic[r]{\includegraphics ...}
%Text...

% For easy quotations: \enquote{text}
% This package is very smart when nesting is applied, otherwise textcmds (see below) provides a shorter command
\usepackage{csquotes}

% For even easier quotations: \qq{text}
\usepackage{textcmds}

%enable margin kerning
\RequirePackage[%
  babel,%
  final,%
  expansion=alltext,%
  protrusion=alltext-nott]{microtype}%
% \texttt{test -- test} keeps the "--" as "--" (and does not convert it to an en dash)
\DisableLigatures{encoding = T1, family = tt* }

%tweak \url{...}
\usepackage{url}
%\urlstyle{same}
%improve wrapping of URLs - hint by http://tex.stackexchange.com/a/10419/9075
\makeatletter
\g@addto@macro{\UrlBreaks}{\UrlOrds}
\makeatother
%nicer // - solution by http://tex.stackexchange.com/a/98470/9075
%DO NOT ACTIVATE -> prevents line breaks
%\makeatletter
%\def\Url@twoslashes{\mathchar`\/\@ifnextchar/{\kern-.2em}{}}
%\g@addto@macro\UrlSpecials{\do\/{\Url@twoslashes}}
%\makeatother

% Diagonal lines in a table - http://tex.stackexchange.com/questions/17745/diagonal-lines-in-table-cell
% Slashbox is not available in texlive (due to licensing) and also gives bad results. This, we use diagbox
%\usepackage{diagbox}

% Required for package pdfcomment later
\usepackage{xcolor}

% For listings
\usepackage{listings}
\lstset{%
  basicstyle=\ttfamily,%
  columns=fixed,%
  basewidth=.5em,%
  xleftmargin=0.5cm,%
  captionpos=b}%
\renewcommand{\lstlistingname}{List.}
% Fix counter as described at https://tex.stackexchange.com/a/28334/9075
\usepackage{chngcntr}
\AtBeginDocument{\counterwithout{lstlisting}{section}}

% Enable nice comments
\usepackage{pdfcomment}
%
\newcommand{\commentontext}[2]{\colorbox{yellow!60}{#1}\pdfcomment[color={0.234 0.867 0.211},hoffset=-6pt,voffset=10pt,opacity=0.5]{#2}}
\newcommand{\commentatside}[1]{\pdfcomment[color={0.045 0.278 0.643},icon=Note]{#1}}
%
% Compatibality with packages todo, easy-todo, todonotes
\newcommand{\todo}[1]{\commentatside{#1}}
% Compatiblity with package fixmetodonotes
\newcommand{\TODO}[1]{\commentatside{#1}}

% Bibliopgraphy enhancements
%  - enable \cite[prenote][]{ref}
%  - enable \cite{ref1,ref2}
% Alternative: \usepackage{cite}, which enables \cite{ref1, ref2} only (otherwise: Error message: "White space in argument")

% Doc: http://texdoc.net/natbib
\usepackage[%
  square,        % for square brackets
  comma,         % use commas as separators
  numbers,       % for numerical citations;
%  sort,          % orders multiple citations into the sequence in which they appear in the list of references;
  sort&compress, % as sort but in addition multiple numerical citations
                 % are compressed if possible (as 3-6, 15);
]{natbib}
% In the bibliography, references have to be formatted as 1., 2., ... not [1], [2], ...
\renewcommand{\bibnumfmt}[1]{#1.}

\ifluatex
  % does not work when using luatex
  % see: https://tex.stackexchange.com/q/419288/9075
\else
  % Prepare more space-saving rendering of the bibliography
  % Source: https://tex.stackexchange.com/a/280936/9075
  \SetExpansion
  [ context = sloppy,
    stretch = 30,
    shrink = 60,
    step = 5 ]
  { encoding = {OT1,T1,TS1} }
  { }
\fi

% Put footnotes below floats
% Source: https://tex.stackexchange.com/a/32993/9075
\usepackage{stfloats}
\fnbelowfloat

% Enable that parameters of \cref{}, \ref{}, \cite{}, ... are linked so that a reader can click on the number an jump to the target in the document
\usepackage{hyperref}
% Enable hyperref without colors and without bookmarks
\hypersetup{hidelinks,
  colorlinks=true,
  allcolors=black,
  pdfstartview=Fit,
  breaklinks=true}
%
% Enable correct jumping to figures when referencing
\usepackage[all]{hypcap}

\usepackage[group-four-digits,per-mode=fraction]{siunitx}

%enable \cref{...} and \Cref{...} instead of \ref: Type of reference included in the link
\usepackage[capitalise,nameinlink]{cleveref}
%Nice formats for \cref
\usepackage{iflang}
\IfLanguageName{ngerman}{
  \crefname{table}{Tab.}{Tab.}
  \Crefname{table}{Tabelle}{Tabellen}
  \crefname{figure}{\figurename}{\figurename}
  \Crefname{figure}{Abbildungen}{Abbildungen}
  \crefname{equation}{Gleichung}{Gleichungen}
  \Crefname{equation}{Gleichung}{Gleichungen}
  \crefname{listing}{\lstlistingname}{\lstlistingname}
  \Crefname{listing}{Listing}{Listings}
  \crefname{section}{Abschnitt}{Abschnitte}
  \Crefname{section}{Abschnitt}{Abschnitte}
  \crefname{paragraph}{Abschnitt}{Abschnitte}
  \Crefname{paragraph}{Abschnitt}{Abschnitte}
  \crefname{subparagraph}{Abschnitt}{Abschnitte}
  \Crefname{subparagraph}{Abschnitt}{Abschnitte}
}{
  \crefname{section}{Sect.}{Sect.}
  \Crefname{section}{Section}{Sections}
  \crefname{listing}{\lstlistingname}{\lstlistingname}
  \Crefname{listing}{Listing}{Listings}
}

%Intermediate solution for hyperlinked refs. See https://tex.stackexchange.com/q/132420/9075 for more information.
\newcommand{\Vlabel}[1]{\label[line]{#1}\hypertarget{#1}{}}
\newcommand{\lref}[1]{\hyperlink{#1}{\FancyVerbLineautorefname~\ref*{#1}}}

\usepackage{xspace}
%\newcommand{\eg}{e.\,g.\xspace}
%\newcommand{\ie}{i.\,e.\xspace}
\newcommand{\eg}{e.\,g.,\ }
\newcommand{\ie}{i.\,e.,\ }

%introduce \powerset - hint by http://matheplanet.com/matheplanet/nuke/html/viewtopic.php?topic=136492&post_id=997377
\DeclareFontFamily{U}{MnSymbolC}{}
\DeclareSymbolFont{MnSyC}{U}{MnSymbolC}{m}{n}
\DeclareFontShape{U}{MnSymbolC}{m}{n}{
  <-6>    MnSymbolC5
  <6-7>   MnSymbolC6
  <7-8>   MnSymbolC7
  <8-9>   MnSymbolC8
  <9-10>  MnSymbolC9
  <10-12> MnSymbolC10
  <12->   MnSymbolC12%
}{}
\DeclareMathSymbol{\powerset}{\mathord}{MnSyC}{180}

\ifluatex
\else
  % Enable copy and paste - also of numbers
  % This has to be done instead of \usepackage{cmap}, because it does not work together with cfr-lm.
  % See: https://tex.stackexchange.com/a/430599/9075
  \input glyphtounicode
  \pdfgentounicode=1
\fi

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

%% END COPYING HERE

% Add copyright
% Do that for the final version or if you send it to colleagues
\iffalse
  %state: intended|submitted|llncs
  %you can add "crop" if the paper should be cropped to the format Springer is publishing
  \usepackage[intended]{llncsconf}

  \conference{name of the conference}

  %in case of "llncs" (final version!)
  %example: llncs{Anonymous et al. (eds). \emph{Proceedings of the International Conference on \LaTeX-Hacks}, LNCS~42. Some Publisher, 2016.}{0042}
  \llncs{book editors and title}{0042} %% 0042 is the start page
\fi

% For demonstration purposes only
\usepackage[math]{blindtext}
\usepackage{mwe}

\begin{document}

\title{Secure Multi-Party Computation with SCALE-MAMBA}
%If Title is too long, use \titlerunning
%\titlerunning{Short Title}

%Single insitute
\author{Jonas Rülfing}
%If there are too many authors, use \authorrunning
%\authorrunning{First Author et al.}
\institute{Seminar: Privacy and Big Data - RWTH Aachen}

%% Multiple insitutes - ALTERNATIVE to the above
% \author{%
%     Firstname Lastname\inst{1} \and
%     Firstname Lastname\inst{2}
% }
%
%If there are too many authors, use \authorrunning
%  \authorrunning{First Author et al.}
%
%  \institute{
%      Insitute 1\\
%      \email{...}\and
%      Insitute 2\\
%      \email{...}
%}

\maketitle

\begin{abstract}
  This thesis takes a look at SMPC especally using SCALE-MAMBA. For that it introduces first what SMPC is,  how SPDZ the precessor of SCALE-MAMBA works and what are the differences. It discusses the functionalities of SCALE-MAMBA, especially the split into online and offline phase, how the beaver tripels are used for sacrificing and in which complexity classes the calculations are done. Finally it compares scale mamba shortly to other competitors in the market and gives an outlook what next technologies are coming right now and are coming in the near future in SMPC.
\end{abstract}

\section{Motivation and Introduction}\label{sec:intromoti}

In todays world, and even more in the future, there are a lot of computations that have to be done in a distributed environment, this might be because there are different parties involved, but also because the amount of data is getting so big that it becomes impossible to have a single entity that processes all the data at once. Furthermore, in a lot of these cases the parties participating in these computations don't want to reveal their own information, because its sensitive or because they don't want to share information with a lot of business value.\\
A basic example is an anonymized voting. In most elections there is a third party that has to aggregate and count the votings made by the votings. For instance in germanys elections this is done by specific people counting a small group of votes and then forwarding the information. The other problem is that it has to be guaranteed that a vote is valid, meaning that the voter has the right to vote and is only voting one time. So if there would be a malicious person in the system, it would be easy for that person to figure out what vote a specific person made.\\
Another use case is the exchange of medical data. Medical data of patients is highly confidential and shoudn't be publiced under n circumstances. But at the same time a lot of different hospitals or doctors have only the data from their small data set of local patients, and for making diagnosis for new patients it is helpful to compare the symptoms to the diagnosis of other doctors.\\
Secure Multi-Party Computation(SMPC) is one possible solution for these use cases. It is a field of research that emerged in the last 30 years, and is gathering a lot of speed in the last years. 
The basic problem is how to evaluate a common function in which multiple parties have different inputs to this function, without revealing the different inputs of the parties. Although there are a lot of theoretical solutions for this problem, real-world implementations are still not perfmorming on the desired level. So in the last years the research shifted more in the area on how to make the existant solutions performant enough while still guaranteeing a desired level of security.\\
This paper's main focus will lay on SCALE-MAMBA, a system developed mainly by the KU Leuven, which is one of the frontrunners in SMPC and a system that combines a lot of research by different teams with the newest ongoing research to provide a platform with highly performant SMPC. Nevertheless, this paper will at first give a short history of SMPC and will try to explain the basic technologies of SMPC to the reader. It will compare SCALE-MAMBA to its predecessors and will also try to give a short comparision to other SMPC systems which use different approaches, especially considering the usability in different use cases. Ultimately, it will try to give an outlook what might be coming in the next years.\\

\section{Cryptography Basics}\label{sec:basics}

Before diving deeper into SMPC, this section will shortly recap on cryptography basics that might be unknown to the reader. These basics are not solely used by SMPC, but nevertheless, play a huge role in a lot of protocols.

\subsection{Message Authentification Codes}
Message Authentification Codes (MACs) are used to guarantee the integrity and authenticity of a message. Basically, a MAC is a tag that is generated for a message and that, if it is correct, verifies the identity of the sender and that the content of the message was not changed.\\
For using MACs we need a tupel of two algorithms \((\text{MAC},\text{VER})\):\\
\begin{itemize}
\item The MAC algorithm \(\text{MAC}(K,m)\rightarrow t\) takes a key \(K\) and a message \(m\) and outputs a tag \(t\)\\
\item The verficiation algorithm \(\text{VER}(K,m,t)\rightarrow 0/1\) takes a key \(K\), a message \(m\) and a tag \(t\) and outputs accept(\(1\)) or reject(\(0\))\\
\end{itemize}
The correctness can be easily seen: \(\forall K: \text{VER}(K,m,\text{MAC}(K,m))=1\)\\

In SCALE-MAMBA really simple MACs are used. We assume that we have a MAC key \( \alpha \). Then the MAC algorithm used is following:\\
$$\text{MAC}(\alpha,m)\rightarrow \alpha * m$$ 
$$\text{VER}(\alpha,m,t)\rightarrow (1 \Leftrightarrow t = \alpha*m) $$

These algorithm has the disadvantage of high storage complexity, but therefore it doesn't take a long time to calculate and it is possible to get the original message with just having the key and the tag. This is a property that is used for an advantage later on in the protocol.\\

\subsection{Homomorphic Encryption}
Encryption as a basic concept is simple. Similiar to MACs, a tupel (E,D) of algorithms is needed.\\
\begin{itemize}
\item The enncryption algorithm \(\text{E}(K,m)\rightarrow c\) takes a key \(K\) and a message \(m\) and outputs a cyphertext \(t\) that is the encryption of the message \(m\)\\
\item The decryption algorithm \(\text{D}(K,c)\rightarrow m\) takes a key \(K\) and a cyphertext \(m\) and outputs the decrypted message \(m\), assuming c was not modified\\
\end{itemize}
Alternatively, in a public key setting, the key for encryption is public and there is a different key for decryption that is privat.\\
The problem is, that modifying the cypertext in any way makes the decryption useless. In some cases it is desirable do do computations on cypertexts, because then it is possible to make calculations without revealing the unencrypted data. That is why Homomorphic Encryption was developed. Let's assume that we have a ring \((R,+,*)\) as our message space. Fully Homomorphic Encryption (FHE) is preserving the ring structure. Therefore, it is possible to do additions and multiplications of cypertexts and when decrypting the result it is the same value as when doing the same additions and multiplications on the real values.\\
Althought FHE has amazing properties for a lot of usecases, in praxis it is way too slow. Because of all the overhead generated by the encryption scheme to guarantee the preservation of the ring structure, the current algorithms are not feasible for most applications.

\subsection{Somewhat Homomorphic Encryption}
That is where Somewhat Homomorphic Encryption (SHE) becomes relevant. SHE only supports one operation, and in most cases that is addition. This takes away a lot of possibilities but, therefore, the performance gain is immense. So immense, that SHE is used a lot in practice, for example in SCALE-MAMBA.\\
The SHE scheme used in SCALE-MAMBA, is based on the so called Brakerski-Gentry-Vaikuntanathan cryptosystem(BGV).\\

\subsection{Adversary types}
Cryptography is the science of protecting information or communication from an adversary. This adversary is normally a third party that wants to get information for his own gain or want to sabotage something because of his own interest. In the are of SMPC we we normally divide between two types of adversaries. In SMPC we expect that everyone but ourselves can be controlled by an adversary.  That means the adversary can control between one and all of the other parties with which we are communicating. The difference to the normal cryptographic adversary model is, that the adversary is one of the parties we are communicating with but not a third party that tries to intercept communication.
\subsubsection{Honest but curious}
This type of adversary tries to gain as much information as possible. But this adversary is obeying to the rules of our protocol, that means it is impossible to know that this party is controlled by a third entity. This adversary tries to retrieve some of the secret information from the other parties, by using all the information he gets during the execution of the protocol.

\subsection{Malicious adversary}
While this adversary also wants to gain all the information, he is not obeying to the rules of the protocol. This type of adverary can send different inputs to manipulate the results he receives. Furthermore, he wants to interrupt the whole protocol. That could mean making the computation impossible, but as well modifying his own input or his own calculations, so the final result is changed and might lead to different implications

It is important to decide between these two types of adversary, because it shows the different securities SMPC has to guarantee. It has to guarantee that no information is unwillingly shared with an illegal party and at the same time verify that the final result has a correct value. Most SMPC protocols try to guarantee safety versus both type of adveraries, bvut there are also approaches that only use passive security, i.e. only protecting versus non malicious adversaries.

\section{SMPC Basics and History}\label{sec:smpc}
\subsection{Beginnings}
As we saw in the introduction, there are a lot current and much more upcoming usecases of calculating the output of a function in a distributed environment with multiple parties while at the same time keeping all the individual inputs secret. The first explorations in these fields started in the 1970s, with theoretical problems like the mental poker problem(how can two players play poker while not being in the same physical location and making sure noone is cheating) and Yaos millionaire problem(how can can two millionaires decide who is richer than the other while not revealing their wealth). \newline
These explorations lead to a group of basic protocols which established the area of Secure Multi-Party Computation(SMPC).\\
 Now we will take a short look on the two most famous of these protocols, which both were the basis for a whole family of algorithms established in the years to follow until today.

\subsection{Yaos Garbled Circuit}
There are typically two groups of SMPC protocols, in which most of the protocols can be categorized. One is based on garbled circuits which were first introduced by Yao in 1986 \cite{4568207}. The basic idea is, that, in a two party setting, one party, Alice, garbles the boolean circuit that has to be evaluated and sends it to the second party, Bob. Bob uses his own input, Alice's encrypted input and the garbled circuit to evaluate the circuit in an ecnrypted way. At the end they exchange the results and can compute the result this way.

\subsection{Shamirs Secret Sharing Scheme}
This paper will not focus more on the family of garbled circuit protocols, but rather on the second family which are based on a so called Linear Secret Sharing Scheme(LSSS). 
These type of protocols were introduced by Adi Shamir in a similiar time, in 1979 \cite{Shamir:1979:SS:359168.359176}.\\
The basic idea of this approach is that multiple parties split their secrets into so called shares and then distribute them to the other parties. This is done by generating random polynomials with a degree depending on the security level and then calculating the shares as points in this polynomial. Computations are executed on these shares, and after the end it is possible to reconstruct the result from the different shares without revealing information about the initial secrets. This is done by using Lagrange interpolation.\\
 There is a huge family of protocols build on this approach, and it is still evolving every day. This paper will focus on the newest of these protocols that were developed in the last 10 years, especially on the SPDZ like protocols.\\
In SMPC, normally all operations are executed in a finite field \(F_p = \{ 0,...,p-1\}\) with a prime p. That means all operations use modulo p at the end on the result to keep the result in the field. Therefore, for now on \(F_p\) will be used and when using an exact application, then p has to be specified depending on the usecase.\\

\subsection{Terminology}
Over the years different terminologies came up in the are of SMPC. Following will be a short introduction into some of them, so their definitions are clear.

\subsubsection{Oblivious Transfer}

\subsubsection{Access structures}
There are a lot of different use cases for SMPC, so of course the adversary model changes depending on the system, the security needs, and the amount of parties involved. it is important to differentiate between these different use cases, because the different needs have big implications for security measurements that have to be taken and for runtime complexity. SMPC should run as fast as possible, while still providing all the neccesary security guarantees.

To describe which party has which rights, and how many parties can be corrupted so the security can still be guaranteed, access structures were introduced.
An access structure describes how many parties are needed to form a so called qualified set. A qualified set is a group of parties which can pool their information to extract secret information..  An access structure is called \(Q_l \) if at least l+1 parties are needed to  form a qualified set. 
An (n,t)-threshold scheme with t<n/l means that at least t+1 parties are needed to form a qualified set. Here, n is the number of total parties, and t is number of possible adversaries while not revealing any secrets.
A full threshold scheme means that all parties are needed to form a qualified set. This implies, that, if all other parties except one are corrupted by an adversary, this one party can still rely on the fact that it secret remains secret. Of course this has a huge impact on overhead and performance, but in some cases it can give needed security guarantees.
An access structure also doesn't have to be symmetric. In some cases there might be different type of parties, where different parties have different positions or levels of trust, so it is possible to specify that some set of parties only has to have the size of 2 to form a qualified set and another one needs the size of 4.\\
Choosing the right access structure for the right use case is really important, because it has a realy big impact on performance and on the desired security level.

TODO: SOUNDNESS SLACK
TODO: LATTICE BASED CRYPTOGRAPHY

Scale-Mamba is an implementation of SPDZ among other protocols. SPDZ is right now the most efficient protocol in use.

Most LSSS-based MPC protocols split the computation into two parts: an of-

ine phase, in which parties interact using expensive" public-key cryptography
to lay the groundwork for an online phase in which only cheap" informationtheoretic
primitives are required. The online phase is where the actual circuit
evaluation takes place. For the access structures considered in this work, namely
Q2 structures, the offline phase is almost as fast as the online phase. Thus the
goal here is to minimize the cost of communication in both phases.

\section{SCALE-MAMBA}\label{sec:scalemamba}

\subsubsection{Overview}

In the last years, the research in SMPC is constantly ongoing, and there are new approaches or improvements to old approaches appearing regulary. But most of these publications only focus on a small area of the whole picture. SCALE-MAMBA\cite{ScaleMambaDocu} is a project that tries to combine the state-of-the-art research into one system. That means, that it is offering kind of an end-to-end system for SMPC.
It is a research system that is continously in development, therefore, there is not a single final version that can be examined. It offers as well a wide range of protocols for different usecases, i.e. it is possible to use the system in different confidurations. That also means for this thesis, that there is not a single process that can be examined. Therefore, this thesis tries to give more an overview how the process works in general, which are the main, especially which are newest and best, parts and where are the problems and bottlenecks.\\
SCALE-MAMBA consists of two parts. The one part is called Secure Computation Algorithms from LEuven (SCALE) and is the SMPC part. It contains all the computational logic and algorithms for multi-party computations. The second part is called Multiparty AlgorithMs Basic Argot (MAMBA) and is a compiler for the SMPC programs.\\

In general, the SMPC protocol is split in two parts, first an offline phase and secondly and online phase. In the online phase only the final evaluation of the function is executed. In the offline phase, prior to the online phase, preparations are done by all the parties to prepare for the evaluation. This includes already calculating values for possible multiplications and MACs, so that the online phase can be executed faster. This phase is not completly offline, because the parties still exchange some information which are later needed for the evaluation. But this phasae is not that time critically, that means it can be already executed a lot earlier than the online phase, so that, when the function evaluation is really needed, the result can be calculated in a rather fast time compared to a system that still has to set up everything. In SCALE-MAMBA, compared to its predecessors, the offline and online phase are completly integrated. Therefore, it is not possible anymore to execute them separately like in SPDZ for example. But it is a much closer approach to a real world application.\\

While SCALE-MAMBA also offers different access structures, the most commonly used security configuration is active security with abort in a full threshold scheme, that means that every except one party could be corrupted, and still no information is leaked, because the evaluation is aborted.

\subsubsection{MAMBA}
For this paper the more interesting part is the algorithmic part. But still, the importance of the compiler shouldn't be undervalued. Because while the algorthmic foundations might be more important from a theoretical point of view, the compiler, written in C++, enables the user to really use these technologies in a system where he can rely on the security gurantees given by the algorithms. The programs can be written for MAMBA in a syntax similiar to python's. In these programs, the user can specify which security settings he wishes to use. He can specify, how he wants his SMPC to exactly be executed, and which security parameters he wants to use. MAMBA is the compiler which then translates the instruction to bytecode and executes them. MAMBA is the part that gives the system the possibility to be properly used and tested. It makes it easy, to change the parameters, to be able to execute comparable tests, and to have a abstraction level for the user.\\

\subsubsection{Offline phase}

In the offline or pre-processing phase, the most important part is the generation of the beaver triples. These triples  are independent of the computation that is later performed in the online phase, but will be used in the online phase to guarantee that multiplications can be done without revealing the secret inputs of the different parties.
Beaver triples are a tuple of three numbers \((a,b,c)\), with \( a,b,c\in \mathcal{F} )\), with \(a\) and \(b\) being chosen uniformly randomly, and \(a*b=c\).

The second thing that is done in the offline phase is the generation of a global MAC key. This MAC key is later used, to authentificate the different shares at the end of the computation. This key is also distributedly generated, i.e. only if all parties pool together they can recover its value.\\

\subsubsection{Zero Proofs of Knowledge}
In the many iterations of the SPDZ like protocols, there were different ways to guarantee that the parties are not cheating. An important technology for that are the so called Zero Proofs of Konwledge Protocol(ZPoKP).

In the offline phase, ZPoKP are used to guarantee that the Beaver triples are calculated correctly, or atleast that they dont divert more than a specified error \(e\) the expected result.

ZPoKP are a technique that was introduced  in 1985 \cite{Goldwasser:1985:KCI:22145.22178}. The idea is that one party can prove to another party that it knows a specific value \(x\), without revealing the value of \(x\) to the other person, nor revealing any other knowledge about the value.

Assumed Alice(in this scenario called the prover) wants to prove to Bob (the verifier) that she knows \( x\) in a Field \( \mathcal{F} \), such that \(f(x)=y\). It is as well assumed, that \( f \) is homomorphic with respect to the field operations in  \( \mathcal{F} \).
\begin{enumerate}
\item Alice chooses a random \( s \in \mathcal{F} \) , calculates \( a = f(s)\) and sends \( a\) to Bob
\item Bob samples a random  \( e \) out of  \( \mathcal{F} \) and sends it to Alice
\item Alice calculates \( z=s+e*x\) and sends it to Bob
\item Bob checks if \( f(z)=a+e*y\)
\end{enumerate}

\textbf{Correctness:}
Bob wants to prove that \( z\) provided by Alice contains the proper \( x\), so that \( f(x)=y\). 
Because f is homomorphic, obviously \( f(z)=f(s)+f(e*x)=f(s)+e*f(x)=a+e*y\) with \(f(s)=a\) and \(f(x)=y\).
So Bob has to calculate \( a+e*y and f(z)\) with the provided \( z\) by Alice. If they are equal, he knows that Alice knows the \( x\) so that \( f(x)=y\).
He can do that, because he got provided the \( a\) and \( z\) by Alice, he knew the \( y\) before the protocol and he chose the \( e\) himself.\\

\textbf{Confidentiality:}
We also have to make sure, that Alice did not share any information regarding her x.
In the firt step she chose a random element s and calulated a=f(s). This information is independent on her secret x, so she didn;t share any information.
When calculating z=s+e*x, she also didnt share any information regarding x, as long it is guaranteed that Bob didnt have any knowledge about the s.
So under the assumption that Alice chooses a new random s each time the protcol is executed, she doesn't share any information regarding her secret x.\\

It is not really important for this thesis, which ZPoKP is exactly used, but it should be said that the protocol used in SCALE-MAMBA is based on Schnorr's protocol.\\

\subsubsection{First improvement with MASCOT}
In one of there earlier works, BDOZ, pairwise MACs were used to authentificate the secret sharing between the parties. It used a pairwise multiplication protocol with linear homomorpic encryption. To guarantee the data integritym pairwise zero-knowledge proofs were used, in total O(n2) of them. 
In SPDZ instead of HE, SHE based on the BGV ecnryption scheme was used, reducing the amount of ZK proofs per triple by a factor of n and as well the overhead per triple, because they used a single ZKPoK for proving thousand of triples.
Because the offline phase was the weak point of the existing SPDZ implementations, in regards to runtime while still managing a sufficiently secure system, in 2016 MASCOT \cite{cryptoeprint:2016:505} was introduced, which improved the speed of the offline phase by a magnitude of 2. MASCOT used Oblivious Transfer instead of SHE to significantly reduce the amount of communication and computation. We will not go into detail into this protocol, but rather look on its successor.

\subsubsection{Overdrive}
Shortly afterwards in 2017 a new protocol for the offline phase called Overdrive was introduced \cite{cryptoeprint:2017:1230}, which once again got better results with SHE instead of OT, improving the run time at least by a factor of 6. Basically, it is a really similiar protocol to MASCOT, but just uses SHE in the places where MASCOT used OT. So explaining Overdrive already should give a good impression how MASCOT works. \\
 Overdrive includes two parts, a protocol called Low Gear protocol for small numbers, and a proctocol called High Gear for larger numbers. In version 1.2 of Scale-mamba only the HighGear protocol was used, even for smaller numbers.\\

The biggest difference between the ZKPoK usage in Overdrive and in SPDZ is, that the parties not anymore prove multiple smaller statements between each other with ZKPoK, but instead make a joint statement and prove thi one toegether. Because ZKPoK are really expensive, this makes a big difference in the overall performance.\\
This way, Overdrive did not improve the amount of communication channels used, but it decreased the computational costs by a factor n.\\
A second big strategy used by SPDZ and Overdrive is the so called Soundness Slack. Because in practice it is to expensive to prove that a result is exact, so it is sufficient to prove that the value, here the product of the beaver triple, exact but with the legalization of a small error. Later on, this error, if existing and introduced by an adversary, will be detected by a different technique called sascrificing.\\
Overdrive uses as well drowning, which is the technique of adding noise to a secret sharing. Because otherwise, the noise of the SHE function would reveal information over the secret. So the noise is drowned by introducing an additional random noise that is bigger than the maximum possible noise.\\

\subsubsection{TopGear}
Recently, an even newer improvement for the offline phase was introduced, called TopGear \cite{cryptoeprint:2019:035}. It is implemented in SCALE-MAMBA since version 1.3\cite{ScaleMambaDocu}, which was released in January 2019. It further improved the HighGear protocol of the Overdrive implementation, by targeting specifically the ZKPoKs.

\subsubsection{Online Phase}

\subsubsection{Preliminaries}

Now we assume that the offline phase was succesfully executed, so all preliminaries are given to execute the online phase. It's main goal is to evaluate a specified circuit with secret inputs by the different parties. So this is the heart of the system, the part where the real computatiuon is executed. 
It is important to note, that the online phase is nearly perfectly optimized when assuming the offline phase was carried out succesfully. The online phase has a runtime complexity in O(n), and therefore there is not a really big performance impact. It is also noteworthy, that in the online phase only only computational and no information theoretical problems are solved. All this leads to the fact, that there is not really any change in the online phase in the SPDZ like protocols in the last years, therefore the online phase is the same like in the earlier version of SPDZ ~\cite{cryptoeprint:2011:535}.\\

We assume following preliminaries are given after the execution of the offline phase:\\
\begin{enumerate}
\item There are \(n\) parties \(P_1,...,P_n\)
\item There is a given function \(f(a_1,...,a_m)\) which the parties want to evaluate together. The inputs are secret, and each input is hold by exactly one party.
\item There is a global MAC key \( \alpha\) that is secret to every party and additively shared, so that every party \( P_i\) has a share  \( \alpha_i\) so that  \( \alpha=\alpha_1+...+\alpha_n\)
\item Each of the secret inputs to \(f\) is additively shared between the parties

\end{enumerate}

Following, it is going to be demonstrated, how addition and multiplication are done in this environment. This is sufficient, because a group with addition and multiplication is already touring complete, so there is no need for any other operation.\\

Sharing of values:\\
After sharing a value \( a \in F_p\), it is represented as following tuple:\\
$\langle a\rangle=(\delta_a, (a_1,...,a_n),(\gamma (a)_1,...,\gamma (a)_n))$
\\With:
\begin{itemize}
\item $ a=(a_1+...+a_n)$
\item $ \gamma (a)=((\gamma (a)_1+...+\gamma (a)_n))=\alpha(a+\delta)$
\item Every party \( P_i\) holds a share of \( a\) as tuple \(\langle a_i\rangle = (\delta_a,a_i,\gamma(a)_i )\)

\end{itemize}

\subsubsection{Addition}
Assumed there are two values \( a\) and \( b\) which should be added together and both are secret inputs from two different parties. Naturally, both of these values are additively shared between all the parties, i.e. every party has a share \( \langle a_i\rangle\) and a share \( \langle b_i\rangle\).\\

Then every party \( P_i\) adds up their local shares, so that:\\
$$\langle c_i\rangle=\langle a_i\rangle+\langle b_i\rangle=((\delta_a+\delta_b),(a_i+b_i),\gamma (a)_i+\gamma (b)_i))$$
\textbf{Correctness}: Afterwards, \( c=a+b\) is obviously additively shared between the parties and can be used for further computations.\\

\subsubsection{Addition with constant}
Adding a public constant is really simple in this representation. Because \( \gamma (a)=\alpha(a+\delta)\), adding a constant \(c\) to a additively shared value \(a\) is done by:\\
$$\langle a\rangle+c=(\delta-c,(a_1+c,a_2,...,a_n),,(\gamma (a)_1,...,\gamma (a)_n)$$
Correctness: It can be seen, that the constant is only added to one shared value and subtracted from \( \delta\). Because of that, \( \alpha(a+\delta)\) stays the same, so the same shared MAC values like before can be used. So \(a+c\) is also properly additively shared.\\

\subsubsection{Multiplication with constant}
Multiplicating with a public constant is as well really simple and can also be done by each party invdividually. It is defined as:\\
$$c*\langle a\rangle=\langle c*a\rangle=(c*\delta,(c*a_1,...,c*a_n),,(c*\gamma (a)_1,...,c*\gamma (a)_n)$$
\textbf{Correctness}:(TO BE DONE\\

\subsubsection{Multiplication}

Multiplication is the only operation,except for distributing the shares and calculating and verifying the end result, where the parties have to interact in the online phase.
Assumed there are two values \( a\) and \( b\) which should be multipled with each other. Naturally, each party has the shares \(\langle a_i\rangle \) and \(\langle b_i\rangle \).\\
For each multiplication operation, we need one of the Beaver triples \( (\langle x_1 \rangle ,\langle y_1\rangle , \langle z_1\rangle )\) generated in the offline phase. In the offline phase we checked the quality of our triples, but we still allowed the possibility that each triple has an error \(e\) so that \(c=a*b+e\). To check the quality of our triple, we use a second triple \( (\langle x_2 \rangle ,\langle y_2\rangle ,\langle z_2\rangle )\) in a procedure called sacrificing it. Both triples are additively shared between all parties like the input values. In SPDZ both triples are completly unrelated to each other, and the sacrificing is done as follows:\\

\begin{enumerate}
\item Choose a random variable \( r\) and open it to the other parties. This is done by each party generating a random share \(r_i\) and then calculating \(r=r_1+...+r_n)\).
\item The parties calculate  \( p=t*\langle x_1 \rangle - \langle x_2 \rangle \) and \( \sigma=\langle y_1 \rangle -  \langle y_2 \rangle \) and partially open the results afterwards
\item The parties calculate \(\phi = t * \langle z_1 \rangle\ - \langle z_2 \rangle - \sigma * \langle x_2 \rangle - p * \langle y_2 \rangle - \sigma * p)\) and then partially open the result
\item If \( \phi \) is 0 then the triples are ok and it is proceeded with \( (\langle x_1 \rangle ,\langle y_1\rangle , \langle z_1\rangle )\). Otherwise the system detected a malicious adversary and it aborts.\\
\end{enumerate}

Partially opening  of values: All players send their respective shares of a value, but not the MAC value, to one chosen party \(P_i\) which summs up all values and then broadcasts the result. In praxis this party is changing for every opening to balance the workload.\\
Fully opening of values: In this case all players broadcast their respective shares and the respective MACs to all players, so all players can verify the result.\\

\textbf{Correctness}: If we simplify \( \phi \), we can see that every term disappears (given that the triples are proper triples):
$$\phi = t*z_1 - z_2 - y_1 + y_1*x_2 + y_2 * x_2 - t * x_1 * y_2 + x_2*y_2 - t*x_1*y_1 + t*x_1*y_2 + x_2*y_1 - x_2*y_2$$
$$ = t*(z_1 - x_1*y_2+x_1*y_2 -x_1*y_1) + y_1*x_2 - x_2*y_1+ x_2*y_2 - z_2 + x_2*y_2 -  x_2*y_2$$
$$= t* (z_1 - x_1*y_1) + x_2*y_2 - z_2 $$
So we can see that \( \phi = 0 \Leftrightarrow x_1 * y_1 = z_1 \land x_2*y_2=z_2 \), so we know the triple is correct. \\
In this process we didn't reveal the triple \( (\langle x_1 \rangle ,\langle y_1\rangle , \langle z_1\rangle ) \), so now we can use it for the multiplication knowing that one the one hand it is still secretly shared and unknown and on the other hand that it was not modified by an adversary.\\

There is a small improvement in SCALE-MAMBA compared to SPDZ regarding the triple generation that was added later in ~\cite{cryptoeprint:2017:123}. The idea is to not sacrifice a random second triple, to guarantee the correctness of the first one, but to sacrifice a similiar triple. For this, the triples in the online phase are already generated in pairs, so that there are are two triples \( (\langle x \rangle ,\langle y \rangle , \langle x*y \rangle ) \) and \( (\langle x' \rangle ,\langle y\rangle , \langle x'*y\rangle ) \). This way the triples can be calculated faster in the offline phase, because one half of the multiplication is the same, and still the triples keep the property to be sacrificed for each other to check the correctness of the other triple.\\\\

In SCALE-MAMBA, this triple sacrifice is also already executed in the offline phase. But for consistency the explanation remains here, to better be able to see the difference between SPDZ and SCALE-MAMBA.\\

Now in the next step the multiplication of \(\langle a \rangle \) and \(\langle b \rangle \) is executed with the help of the triple.
\begin{enumerate}
\item The parties calculate \( \epsilon = \langle a \rangle - \langle x_1 \rangle \) and partially open it
\item The parties calculate \( \delta = \langle b \rangle - \langle y_1 \rangle \) and partially open it
\item The parties calculate \( \langle c \rangle \langle z_1 \rangle + \epsilon * \langle b \rangle + \delta * \langle a \rangle + \epsilon * \delta \) 
\end{enumerate}
Correctness: 
$$  \langle a \rangle * \langle b \rangle = \langle (a-x_1+x_1) \rangle * \langle (b-y_1+y_1) \rangle$$
$$ = ( \epsilon  +\langle  x_1 \rangle ) * ( \delta  + \langle y_1 \rangle )$$
$$ = \langle x_1 \rangle * \langle y_1 \rangle + \epsilon * \langle b \rangle + \delta * \langle a \rangle + \epsilon * \delta =  \langle c \rangle $$

So, as we can see, the calculations are correct and at the same time no information of \(\langle a \rangle \) or \(\langle b \rangle \) was revealed.\\

\subsubsection{Verifying Correctness}
At the end of the computation, we have to find out the total calculated result. After all computations were executed, every party has a share of the final result \( r_i\), a share of the MAC value \( \gamma(r)_i)\) of the final result and a share of the global MAC key \( \alpha_i \). The verification is done as follows:

\begin{enumerate}
\item Every party  opens their share \( r_i\) and their share of the MAC value \( \gamma(r)_i) \)
\item Every party computes \( r = \sum_{i=1}^n r_i \) 
\item Every party computes  \( \gamma(r)_i = \sum_{i=1}^n \gamma(r)_i \)
\item Every party computes  \( \sigma_i =  \gamma(r)_i - \alpha_i * r\) and shares it to all the other parties
\item Every party computes \( \sum_{i=1}^n \sigma_i \) and checks if it is \(0\). If it is  \(0\), the computation was correct.
\end{enumerate}
Correctness: 
$$ \sum_{i=1}^n \sigma_i =  \sum_{i=1}^n  \gamma(r)_i - \alpha_i * r $$
$$=  \sum_{i=1}^n  \gamma(r)_i -  \sum_{i=1}^n  \alpha_i * r $$

If the calculations were correct, the MAC was shared properly and not modified, and the MAC values were not modified, then the difference between both sums will be 0. If it is not 0, the protocol detects that there is a malicious adversary and aborts.\\

\subsubsection{Upcoming changes to SCALE-MAMBA}
SCALE-MAMBA is a research system, so there will be a lot more research time spend on improving the system further. An important aspect is the usability in the practice. Because online and offline phase are integrated into each other, the complete runtime sometimes takes a lot of time. This is especially the case in full threshold access structures. Because of the integration, it is right now not possible to execute the offline phase earlier. One of the currently ongoing improvements on SCALE-MAMBA tackles exactly this problem. The developer team tries to give MAMBA the possibility to compile code just in time. That would offere the possibility, that the offline data is calculated before and then, when the need arises, a computatation could be specified in MAMBA and then intime compiled and executed, without having to wait for the expensive offline calculations.\\

\section{Comparison}\label{sec:comparison}

a.	What else is out there?
b.	Different use cases – when which SMPC tool is more useful; what are the strengths of SCALE-MAMBA compared to other ones?
	- Security! High performant for such a secure system
	- In some cases not the highest security is needed; application for a normal user, not for a company, have much less security requirements but much higher requirements on performance
	- Active security is not always required, in non-critical systems it is often sufficient to know that noone knows the users data
c.	Factual comparisons
	- YAO/LSSS --> LSSS protocols right now are more suited for the praxis

	- SCAPI
	-SCAPI is an open-source Java library for implementing secure two-party and multiparty computation protocols (SCAPI stands for the “Secure Computation API”). It provides a reliable, 	efficient, and highly flexible cryptographic infrastructure.
SCAPI provides only passive security

https://scapi.readthedocs.io/en/latest/intro.html

\section{Resume}\label{sec:outlook}

\subsubsection{Outlook}
Their are a lot of future applications for Secure Multiparty Computation. With the threoretical research of the early years much more turning to praxis oriented research, todays SMPC protocols, and especially the successors of SPDZ, have a bright future. With already more and more applications starting to use SMPC in systems that need to guarantee security of private data, the list will only grow more and more with the performance of the protocols and the same time the available computation power improving. \\
With SCALE-MAMBA being one of the first performant systems that can be really used in a real time scenario, it will probably not take a lot of more years until there will be proper commercial solutions. And if big companies start to adapt and use these theoretical foundations, then, in my opinion, SMPC will be a really important tool in computer science and especially in distrbuted environments. There are a lot more possible applications that could be coming in the next years, for example in the financial branche SMPC could play a huge role.\\
SMPC can be a technology that shapes the future, or it can be just a stepstone for a new technology to emerge. But definetly, privacy and security plays already a huge role in our society, and this will not diminish in any close future.

\subsubsection{Conclusion}

In this paper we could see how SMPC works in general, and, especially, how SCALE-MAMBA works and how it evolved over the years in comparison to its predecessor, SPDZ. It was shown, how the protocol is split into a primary offline phase in which the computational expensive calculations are executed and all the preparations are done for the later circuit evaluation, like calculating the Beaver Triples, and how afterwards in the online phase the circuit is evaluated, how the triples are used to execute a multiplication without sharing the secret inputs and how the results can be verified at the end.

\renewcommand{\bibsection}{\section*{References}} % requried for natbib to have "References" printed and as section*, not chapter*
% Use natbib compatbile splncsnat style.
% It does provide all features of splncs03, but is developed in a clean way.
% Source: http://phaseportrait.blogspot.de/2011/02/natbib-compatible-bibtex-style-bst-file.html
\bibliographystyle{splncsnat}
\begingroup
  \ifluatex
    %try to activate if bibliography looks ugly
    %\sloppy
  \else
    \microtypecontext{expansion=sloppy}
  \fi
  \small % ensure correct font size for the bibliography
  \bibliography{paper}
\endgroup

% Enfore empty line after bibliography
\ \\
%

\end{document}

### 2019-03-11 paper.tex
Added online phase/ only thing missing is the improvemetn with different beaver triples.

@@
 \subsubsection{Oblivious Transfer}
+While Oblivious Transfer (OT) is not of direct importance for this thesis, the general concept should still be known to the reader.
 
@@
 Choosing the right access structure for the right use case is really important, because it has a realy big impact on performance and on the desired security level.
-
-TODO: SOUNDNESS SLACK
-TODO: LATTICE BASED CRYPTOGRAPHY
-
-Scale-Mamba is an implementation of SPDZ among other protocols. SPDZ is right now the most efficient protocol in use.
-
-Most LSSS-based MPC protocols split the computation into two parts: an of-
-
-ine phase, in which parties interact using expensive" public-key cryptography
-to lay the groundwork for an online phase in which only cheap" informationtheoretic
-primitives are required. The online phase is where the actual circuit
-evaluation takes place. For the access structures considered in this work, namely
-Q2 structures, the offline phase is almost as fast as the online phase. Thus the
-goal here is to minimize the cost of communication in both phases.
 
@@
 This includes already calculating values for possible multiplications and MACs, so that the online phase can be executed faster.
+In this phase the expensive calculations are done, like public key cryptography or authentifications, so that the online phase only needs basic primitives.
 This phase is not completly offline, because the parties still exchange some information which are later needed for the evaluation.
@@
 So this is the heart of the system, the part where the real computatiuon is executed.
+\\
 It is important to note, that the online phase is nearly perfectly optimized when assuming the offline phase was carried out succesfully.
@@
 $$c*\langle a\rangle=\langle c*a\rangle=(c*\delta,(c*a_1,...,c*a_n),,(c*\gamma (a)_1,...,c*\gamma (a)_n)$$
-\textbf{Correctness}:(TO BE DONE\\
+\textbf{Correctness}: We have to examine the different parts, to see each parts correctness: \\
+
+\begin{enumerate}
+\item \(  c*a = c * (a_1+....+a_n) = c*a_1+...+c*a_n \)
+\item \( \gamma (c*a) = \alpha*(c*a) = c*(\alpha * a) = c * \gamma (a) = c* ( \gamma (a)_1+...+\gamma (a)_n)  \)
+\end{enumerate}
 

### 2019-04-13 paper.tex
Sat work

@@
 \begin{abstract}
-  This thesis takes a look at SMPC especally using SCALE-MAMBA.
-For that it introduces first what SMPC is,  how SPDZ the precessor of SCALE-MAMBA works and what are the differences.
+This paper takes a look at Secure Multi-Party Computation(SMPC), especially using SCALE-MAMBA.
+For that, it introduces first what SMPC is, how SPDZ, the predecessor of SCALE-MAMBA, works and what are the differences.
 It discusses the functionalities of SCALE-MAMBA, especially the split into online and offline phase, how the beaver tripels are used for sacrificing and in which complexity classes the calculations are done.
@@
 
-In todays world, and even more in the future, there are a lot of computations that have to be done in a distributed environment, this might be because there are different parties involved, but also because the amount of data is getting so big that it becomes impossible to have a single entity that processes all the data at once.
-Furthermore, in a lot of these cases the parties participating in these computations don't want to reveal their own information, because its sensitive or because they don't want to share information with a lot of business value.\\
-A basic example is an anonymized voting.
-In most elections there is a third party that has to aggregate and count the votings made by the votings.
-For instance in germanys elections this is done by specific people counting a small group of votes and then forwarding the information.
-The other problem is that it has to be guaranteed that a vote is valid, meaning that the voter has the right to vote and is only voting one time.
-So if there would be a malicious person in the system, it would be easy for that person to figure out what vote a specific person made.\\
+In todays world, and even more in the future, there are a lot of computations that have to be done in a distributed environment, because in a lot of use cases multiple different parties are involved, but also because the amount of data is getting so big that it often becomes impossible to have a single entity that processes and stores all the data.
+Furthermore, in a lot of these cases the parties participating in these computations do not want to reveal their own information, because it is sensitive information or because they do not want to share information with a lot of business value.\\
+A basic example is an anonymous voting.
+In most elections there is a third party that has to aggregate and count the votes made by the voters.
+For instance, in Germany's elections this is done by specific people counting a small amount of votes and then forwarding the information to a central institution.
+The other problem is that it has to be guaranteed that a vote is valid, meaning that each voter holds the right to vote and is only voting one time.
+So if there would be a malicious person in this specific system, it would be easy for that person to gather information over a small subset of voters or even manipulate the local results.\\
 Another use case is the exchange of medical data.
-Medical data of patients is highly confidential and shoudn't be publiced under n circumstances.
-But at the same time a lot of different hospitals or doctors have only the data from their small data set of local patients, and for making diagnosis for new patients it is helpful to compare the symptoms to the diagnosis of other doctors.\\
+Medical data of patients is highly confidential and should never be published.
+But at the same time, a lot of different hospitals or doctors have only the data from their small data set of local patients, and for making founded diagnosis for new patients it is helpful to compare the symptoms to the diagnoses of other doctors.\\
 Secure Multi-Party Computation(SMPC) is one possible solution for these use cases. It is a field of research that emerged in the last 30 years, and is gathering a lot of speed in the last years. 
-The basic problem is how to evaluate a common function in which multiple parties have different inputs to this function, without revealing the different inputs of the parties.
-Although there are a lot of theoretical solutions for this problem, real-world implementations are still not perfmorming on the desired level.
-So in the last years the research shifted more in the area on how to make the existant solutions performant enough while still guaranteeing a desired level of security.\\
-This paper's main focus will lay on SCALE-MAMBA, a system developed mainly by the KU Leuven, which is one of the frontrunners in SMPC and a system that combines a lot of research by different teams with the newest ongoing research to provide a platform with highly performant SMPC.
-Nevertheless, this paper will at first give a short history of SMPC and will try to explain the basic technologies of SMPC to the reader.
-It will compare SCALE-MAMBA to its predecessors and will also try to give a short comparision to other SMPC systems which use different approaches, especially considering the usability in different use cases.
+The basic problem is how to evaluate a common function, with multiple parties having different inputs to this function, without revealing the different inputs of the parties.
+Although there are a lot of theoretical solutions for this problem, real-world implementations are still not performing on the desired level.
+So in the last years the research shifted more into the area of how to make the existent solutions performant enough while still guaranteeing a desired level of security.\\
+This paper's main focus will lay on SCALE-MAMBA, a system developed mainly by the Katholieke Universiteit Leuven, which is one of the forerunners in SMPC and a system that combines a lot of research by different teams with the newest ongoing research to provide a platform with highly performant SMPC.
+Nevertheless, this paper will at first give a short introduction into the history of SMPC and will try to explain its basic technologies to the reader.
+It will compare SCALE-MAMBA to its predecessors and will also try to give a short comparison to other SMPC systems which use different approaches, especially considering the usability in different use cases.
 Ultimately, it will try to give an outlook what might be coming in the next years.\\
@@
 
-Before diving deeper into SMPC, this section will shortly recap on cryptography basics that might be unknown to the reader.
+Before delving deeper into SMPC, this section will shortly recap on cryptography basics that might be unknown to the reader.
 These basics are not solely used by SMPC, but nevertheless, play a huge role in a lot of protocols.
@@
 \section{SMPC Basics and History}\label{sec:smpc}
+
 \subsection{Beginnings}
-As we saw in the introduction, there are a lot current and much more upcoming usecases of calculating the output of a function in a distributed environment with multiple parties while at the same time keeping all the individual inputs secret.
-The first explorations in these fields started in the 1970s, with theoretical problems like the mental poker problem(how can two players play poker while not being in the same physical location and making sure noone is cheating) and Yaos millionaire problem(how can can two millionaires decide who is richer than the other while not revealing their wealth).
+As we saw in the introduction, there are a lot current and much more upcoming use cases of calculating the output of a function in a distributed environment with multiple parties while at the same time keeping all the individual inputs secret.
+The first explorations in these fields started in the 1970s, with theoretical problems like the mental poker problem(how can two players play poker while not being in the same physical location and making sure no one is cheating) and Yaos millionaire problem(how can can two millionaires decide who is richer than the other while not revealing their wealth).
 \newline
 These explorations lead to a group of basic protocols which established the area of Secure Multi-Party Computation(SMPC).\\
+
+\subsection{Problem statement}
+Formally, the SMPC problem can be stated as follows:\\
+\begin{itemize}
+\item Given a finite field \( \mathcal{F}_p=\{0,...,p-1\}\) with the modulo operation regarding the prime p
+\item Given \(n\) parties \(P_1,...,P_n\) participating in the computation
+\item Given \(m\) secret inputs \(x_1,...,x_m\) which are each hold by one of the parties
+\item Given a function \(f(x_1,...,x_n)\) which depends on the secret inputs
+\item Goal: Evaluating the function \(f(x_1,...,x_n)\) while guaranteeing secrecy and correctness
+\end{itemize}
+Secrecy: If party \(P_i\) holds secret \(x_j\), then at the end of the evaluation no information regarding \(x_j\) is revealed to the other parties.\\
+Correctness: At the end of the computation, the result is correct.\\
+
  Now we will take a short look on the two most famous of these protocols, which both were the basis for a whole family of algorithms established in the years to follow until today.
@@
 
+\section{Implementations}
+Because SCALE-MAMBA was only released in 2018, there are not a lot documented usecases in praxis.
+Nevertheless, because of the structural and alogirthmic similarities to the earlier versions of SPDZ, taking a look on SPDZ implementations should give a general understanding of the possible usage of SCALE-MAMBA.\\
+
+\subsection{SPDZ 2.0 in medical healthcare}
+In medical healthcare, preserving privacy plays a huge role because patient records hold highly sensitive data, thus, its a natural area for the application of privacy protecting techonologies.
+
+https://www.researchgate.net/publication/328040632A\_New\_Approach\_to\_Privacy-Preserving\_Clinical\_Decision\_Support\_Systems\_for\_HIV\_Treatment
+https://arxiv.org/pdf/1901.00329.pdf
+https://www.researchgate.net/publication/324077995\_SEMBASEcure\_multi-biometric\_authentication
+
 \section{Comparison}\label{sec:comparison}
@@
 
+https://www.researchgate.net/publication/309817066\_Cheater\_Detection\_in\_SPDZ\_Multiparty\_Computation
+https://www.researchgate.net/publication/330091440\_Breaking\_MPC\_implementations\_through\_compression
+
 \subsubsection{Conclusion}

### 2019-04-13 paper.tex
sat eve

@@
 
-\section{Cryptography Basics}\label{sec:basics}
+\section{Cryptographic Basics}\label{sec:basics}
 
@@
 
-\subsection{Adversary types}
-Cryptography is the science of protecting information or communication from an adversary.
-This adversary is normally a third party that wants to get information for his own gain or want to sabotage something because of his own interest.
-In the are of SMPC we we normally divide between two types of adversaries.
-In SMPC we expect that everyone but ourselves can be controlled by an adversary.
-That means the adversary can control between one and all of the other parties with which we are communicating.
-The difference to the normal cryptographic adversary model is, that the adversary is one of the parties we are communicating with but not a third party that tries to intercept communication.
+\section{SMPC Basics and History}\label{sec:smpc}
+
+\subsection{Beginnings}
+As we saw in the introduction, there are a lot current and much more upcoming use cases of calculating the output of a function in a distributed environment with multiple parties while at the same time keeping all the individual inputs secret.
+The first explorations in these fields started in the 1970s, with theoretical problems like the mental poker problem(how can two players play poker while not being in the same physical location and making sure no one is cheating) and Yao's millionaire problem(how can can two millionaires decide who is richer than the other while not revealing their wealth).
+\\
+These explorations lead to a group of basic protocols which established the area of Secure Multi-Party Computation(SMPC).\\
+
+\subsection{Problem statement}
+Formally, the SMPC problem can be stated as follows:
+\begin{itemize}
+\item Given a finite field \( \mathcal{F}_p=\{0,...,p-1\}\) with a prime p
+\item Given \(n\) parties \(P_1,...,P_n\) participating in the computation
+\item Given \(m\) secret inputs \(x_1,...,x_m\) with each of them hold by one of the parties
+\item Given a function \(f(x_1,...,x_m)\)
+\item Goal: Evaluating the function \(f(x_1,...,x_m)\) while guaranteeing secrecy and correctness
+\end{itemize}
+\textbf{Secrecy:} If party \(P_i\) holds secret \(x_j\), then at the end of the evaluation no information regarding \(x_j\) is revealed to the other parties.\\
+\textbf{Correctness:} At the end of the computation, the result is correct.\\\\
+In a lot of cases each party has exactly one corresponding secret input, but this is not necessarily required. There can also be parties without input and parties with multiple inputs.\\
+Now we will take a short look on two of the first protocols, which both are the basis for a whole family of SMPC schemes.
+
+\subsection{Yao's Garbled Circuit}
+There are typically two families, into which most of the SMPC protocols can be categorized.
+One is based on garbled circuits which were first introduced by Yao in 1986 \cite{4568207}.
+The basic idea is that, in a two party setting, one party, Alice, garbles the boolean circuit that has to be evaluated and sends it to the second party, Bob.
+Bob uses his own input, Alice's encrypted input and the garbled circuit to evaluate the circuit in an encrypted way.
+At the end they exchange the results and can compute the result this way.
+
+\subsection{Shamirs Secret Sharing Scheme}
+This paper will not further discuss the family of garbled circuit protocols, but rather the second family which are protocols based on so called Linear Secret Sharing. 
+These type of protocols were introduced by Adi Shamir at a similar time, in 1979 \cite{Shamir:1979:SS:359168.359176}.\\
+The basic idea of this approach is that multiple parties split their secrets into so called shares and then distribute them to the other parties.
+In Shamir's scheme this is done by generating random polynomials with a degree depending on the security level and then calculating the shares as points in this polynomial.
+Computations are executed on these shares, and at the end it is possible to reconstruct the result from the different shares without revealing information about the initial secrets.
+This is done by using Lagrange interpolation.\\
+ There is a huge family of protocols build on this approach, and it is still evolving every day.
+This paper will focus on the newest of these protocols that were developed in the last 10 years, especially on the SPDZ like protocols.\\
+
+Before looking into exact implementations of linear secret sharing schemes, following, a short look will taken at the general adversary model and required basic definitions.
+
+\subsection{Adversary Model}
+In cryptography, the adversary which is considered is normally a third party that wants to get information for his own gain or want to sabotage something because of his own interest.
+In SMPC it is expected that the adversary is not only participating as one party, but could could control multiple parties.
+That means a potential adversary can control between one and all but one of the parties which are participating.
+The difference to the normal cryptographic adversary model is, that the adversary is controlling parties that are participating in the scheme and not a third party that tries to intercept communication.\\
+In the area of SMPC there is normally a differentiation between two types of adversaries.
+
 \subsubsection{Honest but curious}
-This type of adversary tries to gain as much information as possible.
-But this adversary is obeying to the rules of our protocol, that means it is impossible to know that this party is controlled by a third entity.
+The honest but curious adversary tries to gain as much information as possible while obeying to the rules of the protocol.
+Hence, it is impossible to know that this party is controlled by a third entity.
 This adversary tries to retrieve some of the secret information from the other parties, by using all the information he gets during the execution of the protocol.
@@
 While this adversary also wants to gain all the information, he is not obeying to the rules of the protocol.
-This type of adverary can send different inputs to manipulate the results he receives.
+This type of adversary can send different inputs to manipulate the results he receives.
 Furthermore, he wants to interrupt the whole protocol.
@@
 
-\section{SMPC Basics and History}\label{sec:smpc}
-
-\subsection{Beginnings}
-As we saw in the introduction, there are a lot current and much more upcoming use cases of calculating the output of a function in a distributed environment with multiple parties while at the same time keeping all the individual inputs secret.
-The first explorations in these fields started in the 1970s, with theoretical problems like the mental poker problem(how can two players play poker while not being in the same physical location and making sure no one is cheating) and Yaos millionaire problem(how can can two millionaires decide who is richer than the other while not revealing their wealth).
-\newline
-These explorations lead to a group of basic protocols which established the area of Secure Multi-Party Computation(SMPC).\\
-
-\subsection{Problem statement}
-Formally, the SMPC problem can be stated as follows:\\
-\begin{itemize}
-\item Given a finite field \( \mathcal{F}_p=\{0,...,p-1\}\) with the modulo operation regarding the prime p
-\item Given \(n\) parties \(P_1,...,P_n\) participating in the computation
-\item Given \(m\) secret inputs \(x_1,...,x_m\) which are each hold by one of the parties
-\item Given a function \(f(x_1,...,x_n)\) which depends on the secret inputs
-\item Goal: Evaluating the function \(f(x_1,...,x_n)\) while guaranteeing secrecy and correctness
-\end{itemize}
-Secrecy: If party \(P_i\) holds secret \(x_j\), then at the end of the evaluation no information regarding \(x_j\) is revealed to the other parties.\\
-Correctness: At the end of the computation, the result is correct.\\
-
- Now we will take a short look on the two most famous of these protocols, which both were the basis for a whole family of algorithms established in the years to follow until today.
-
-\subsection{Yaos Garbled Circuit}
-There are typically two groups of SMPC protocols, in which most of the protocols can be categorized.
-One is based on garbled circuits which were first introduced by Yao in 1986 \cite{4568207}.
-The basic idea is, that, in a two party setting, one party, Alice, garbles the boolean circuit that has to be evaluated and sends it to the second party, Bob.
-Bob uses his own input, Alice's encrypted input and the garbled circuit to evaluate the circuit in an ecnrypted way.
-At the end they exchange the results and can compute the result this way.
-
-\subsection{Shamirs Secret Sharing Scheme}
-This paper will not focus more on the family of garbled circuit protocols, but rather on the second family which are based on a so called Linear Secret Sharing Scheme(LSSS). 
-These type of protocols were introduced by Adi Shamir in a similiar time, in 1979 \cite{Shamir:1979:SS:359168.359176}.\\
-The basic idea of this approach is that multiple parties split their secrets into so called shares and then distribute them to the other parties.
-This is done by generating random polynomials with a degree depending on the security level and then calculating the shares as points in this polynomial.
-Computations are executed on these shares, and after the end it is possible to reconstruct the result from the different shares without revealing information about the initial secrets.
-This is done by using Lagrange interpolation.\\
- There is a huge family of protocols build on this approach, and it is still evolving every day.
-This paper will focus on the newest of these protocols that were developed in the last 10 years, especially on the SPDZ like protocols.\\
-In SMPC, normally all operations are executed in a finite field \(F_p = \{ 0,...,p-1\}\) with a prime p.
-That means all operations use modulo p at the end on the result to keep the result in the field.
-Therefore, for now on \(F_p\) will be used and when using an exact application, then p has to be specified depending on the usecase.\\
-
-\subsection{Terminology}
-Over the years different terminologies came up in the are of SMPC. Following will be a short introduction into some of them, so their definitions are clear.
-
-\subsubsection{Oblivious Transfer}
-While Oblivious Transfer (OT) is not of direct importance for this thesis, the general concept should still be known to the reader.
-
 \subsubsection{Access structures}
 There are a lot of different use cases for SMPC, so of course the adversary model changes depending on the system, the security needs, and the amount of parties involved.
-it is important to differentiate between these different use cases, because the different needs have big implications for security measurements that have to be taken and for runtime complexity.
-SMPC should run as fast as possible, while still providing all the neccesary security guarantees.
-
-To describe which party has which rights, and how many parties can be corrupted so the security can still be guaranteed, access structures were introduced.
+it is important to differentiate between these different use cases, because the different needs have big implications for security measurements that have to be taken and, thus, for runtime complexity.
+SMPC should run as fast as possible, while still providing all the necessary security guarantees.\\
+To describe which party has which rights, and how many parties can be corrupted, so that the security can still be guaranteed, access structures were introduced.
 An access structure describes how many parties are needed to form a so called qualified set.
-A qualified set is a group of parties which can pool their information to extract secret information..
-An access structure is called \(Q_l \) if at least l+1 parties are needed to  form a qualified set.
-An (n,t)-threshold scheme with t<n/l means that at least t+1 parties are needed to form a qualified set.
-Here, n is the number of total parties, and t is number of possible adversaries while not revealing any secrets.
+A qualified set is a group of parties which can pool their information to extract secret information.
+An access structure is called \(Q_l \)-access structure if at least \(l+1\) parties are needed to  form a qualified set.
+An \((n,t)\)-threshold scheme with \(t<\frac{n}{l}\) means that at least \(t+1\) parties are needed to form a qualified set.
+Here, \(n\) is the number of total parties, and \(t\) is number of possible adversaries while not revealing any secrets.\\
 A full threshold scheme means that all parties are needed to form a qualified set.
 This implies, that, if all other parties except one are corrupted by an adversary, this one party can still rely on the fact that it secret remains secret.
-Of course this has a huge impact on overhead and performance, but in some cases it can give needed security guarantees.
-An access structure also doesn't have to be symmetric.
-In some cases there might be different type of parties, where different parties have different positions or levels of trust, so it is possible to specify that some set of parties only has to have the size of 2 to form a qualified set and another one needs the size of 4.\\
+Of course this has a huge impact on overhead and performance, but in some cases it can provide needed security guarantees.
+An access structure also does not have to be symmetric.
+In some cases there might be different type of parties, where different parties have different positions or levels of trust, so it is, for example, possible to specify that some set of parties only has to have the size of 2 to form a qualified set and another one needs the size of 4.\\
 Choosing the right access structure for the right use case is really important, because it has a realy big impact on performance and on the desired security level.
@@
 
-In the last years, the research in SMPC is constantly ongoing, and there are new approaches or improvements to old approaches appearing regulary.
+In the last years, research in SMPC is constantly ongoing, and there are new approaches or improvements to old approaches appearing regulary.
 But most of these publications only focus on a small area of the whole picture.
 SCALE-MAMBA\cite{ScaleMambaDocu} is a project that tries to combine the state-of-the-art research into one system.
-That means, that it is offering kind of an end-to-end system for SMPC.
-It is a research system that is continously in development, therefore, there is not a single final version that can be examined.
-It offers as well a wide range of protocols for different usecases, i.e.
-it is possible to use the system in different confidurations.
-That also means for this thesis, that there is not a single process that can be examined.
-Therefore, this thesis tries to give more an overview how the process works in general, which are the main, especially which are newest and best, parts and where are the problems and bottlenecks.\\
+That means that it is offering an end-to-end system for SMPC.\\
+It is a research system that is continuously in development, therefore, there is not a single final version that can be examined.
+It offers a wide range of protocols for different use cases, i.e.
+it is possible to use the system in different configurations.
+That also means for this paper, that there is not a single protocol that can be examined.
+Thus, this paper tries to give more of an overview how the process works in general, which are the main, especially which are newest and best, components and where are the problems and bottlenecks.\\
 SCALE-MAMBA consists of two parts.
-The one part is called Secure Computation Algorithms from LEuven (SCALE) and is the SMPC part.
+One part is called Secure Computation Algorithms from LEuven (SCALE) and is the SMPC part.
 It contains all the computational logic and algorithms for multi-party computations.
@@
 
-In general, the SMPC protocol is split in two parts, first an offline phase and secondly and online phase.
+In general, the SCALE protocol is split in two parts, first an offline phase and afterwards an online phase.
 In the online phase only the final evaluation of the function is executed.

### 2019-04-14 paper.tex
small progress

@@
  There is a huge family of protocols build on this approach, and it is still evolving every day.
-This paper will focus on the newest of these protocols that were developed in the last 10 years, especially on the SPDZ like protocols.\\
+This paper will focus on the newest of these protocols that were developed in the last 10 years, especially on the protocols which are based on or similiar to SPDZ.\\
 
@@
 
+\subsubsection{Secret sharing}
+
+As mentioned before, SCALE-MAMBA, and in general similiar protocols like SPDZ, use linear secret sharing.\\
+Given party \(P_i\) which holds a secret value \( a \in \mathcal{F}_p\) and a share of the global MAC key \( \alpha\).
+Then \( P_i\) can additvely share \( a\) with the other \(n-1\) parties as follows:
+\begin{itemize}
+\item Choose random \(\delta_a \in \mathcal{F}_p\)
+\item Compute tag $ \gamma (a)= \alpha(a+\delta)$
+\item Choose random \(a_1,...,a_n \in \mathcal{F}_p\) so that $ a=(a_1+...+a_n)$
+\item Choose random \((\gamma (a)_1+...+\gamma (a)_n) \in \mathcal{F}_p\) so that $\gamma (a)=((\gamma (a)_1+...+\gamma (a)_n))$
+\item Send to each party \( P_j \) \( \delta_a \) and the shares \(a_j \) and, \( \gamma (a)_j \)
+\item Now party \( P_j\) holds a share of \( a\) as tuple \(\langle a_i\rangle = (\delta_a,a_i,\gamma(a)_i )\)
+\end{itemize}
+
+In SCALE-MAMBA, this value is then represented as followin tuple:\\
+$\langle a\rangle=(\delta_a, (a_1,...,a_n),(\gamma (a)_1,...,\gamma (a)_n))$
+\\With:
+\begin{itemize}
+\item $ a=(a_1+...+a_n)$
+\item $ \gamma (a)=((\gamma (a)_1+...+\gamma (a)_n))=\alpha(a+\delta)$
+\item Every party \( P_i\) holds a share of \( a\) as tuple \(\langle a_i\rangle = (\delta_a,a_i,\gamma(a)_i )\)
+
+\end{itemize}
+
+ and the corresponding tag \( \gamma(a)\). Then th
+Sharing of values:\\
+After sharing a value \( a \in F_p\), it is represented as following tuple:\\
+$\langle a\rangle=(\delta_a, (a_1,...,a_n),(\gamma (a)_1,...,\gamma (a)_n))$
+\\With:
+\begin{itemize}
+\item $ a=(a_1+...+a_n)$
+\item $ \gamma (a)=((\gamma (a)_1+...+\gamma (a)_n))=\alpha(a+\delta)$
+\item Every party \( P_i\) holds a share of \( a\) as tuple \(\langle a_i\rangle = (\delta_a,a_i,\gamma(a)_i )\)
+
+\end{itemize}
+
+Partially opening  of values: All players send their respective shares of a value, but not the MAC value, to one chosen party \(P_i\) which summs up all values and then broadcasts the result.
+In praxis this party is changing for every opening to balance the workload.\\
+Fully opening of values: In this case all players broadcast their respective shares and the respective MACs to all players, so all players can verify the result.\\
+
 \subsubsection{Offline phase}
@@
 
-Now we assume that the offline phase was succesfully executed, so all preliminaries are given to execute the online phase.
+Now it is assumed that the offline phase was succesfully executed, so all preliminaries are given to execute the online phase.
 It's main goal is to evaluate a specified circuit with secret inputs by the different parties.
@@
 The online phase has a runtime complexity in O(n), and therefore there is not a really big performance impact.
-It is also noteworthy, that in the online phase only only computational and no information theoretical problems are solved.
+It is also noteworthy, that in the online phase only computational and no information theoretical problems are solved.
 All this leads to the fact, that there is not really any change in the online phase in the SPDZ like protocols in the last years, therefore the online phase is the same like in the earlier version of SPDZ  ~\cite{cryptoeprint:2011:535}.\\
@@
 Following, it is going to be demonstrated, how addition and multiplication are done in this environment.
-This is sufficient, because a group with addition and multiplication is already touring complete, so there is no need for any other operation.\\
+This is sufficient, because a group with addition and multiplication is already touring complete, so there is no need for any other operation.
+Furthermore, addition and multiplications with constants are also discussed.\\
 

### 2019-04-14 paper.tex
small progress

@@
 \item Send to each party \( P_j \) \( \delta_a \) and the shares \(a_j \) and, \( \gamma (a)_j \)
-\item Now party \( P_j\) holds a share of \( a\) as tuple \(\langle a_i\rangle = (\delta_a,a_i,\gamma(a)_i )\)
+\item Now each party \( P_j\) holds a share of \( a\) as tuple \(\langle a_i\rangle = (\delta_a,a_i,\gamma(a)_i )\)
 \end{itemize}
+
+Now the value of \(a\) and the corresponding tag \(\gamma(a)\) can only be reconstructed if, and only if, all \(n\) secret shares are known.
+At the same time, no bit of information regarding \(a\) is revealed, as long as not all secret shares are known.\\
+Correctness: Assume that $n-1$ secret shares are known to an adversary, so all secret shares except one $a_x \in \mathcal{F}_p $.
+Then the adversary can calculate SUM=A-A_x.
+Nevertheless, because $a_x$ was drawn from an independent distribution, knowing SUM reveals not a single bit of information about $a$.\\
 

### 2019-04-16 paper.tex
Midday commit

--- 2019-04-14 paper.tex
@@
 Then \( P_i\) can additvely share \( a\) with the other \(n-1\) parties as follows:
-\begin{itemize}
+\begin{enumerate}
 \item Choose random \(\delta_a \in \mathcal{F}_p\)
@@
 \item Send to each party \( P_j \) \( \delta_a \) and the shares \(a_j \) and, \( \gamma (a)_j \)
-\item Now party \( P_j\) holds a share of \( a\) as tuple \(\langle a_i\rangle = (\delta_a,a_i,\gamma(a)_i )\)
-\end{itemize}
+\item Now each party \( P_j\) holds a share of \( a\) as tuple \(\langle a_j\rangle = (\delta_a,a_j,\gamma(a)_j )\)
+\end{enumerate}
 
-In SCALE-MAMBA, this value is then represented as followin tuple:\\
-$\langle a\rangle=(\delta_a, (a_1,...,a_n),(\gamma (a)_1,...,\gamma (a)_n))$
-\\With:
-\begin{itemize}
-\item $ a=(a_1+...+a_n)$
-\item $ \gamma (a)=((\gamma (a)_1+...+\gamma (a)_n))=\alpha(a+\delta)$
-\item Every party \( P_i\) holds a share of \( a\) as tuple \(\langle a_i\rangle = (\delta_a,a_i,\gamma(a)_i )\)
+Now the value of \(a\) and the corresponding tag \(\gamma(a)\) can only be reconstructed if, and only if, all \(n\) secret shares are known.
+At the same time, no bit of information regarding \(a\) is revealed, as long as not all secret shares are known.\\
+\textbf{Correctness:} Assume that $n-1$ secret shares are known to an adversary, so all secret shares except one $a_x \in \mathcal{F}_p $.
+Then the adversary can calculate $\sum_{i \neq x} a_i = a-a_x$.
+Nevertheless, because $a_x$ was drawn from an independent distribution and the adversary has no information over the value of \(a_x\), knowing $a-a_x$ reveals not a single bit of information about $a$.\\
 
-\end{itemize}
+\subsubsection{Revealing secrets}
 
- and the corresponding tag \( \gamma(a)\). Then th
-Sharing of values:\\
-After sharing a value \( a \in F_p\), it is represented as following tuple:\\
-$\langle a\rangle=(\delta_a, (a_1,...,a_n),(\gamma (a)_1,...,\gamma (a)_n))$
-\\With:
-\begin{itemize}
-\item $ a=(a_1+...+a_n)$
-\item $ \gamma (a)=((\gamma (a)_1+...+\gamma (a)_n))=\alpha(a+\delta)$
-\item Every party \( P_i\) holds a share of \( a\) as tuple \(\langle a_i\rangle = (\delta_a,a_i,\gamma(a)_i )\)
+For obvious reasons, the secrets also have to be revealed again. SCALE-MAMBA uses two different types of revealing them.\\
 
-\end{itemize}
-
-Partially opening  of values: All players send their respective shares of a value, but not the MAC value, to one chosen party \(P_i\) which summs up all values and then broadcasts the result.
-In praxis this party is changing for every opening to balance the workload.\\
-Fully opening of values: In this case all players broadcast their respective shares and the respective MACs to all players, so all players can verify the result.\\
+\textbf{Partial Reveal}
+All players send their respective shares of a value, but not the tag share, to one chosen party \(P_i\) which summs up all values and then broadcasts the result.
+The tag stays secretly shared between the parties.\\
+In praxis, the party computing the value is changing for every reveal to keep the workload balanced.\\
+\textbf{Full Reveal}
+In this case all players broadcast their respective shares and the respective MACs to all players, so all players can calculate the secret and verify the result with the corresponding tag.\\
 
@@
 \item There is a global MAC key \( \alpha\) that is secret to every party and additively shared, so that every party \( P_i\) has a share  \( \alpha_i\) so that  \( \alpha=\alpha_1+...+\alpha_n\)
-\item Each of the secret inputs to \(f\) is additively shared between the parties
-
+\item There are $m$ secret inputs $a_1,...,a_m$ which are each hold by exactly one party
 \end{enumerate}
 
+First, the secret inputs  $a_1,...,a_m$ are additvely shared with the other parties as described before in [Link to secret sharing in SM XXX].
+So now, each party $P_j$ holds a share ${(a_i)}_j$ of every secret $a_i$.\\
+
 Following, it is going to be demonstrated, how addition and multiplication are done in this environment.
-This is sufficient, because a group with addition and multiplication is already touring complete, so there is no need for any other operation.
-Furthermore, addition and multiplications with constants are also discussed.\\
-
-Sharing of values:\\
-After sharing a value \( a \in F_p\), it is represented as following tuple:\\
-$\langle a\rangle=(\delta_a, (a_1,...,a_n),(\gamma (a)_1,...,\gamma (a)_n))$
-\\With:
-\begin{itemize}
-\item $ a=(a_1+...+a_n)$
-\item $ \gamma (a)=((\gamma (a)_1+...+\gamma (a)_n))=\alpha(a+\delta)$
-\item Every party \( P_i\) holds a share of \( a\) as tuple \(\langle a_i\rangle = (\delta_a,a_i,\gamma(a)_i )\)
-
-\end{itemize}
+These two operations are sufficient, because a group with addition and multiplication is already touring complete, so there is no need for any other operation[TODO: ADD reference].
+Furthermore, addition and multiplications with constants are also discussed, because they can be done in a simple way that saves computation time.\\
 
 \subsubsection{Addition}
-Assumed there are two values \( a\) and \( b\) which should be added together and both are secret inputs from two different parties.
+Assumed there are two values \( a\) and \( b\) which should be added together, and both are secret inputs from two different parties.
 Naturally, both of these values are additively shared between all the parties, i.e.
@@
 $$\langle c_i\rangle=\langle a_i\rangle+\langle b_i\rangle=((\delta_a+\delta_b),(a_i+b_i),\gamma (a)_i+\gamma (b)_i))$$
-\textbf{Correctness}: Afterwards, \( c=a+b\) is obviously additively shared between the parties and can be used for further computations.\\
+\textbf{Correctness:} Afterwards, \( c=a+b\) is obviously additively shared between the parties and can be used for further computations.[TODO: Explain correctness more in detail\\
 
 \subsubsection{Addition with constant}
-Adding a public constant is really simple in this representation. Because \( \gamma (a)=\alpha(a+\delta)\), adding a constant \(c\) to a additively shared value \(a\) is done by:\\
+Adding a public constant $c$ to a secretly shared value $ \langle a\rangle $ is really simple in this representation.
+Because \( \gamma (a)=\alpha(a+\delta)\), adding a constant \( c \) to a additively shared value \( a \) is done by:\\
 $$\langle a\rangle+c=(\delta-c,(a_1+c,a_2,...,a_n),,(\gamma (a)_1,...,\gamma (a)_n)$$
-Correctness: It can be seen, that the constant is only added to one shared value and subtracted from \( \delta\).
+\textbf{Correctness:} It can be seen, that the constant is only added to one shared value and subtracted from \( \delta\).
 Because of that, \( \alpha(a+\delta)\) stays the same, so the same shared MAC values like before can be used.
-So \(a+c\) is also properly additively shared.\\
+So \( \langle a+c \rangle \) is also properly additively shared.\\
 

### 2019-04-17 paper.tex
xx

@@
 \item There are $m$ secret inputs $a_1,...,a_m$ which are each hold by exactly one party
+\item There is a sufficiently large queue of beaver triples which are secretly shared
 \end{enumerate}
@@
 $$\langle c_i\rangle=\langle a_i\rangle+\langle b_i\rangle=((\delta_a+\delta_b),(a_i+b_i),\gamma (a)_i+\gamma (b)_i))$$
-\textbf{Correctness:} Afterwards, \( c=a+b\) is obviously additively shared between the parties and can be used for further computations.[TODO: Explain correctness more in detail\\
+\textbf{Correctness:} Afterwards, \( c=a+b\) is additively shared between the parties and can be used for further computations, because when fully revealing $\langle c\rangle$ following statements hold:
+\begin{enumerate}
+\item $a+b=(a_1+...+a_n)+(b_1+...+b_n)=(a_1+b+1)+...+(a_n+b_n)=c$
+\item $\delta_a+\delta_b=\delta_c$
+\item $\gamma(a)+\gamma(b)=\alpha*(\delta_a+a)+\alpha*(\delta_b+b)=\alpha*(\delta_a+\delta_b+a+b)=\alpha*(\delta_c+c)=\gamma(c)$
+\end{enumerate}
 
 \subsubsection{Addition with constant}
-Adding a public constant $c$ to a secretly shared value $ \langle a\rangle $ is really simple in this representation.
-Because \( \gamma (a)=\alpha(a+\delta)\), adding a constant \( c \) to a additively shared value \( a \) is done by:\\
+Adding a public constant $c$ to a secretly shared value $ \langle a\rangle $ is really simple in this representation, because \( \gamma (a)=\alpha(a+\delta)\) holds:\\
 $$\langle a\rangle+c=(\delta-c,(a_1+c,a_2,...,a_n),,(\gamma (a)_1,...,\gamma (a)_n)$$
-\textbf{Correctness:} It can be seen, that the constant is only added to one shared value and subtracted from \( \delta\).
-Because of that, \( \alpha(a+\delta)\) stays the same, so the same shared MAC values like before can be used.
-So \( \langle a+c \rangle \) is also properly additively shared.\\
+\textbf{Correctness:} It can be seen, that the constant is only added to one share $a_1$ of the value and subtracted from \( \delta_a\). \( \langle a+c \rangle \) is properly secretly shared because:
+\begin{itemize}
+\item $a+c=(a_1+...+a_n)+c=(a+1+c)+a_2+...+a_n$
+\item \( \gamma(a)=\alpha*(a+\delta_a)=\alpha*((a+c)+(\delta_a-c))=\gamma(a+c)\)
+\end{itemize}
 
 \subsubsection{Multiplication with constant}
-Multiplicating with a public constant is as well really simple and can also be done by each party invdividually. It is defined as:\\
-$$c*\langle a\rangle=\langle c*a\rangle=(c*\delta,(c*a_1,...,c*a_n),,(c*\gamma (a)_1,...,c*\gamma (a)_n)$$
-\textbf{Correctness}: We have to examine the different parts, to see each parts correctness: \\
-
+Multiplication with a public constant is as well really simple, and can also be done by each party individually without any communication between the parties. It is defined as:\\
+$$c*\langle a\rangle=\langle c*a\rangle=(c*\delta,(c*a_1,...,c*a_n),(c*\gamma (a)_1,...,c*\gamma (a)_n)$$
+\textbf{Correctness}: We have to examine the different parts, to see each parts correctness:
 \begin{enumerate}
 \item \(  c*a = c * (a_1+....+a_n) = c*a_1+...+c*a_n \)
-\item \( \gamma (c*a) = \alpha*(c*a) = c*(\alpha * a) = c * \gamma (a) = c* ( \gamma (a)_1+...+\gamma (a)_n)  \)
+\item \( c*\gamma (a)_1+...+c*\gamma (a)_n=c*\gamma (a) = c*(\alpha*(a+\delta_a)) = \alpha*(c*a+c*\delta_a))=\gamma(c*a)\)
 \end{enumerate}
@@
 
-Multiplication is the only operation,except for distributing the shares and calculating and verifying the end result, where the parties have to interact in the online phase.
-Assumed there are two values \( a\) and \( b\) which should be multipled with each other. Naturally, each party has the shares \(\langle a_i\rangle \) and \(\langle b_i\rangle \).\\
+Multiplication is the only operation, except for distributing the shares and calculating and verifying the end result, where the parties have to interact in the online phase.
+Assumed there are two secretly shared values \( \langle a \rangle\) and \( \langle b \rangle\) which should be multiplied with each other.
+Naturally, each party has the shares \(\langle a_i\rangle \) and \(\langle b_i\rangle \).\\
 For each multiplication operation, we need one of the Beaver triples \( (\langle x_1 \rangle ,\langle y_1\rangle , \langle z_1\rangle )\) generated in the offline phase.
@@
 \end{enumerate}
-
-Partially opening  of values: All players send their respective shares of a value, but not the MAC value, to one chosen party \(P_i\) which summs up all values and then broadcasts the result.
-In praxis this party is changing for every opening to balance the workload.\\
-Fully opening of values: In this case all players broadcast their respective shares and the respective MACs to all players, so all players can verify the result.\\
 
@@
 
-Now in the next step the multiplication of \(\langle a \rangle \) and \(\langle b \rangle \) is executed with the help of the triple.
+Multiplication is the only operation, except for distributing the shares and calculating and verifying the end result, where the parties have to interact in the online phase.
+Assumed there are two secretly shared values \( \langle a \rangle\) and \( \langle b \rangle\) which should be multiplied with each other.
+Naturally, each party $P_i$ holds the shares \(\langle a_i\rangle \) and \(\langle b_i\rangle \).\\
+For each multiplication operation, one of the Beaver triples \( (\langle x \rangle ,\langle y\rangle , \langle z\rangle )\) generated in the offline phase is needed, which is taken out of the queue.\\ The multiplication is done with following protocol:
 \begin{enumerate}
-\item The parties calculate \( \epsilon = \langle a \rangle - \langle x_1 \rangle \) and partially open it
-\item The parties calculate \( \delta = \langle b \rangle - \langle y_1 \rangle \) and partially open it
-\item The parties calculate \( \langle c \rangle \langle z_1 \rangle + \epsilon * \langle b \rangle + \delta * \langle a \rangle + \epsilon * \delta \) 
+\item The parties calculate \( \epsilon = \langle a \rangle - \langle x \rangle \) and partially open it
+\item The parties calculate \( \delta = \langle b \rangle - \langle y \rangle \) and partially open it
+\item The parties calculate \( \langle c \rangle \langle z \rangle + \epsilon * \langle b \rangle + \delta * \langle a \rangle + \epsilon * \delta \) 
 \end{enumerate}
 Correctness: 
-$$  \langle a \rangle * \langle b \rangle = \langle (a-x_1+x_1) \rangle * \langle (b-y_1+y_1) \rangle$$
-$$ = ( \epsilon  +\langle  x_1 \rangle ) * ( \delta  + \langle y_1 \rangle )$$
-$$ = \langle x_1 \rangle * \langle y_1 \rangle + \epsilon * \langle b \rangle + \delta * \langle a \rangle + \epsilon * \delta =  \langle c \rangle $$
+$$  \langle a \rangle * \langle b \rangle = \langle (a-x+x) \rangle * \langle (b-y+y) \rangle$$
+$$ = ( \epsilon  +\langle  x \rangle ) * ( \delta  + \langle y \rangle )$$
+$$ = \langle x \rangle * \langle y \rangle + \epsilon * \langle b \rangle + \delta * \langle a \rangle + \epsilon * \delta =  \langle c \rangle $$
 
-So, as we can see, the calculations are correct and at the same time no information of \(\langle a \rangle \) or \(\langle b \rangle \) was revealed.\\
+So, as we can see, the calculations are correct and at the same time no information about \(\langle a \rangle \) or \(\langle b \rangle \) was revealed.\\
 
 \subsubsection{Verifying Correctness}
-At the end of the computation, we have to find out the total calculated result.
-After all computations were executed, every party has a share of the final result \( r_i\), a share of the MAC value \( \gamma(r)_i)\) of the final result and a share of the global MAC key \( \alpha_i \).
+At the end of the computation, we have to find out the total result $r$.
+After all computations were executed, every party has a share of the final result \( r_i\), a share of the tag \( \gamma(r)_i)\) of the final result and a share of the global MAC key \( \alpha_i \).
 The verification is done as follows:
@@
 
-If the calculations were correct, the MAC was shared properly and not modified, and the MAC values were not modified, then the difference between both sums will be 0.
+If the calculations were correct, the MAC was shared properly and not modified, and the tag values were not modified, then the difference between both sums will be 0.
 If it is not 0, the protocol detects that there is a malicious adversary and aborts.\\
+
+\subsubsection{Rumtime Complexity}
+Addition: O(1)
+Mutliplication: O(1)
+Secret sharing: O(n) for every party.
+Be d the depth of the circuit (
 

### 2019-04-19 paper.tex
X

@@
 
-In the offline or pre-processing phase, the most important part is the generation of the beaver triples.
-These triples  are independent of the computation that is later performed in the online phase, but will be used in the online phase to guarantee that multiplications can be done without revealing the secret inputs of the different parties.
-Beaver triples are a tuple of three numbers \((a,b,c)\), with \( a,b,c\in \mathcal{F} )\), with \(a\) and \(b\) being chosen uniformly randomly, and \(a*b=c\).
-
-The second thing that is done in the offline phase is the generation of a global MAC key.
+There are two main things done in the offline phase.
+The first one is the generation of a global MAC key $\alpha$.
 This MAC key is later used, to authentificate the different shares at the end of the computation.
 This key is also distributedly generated, i.e.
-only if all parties pool together they can recover its value.\\
+only if all parties pool together their shares they can recover its value.\\
+
+The second and and most important part in the offline phase, is the generation of the beaver triples.
+These triples  are independent of the computation that is later performed in the online phase, but will be used in the online phase to guarantee that multiplications can be done without revealing the secret inputs of the different parties.
+For each multiplication, one of the triples is needed.
+
+\subsubsection{Beaver Triples}
+
+Beaver triples are triples of three numbers \((a,b,c)\), with \( a,b,c\in \mathcal{F}_p \), with \(a\) and \(b\) being chosen uniformly randomly, and \(a*b=c\).
+In the offline or pre-processing phase, the most important part is the generation of the beaver triples.
+These triples  are independent of the computation that is later performed in the online phase, but will be used in the online phase to guarantee that multiplications can be done without revealing the secret inputs of the different parties.\\
+
+For each multiplication operation in the online phase, we need one of the Beaver triples \( (\langle x_1 \rangle ,\langle y_1\rangle , \langle z_1\rangle )\) generated in the offline phase.
+In the offline phase we checked the quality of our triples, but we still allowed the possibility that each triple has an error \(e\) so that \(c=a*b+e\).
+To check the quality of our triple, we use a second triple \( (\langle x_2 \rangle ,\langle y_2\rangle ,\langle z_2\rangle )\) in a procedure called sacrificing it.
+Both triples are additively shared between all parties like the input values.
+In SPDZ both triples are completly unrelated to each other, and the sacrificing is done as follows:\\
+
+Sacrificing of triples
+Even though the triples are generated and each party holds the respective shares, there is still an allowed error \(e\) for each triple \( (\langle x_1 \rangle ,\langle y_1\rangle , \langle z_1\rangle )\) so that \(z_1=x_1*y_1+e\).
+To guarantee that the triples are really sound, the correctness has to be checked.
+This is done with a second triple \( (\langle x_2 \rangle ,\langle y_2\rangle ,\langle z_2\rangle )\) in a procedure called sacrificing, which uses the second triple to guarantee the soundness of the first.
+In SPDZ that is done as follows:\\
+\begin{enumerate}
+\item Choose a random variable \( r\) and open it to the other parties. This is done by each party generating a random share \(r_i\) and then calculating \(r=r_1+...+r_n\).
+\item The parties calculate  \( p=r*\langle x_1 \rangle - \langle x_2 \rangle \) and \( \sigma=\langle y_1 \rangle -  \langle y_2 \rangle \) and partially open the results afterwards
+\item The parties calculate \(\phi = r * \langle z_1 \rangle\ - \langle z_2 \rangle - \sigma * \langle x_2 \rangle - p * \langle y_2 \rangle - \sigma * p\) and then partially open the result
+\item If \( \phi \) is 0 then the triples are correct and it is proceeded with \( (\langle x_1 \rangle ,\langle y_1\rangle , \langle z_1\rangle )\) as a correct, secretly shared and not revealed triple.
+Otherwise the system detected a malicious adversary and it aborts.\\
+\end{enumerate}
+
+Remark: As shown later in chapter [X], addition of two secretly shared values and multiplication of a secretly shared value with a public constant can be done individually by every party, so that, when later revealing the results of each party, the correct result can be computed by the parties together.
+Accordingly, beforehand operations can be assumed as working as seen later.
+
+\textbf{Correctness}: Simplifying \( \phi \), it can be see that every term disappears (given that the triples are proper triples):
+$$\phi = t*z_1 - z_2 - y_1 + y_1*x_2 + y_2 * x_2 - t * x_1 * y_2 + x_2*y_2 - t*x_1*y_1 + t*x_1*y_2 + x_2*y_1 - x_2*y_2$$
+$$ = t*(z_1 - x_1*y_2+x_1*y_2 -x_1*y_1) + y_1*x_2 - x_2*y_1+ x_2*y_2 - z_2 + x_2*y_2 -  x_2*y_2$$
+$$= t* (z_1 - x_1*y_1) + x_2*y_2 - z_2 $$
+Therefore, \( \phi = 0 \Leftrightarrow x_1 * y_1 = z_1 \land x_2*y_2=z_2 \), so this only holds if both triples are correct. \\
+In this process the triple \( (\langle x_1 \rangle ,\langle y_1\rangle , \langle z_1\rangle ) \) wasn't revealed, so now it can be used for the multiplication knowing that one the one hand it is still secretly shared and unknown and on the other hand that it was not modified by an adversary.\\
+
+Sacrificing of triples in SCALE-MAMBA
+
+In later versions of SPDZ, since ~\cite{cryptoeprint:2017:123}, and in the current implementations of SCALE-MAMBA, a slightly modified version of aforementioned protocol is used.
+The idea is to not sacrifice a random second triple, to guarantee the correctness of the first one, but to sacrifice a similar triple.
+For this, the triples are directly generated in pairs, so that there are are two triples \( (\langle x \rangle ,\langle y \rangle , \langle z \rangle ) \) and \( (\langle x' \rangle ,\langle y\rangle , \langle z'\rangle ) \) with $z=x*y$ and $z'=x'*y$.
+With $y_1=y_2$, the formula from SPDZ becomes directly much simpler.
+\begin{enumerate}
+\item Choose a random variable \( r\) and open it to the other parties. This is done by each party generating a random share \(r_i\) and then calculating \(r=r_1+...+r_n\).
+\item The parties calculate  \( p=r*\langle x \rangle - \langle x' \rangle \) and partially open the result afterwards
+\item The parties calculate \(\phi = r * \langle z \rangle\ - \langle z' \rangle - p * \langle y \rangle\) and then partially open the result
+\item If \( \phi \) is 0 then the triples are correct and it is proceeded with \( (\langle x \rangle ,\langle y\rangle , \langle z\rangle )\) as a correct, secretly shared and not revealed triple.
+Otherwise the system detected a malicious adversary and it aborts.\\
+\end{enumerate}
+
+\textbf{Correctness}: Parallel to before, simplifying \( \phi \) leads to the same implications:\\
+$$\phi =r*z-z'-p*y=r*z-z'-(r*x-x')*y=r*(z-x*y)+(x'*y-z')$$
+Therefore, \( \phi = 0 \Leftrightarrow x * y = z \land x'*y=z' \)
+
+Comparison in Sacrificing
+This small improvement in the protocol nearly halved the mathematical operations necessary for the creation of one beaver triple.
+This is huge, because the triple generation is one of the most time consuming task, so it sped up the offline phase.
+In following table, the differences between both sacrificing protocols can be seen:
+
+\begin{center}
+\begin{tabular}{ | c | c | c | }
+\hline
+& SPDZ & SCALE-MAMBA \\
+\hline
+ Additions & 6 & 3 \\ 
+\hline
+ Multiplications with constants & 5 & 3 \\  
+\hline
+Partial reveals & 3 &2 \\
+\hline
+\end{tabular}
+\end{center}
+
+In SPDZ: Sacrifice in online phase???
 
@@
 Naturally, each party $P_i$ holds the shares \(\langle a_i\rangle \) and \(\langle b_i\rangle \).\\
-For each multiplication operation, one of the Beaver triples \( (\langle x \rangle ,\langle y\rangle , \langle z\rangle )\) generated in the offline phase is needed, which is taken out of the queue.\\ The multiplication is done with following protocol:
+For each multiplication operation, one of the secretly shared Beaver triples \( (\langle x \rangle ,\langle y\rangle , \langle z\rangle )\) generated in the offline phase is needed, which is taken out of the queue.\\ The multiplication is done with following protocol:
 \begin{enumerate}
@@
 \item The parties calculate \( \delta = \langle b \rangle - \langle y \rangle \) and partially open it
-\item The parties calculate \( \langle c \rangle \langle z \rangle + \epsilon * \langle b \rangle + \delta * \langle a \rangle + \epsilon * \delta \) 
+\item The parties calculate \( \langle c \rangle = \langle z \rangle + \epsilon * \langle b \rangle + \delta * \langle a \rangle + \epsilon * \delta \) 
 \end{enumerate}
-Correctness: 
-$$  \langle a \rangle * \langle b \rangle = \langle (a-x+x) \rangle * \langle (b-y+y) \rangle$$
+\textbf{Correctness:} 
+$$  \langle a \rangle * \langle b \rangle = \langle ( \langle a \rangle- \langle x \rangle+ \langle x \rangle) \rangle * \langle ( \langle b \rangle- \langle y \rangle+ \langle y \rangle) \rangle$$
 $$ = ( \epsilon  +\langle  x \rangle ) * ( \delta  + \langle y \rangle )$$
@@
 
-So, as we can see, the calculations are correct and at the same time no information about \(\langle a \rangle \) or \(\langle b \rangle \) was revealed.\\
+As it can be seen, the computations are correct and calculate the expected result, and at the same time no information about \(\langle a \rangle \) or \(\langle b \rangle \) is revealed.\\
 
-\subsubsection{Verifying Correctness}
-At the end of the computation, we have to find out the total result $r$.
+\textbf{Runtime:} With the help of the beaver triple, it was possible to express a multiplication as a combination of 5 additions, 2 multiplications with constants and 1 multiplication of two constants by each party.
+Additionally two values were partially opened, but this can be done in constant time too.
+So a multiplication can be done in \(O(1)\).\\
+
+\subsubsection{Determining the Result and Verifying Correctness}
+At the end of the circuit evaluation, the total result $r$ has to be computed.
 After all computations were executed, every party has a share of the final result \( r_i\), a share of the tag \( \gamma(r)_i)\) of the final result and a share of the global MAC key \( \alpha_i \).
@@
 
-\subsubsection{Rumtime Complexity}
-Addition: O(1)
-Mutliplication: O(1)
-Secret sharing: O(n) for every party.
-Be d the depth of the circuit (
+\subsubsection{Runtime Complexity}
+Because all the mathematical operations can be executed in O(1) given the preliminaries from the offline phase, the mathematical evaluation of the circuit can be executed in O(d) with d being the amount of operations the circuit consists of.
+Additionally, each party can share a secret in O(n), so the total complexity class of the online phase is linear.
+Therefore, the online phase is really well scalable.\\
 

### 2019-04-19 paper.tex
X

@@
 The first one is the generation of a global MAC key $\alpha$.
-This MAC key is later used, to authentificate the different shares at the end of the computation.
-This key is also distributedly generated, i.e.
+This MAC key is later used, to authenticate the different shares at the end of the computation.
+This key is also distributively generated, i.e.
 only if all parties pool together their shares they can recover its value.\\
@@
 Both triples are additively shared between all parties like the input values.
-In SPDZ both triples are completly unrelated to each other, and the sacrificing is done as follows:\\
+In SPDZ both triples are completely unrelated to each other, and the sacrificing is done as follows:\\
 
-Sacrificing of triples
-Even though the triples are generated and each party holds the respective shares, there is still an allowed error \(e\) for each triple \( (\langle x_1 \rangle ,\langle y_1\rangle , \langle z_1\rangle )\) so that \(z_1=x_1*y_1+e\).
-To guarantee that the triples are really sound, the correctness has to be checked.
-This is done with a second triple \( (\langle x_2 \rangle ,\langle y_2\rangle ,\langle z_2\rangle )\) in a procedure called sacrificing, which uses the second triple to guarantee the soundness of the first.
-In SPDZ that is done as follows:\\
-\begin{enumerate}
-\item Choose a random variable \( r\) and open it to the other parties. This is done by each party generating a random share \(r_i\) and then calculating \(r=r_1+...+r_n\).
-\item The parties calculate  \( p=r*\langle x_1 \rangle - \langle x_2 \rangle \) and \( \sigma=\langle y_1 \rangle -  \langle y_2 \rangle \) and partially open the results afterwards
-\item The parties calculate \(\phi = r * \langle z_1 \rangle\ - \langle z_2 \rangle - \sigma * \langle x_2 \rangle - p * \langle y_2 \rangle - \sigma * p\) and then partially open the result
-\item If \( \phi \) is 0 then the triples are correct and it is proceeded with \( (\langle x_1 \rangle ,\langle y_1\rangle , \langle z_1\rangle )\) as a correct, secretly shared and not revealed triple.
-Otherwise the system detected a malicious adversary and it aborts.\\
-\end{enumerate}
-
-Remark: As shown later in chapter [X], addition of two secretly shared values and multiplication of a secretly shared value with a public constant can be done individually by every party, so that, when later revealing the results of each party, the correct result can be computed by the parties together.
-Accordingly, beforehand operations can be assumed as working as seen later.
-
-\textbf{Correctness}: Simplifying \( \phi \), it can be see that every term disappears (given that the triples are proper triples):
-$$\phi = t*z_1 - z_2 - y_1 + y_1*x_2 + y_2 * x_2 - t * x_1 * y_2 + x_2*y_2 - t*x_1*y_1 + t*x_1*y_2 + x_2*y_1 - x_2*y_2$$
-$$ = t*(z_1 - x_1*y_2+x_1*y_2 -x_1*y_1) + y_1*x_2 - x_2*y_1+ x_2*y_2 - z_2 + x_2*y_2 -  x_2*y_2$$
-$$= t* (z_1 - x_1*y_1) + x_2*y_2 - z_2 $$
-Therefore, \( \phi = 0 \Leftrightarrow x_1 * y_1 = z_1 \land x_2*y_2=z_2 \), so this only holds if both triples are correct. \\
-In this process the triple \( (\langle x_1 \rangle ,\langle y_1\rangle , \langle z_1\rangle ) \) wasn't revealed, so now it can be used for the multiplication knowing that one the one hand it is still secretly shared and unknown and on the other hand that it was not modified by an adversary.\\
-
-Sacrificing of triples in SCALE-MAMBA
-
-In later versions of SPDZ, since ~\cite{cryptoeprint:2017:123}, and in the current implementations of SCALE-MAMBA, a slightly modified version of aforementioned protocol is used.
-The idea is to not sacrifice a random second triple, to guarantee the correctness of the first one, but to sacrifice a similar triple.
-For this, the triples are directly generated in pairs, so that there are are two triples \( (\langle x \rangle ,\langle y \rangle , \langle z \rangle ) \) and \( (\langle x' \rangle ,\langle y\rangle , \langle z'\rangle ) \) with $z=x*y$ and $z'=x'*y$.
-With $y_1=y_2$, the formula from SPDZ becomes directly much simpler.
-\begin{enumerate}
-\item Choose a random variable \( r\) and open it to the other parties. This is done by each party generating a random share \(r_i\) and then calculating \(r=r_1+...+r_n\).
-\item The parties calculate  \( p=r*\langle x \rangle - \langle x' \rangle \) and partially open the result afterwards
-\item The parties calculate \(\phi = r * \langle z \rangle\ - \langle z' \rangle - p * \langle y \rangle\) and then partially open the result
-\item If \( \phi \) is 0 then the triples are correct and it is proceeded with \( (\langle x \rangle ,\langle y\rangle , \langle z\rangle )\) as a correct, secretly shared and not revealed triple.
-Otherwise the system detected a malicious adversary and it aborts.\\
-\end{enumerate}
-
-\textbf{Correctness}: Parallel to before, simplifying \( \phi \) leads to the same implications:\\
-$$\phi =r*z-z'-p*y=r*z-z'-(r*x-x')*y=r*(z-x*y)+(x'*y-z')$$
-Therefore, \( \phi = 0 \Leftrightarrow x * y = z \land x'*y=z' \)
-
-Comparison in Sacrificing
-This small improvement in the protocol nearly halved the mathematical operations necessary for the creation of one beaver triple.
-This is huge, because the triple generation is one of the most time consuming task, so it sped up the offline phase.
-In following table, the differences between both sacrificing protocols can be seen:
-
-\begin{center}
-\begin{tabular}{ | c | c | c | }
-\hline
-& SPDZ & SCALE-MAMBA \\
-\hline
- Additions & 6 & 3 \\ 
-\hline
- Multiplications with constants & 5 & 3 \\  
-\hline
-Partial reveals & 3 &2 \\
-\hline
-\end{tabular}
-\end{center}
-
-In SPDZ: Sacrifice in online phase???
+GENERATION OF BEAVER TRIPLES
 
@@
 
-In the offline phase, ZPoKP are used to guarantee that the Beaver triples are calculated correctly, or atleast that they dont divert more than a specified error \(e\) the expected result.
+In the offline phase, ZPoKP are used to guarantee that the Beaver triples are calculated correctly before the sacrificing, or atleast that they dont divert more than a specified error \(e\) from the expected result.
 
@@
 It further improved the HighGear protocol of the Overdrive implementation, by targeting specifically the ZKPoKs.
+
+\subsubsection{Sacrificing of triples}
+Even though the triples are generated and each party holds the respective shares, there is still an allowed error \(e\) for each triple \( (\langle x_1 \rangle ,\langle y_1\rangle , \langle z_1\rangle )\) so that \(z_1=x_1*y_1+e\).
+To guarantee that the triples are really sound, the correctness has to be checked.
+This is done with a second triple \( (\langle x_2 \rangle ,\langle y_2\rangle ,\langle z_2\rangle )\) in a procedure called sacrificing, which uses the second triple to guarantee the soundness of the first.
+In SPDZ that is done as follows:\\
+\begin{enumerate}
+\item Choose a random variable \( r\) and open it to the other parties. This is done by each party generating a random share \(r_i\) and then calculating \(r=r_1+...+r_n\).
+\item The parties calculate  \( p=r*\langle x_1 \rangle - \langle x_2 \rangle \) and \( \sigma=\langle y_1 \rangle -  \langle y_2 \rangle \) and partially open the results afterwards
+\item The parties calculate \(\phi = r * \langle z_1 \rangle\ - \langle z_2 \rangle - \sigma * \langle x_2 \rangle - p * \langle y_2 \rangle - \sigma * p\) and then partially open the result
+\item If \( \phi \) is 0 then the triples are correct and it is proceeded with \( (\langle x_1 \rangle ,\langle y_1\rangle , \langle z_1\rangle )\) as a correct, secretly shared and not revealed triple.
+Otherwise the system detected a malicious adversary and it aborts.\\
+\end{enumerate}
+
+Remark: As shown later in chapter [X], addition of two secretly shared values and multiplication of a secretly shared value with a public constant can be done individually by every party, so that, when later revealing the results of each party, the correct result can be computed by the parties together.
+Accordingly, beforehand operations can be assumed as working as seen later.
+
+\textbf{Correctness}: Simplifying \( \phi \), it can be see that every term disappears (given that the triples are proper triples):
+$$\phi = t*z_1 - z_2 - y_1 + y_1*x_2 + y_2 * x_2 - t * x_1 * y_2 + x_2*y_2 - t*x_1*y_1 + t*x_1*y_2 + x_2*y_1 - x_2*y_2$$
+$$ = t*(z_1 - x_1*y_2+x_1*y_2 -x_1*y_1) + y_1*x_2 - x_2*y_1+ x_2*y_2 - z_2 + x_2*y_2 -  x_2*y_2$$
+$$= t* (z_1 - x_1*y_1) + x_2*y_2 - z_2 $$
+Therefore, \( \phi = 0 \Leftrightarrow x_1 * y_1 = z_1 \land x_2*y_2=z_2 \), so this only holds if both triples are correct. \\
+In this process the triple \( (\langle x_1 \rangle ,\langle y_1\rangle , \langle z_1\rangle ) \) wasn't revealed, so now it can be used for the multiplication knowing that one the one hand it is still secretly shared and unknown and on the other hand that it was not modified by an adversary.\\
+
+\textbf{Sacrificing of triples in SCALE-MAMBA}
+
+In later versions of SPDZ, since ~\cite{cryptoeprint:2017:123}, and in the current implementations of SCALE-MAMBA, a slightly modified version of aforementioned protocol is used.
+The idea is to not sacrifice a random second triple, to guarantee the correctness of the first one, but to sacrifice a similar triple.
+For this, the triples are directly generated in pairs, so that there are are two triples \( (\langle x \rangle ,\langle y \rangle , \langle z \rangle ) \) and \( (\langle x' \rangle ,\langle y\rangle , \langle z'\rangle ) \) with $z=x*y$ and $z'=x'*y$.
+With $y_1=y_2$, the formula from SPDZ becomes directly much simpler.
+\begin{enumerate}
+\item Choose a random variable \( r\) and open it to the other parties. This is done by each party generating a random share \(r_i\) and then calculating \(r=r_1+...+r_n\).
+\item The parties calculate  \( p=r*\langle x \rangle - \langle x' \rangle \) and partially open the result afterwards
+\item The parties calculate \(\phi = r * \langle z \rangle\ - \langle z' \rangle - p * \langle y \rangle\) and then partially open the result
+\item If \( \phi \) is 0 then the triples are correct and it is proceeded with \( (\langle x \rangle ,\langle y\rangle , \langle z\rangle )\) as a correct, secretly shared and not revealed triple.
+Otherwise the system detected a malicious adversary and it aborts.\\
+\end{enumerate}
+
+\textbf{Correctness}: Parallel to before, simplifying \( \phi \) leads to the same implications:\\
+$$\phi =r*z-z'-p*y=r*z-z'-(r*x-x')*y=r*(z-x*y)+(x'*y-z')$$
+Therefore, \( \phi = 0 \Leftrightarrow x * y = z \land x'*y=z' \)
+
+\textbf{Comparison of Sacrificing}
+This small improvement in the protocol nearly halved the mathematical operations necessary for the scarification process.
+This is huge, because the triple generation is one of the most time consuming task, so it sped up the offline phase.
+In following table, the differences between both sacrificing protocols can be seen:
+
+\begin{center}
+\begin{tabular}{ | c | c | c | }
+\hline
+& SPDZ & SCALE-MAMBA \\
+\hline
+ Additions & 6 & 3 \\ 
+\hline
+ Multiplications with constants & 5 & 3 \\  
+\hline
+Partial reveals & 3 &2 \\
+\hline
+\end{tabular}
+\end{center}
+
+In SPDZ: Sacrifice in online phase???
 
@@
 \subsection{SPDZ 2.0 in medical healthcare}
-In medical healthcare, preserving privacy plays a huge role because patient records hold highly sensitive data, thus, its a natural area for the application of privacy protecting techonologies.
+In medical healthcare, preserving privacy plays a huge role because patient records hold highly sensitive data, thus, its a natural area for the application of privacy protecting technologies.
+In [TODO: REFERENCE] a clinical decision support system was implemented with SPDZ 2.0.
+The idea is that the system provides the optimal treatment for a patient with HIV based on the  records of similar patients in different hospitals.
+To preserve privacy, the hospitals compute the efficiency of one treatment for one patient together with SMPC.\\
+The researchers used a setup of 20.000 patient records split between the hospitals, and were able to compute the effectiveness of 100 treatments for one patient in 24 minutes in the online phase.
+For calculating the effectiveness of one treatment, their algorithm needed 40 million multiplications.
+They only implemented the online phase of SPDZ, but showed that it scaled linearly with the amount of patient records.
+
+TODO: CHECK HOW MANY HOSPITALS
+
+Supposedly, the offline phase took them around 22 minutes.
 
@@
 https://www.researchgate.net/publication/324077995\_SEMBASEcure\_multi-biometric\_authentication
+
+https://www.moodle.ch/lms/pluginfile.php/6253/mod\_book/chapter/25/mpc-book.pdf
 

### 2019-04-19 paper.tex
Freitag abend

@@
 \begin{abstract}
-This paper takes a look at Secure Multi-Party Computation(SMPC), especially using SCALE-MAMBA.
-For that, it introduces first what SMPC is, how SPDZ, the predecessor of SCALE-MAMBA, works and what are the differences.
-It discusses the functionalities of SCALE-MAMBA, especially the split into online and offline phase, how the beaver tripels are used for sacrificing and in which complexity classes the calculations are done.
-Finally it compares scale mamba shortly to other competitors in the market and gives an outlook what next technologies are coming right now and are coming in the near future in SMPC.
+This paper examines Secure Multi-Party Computation(SMPC), especially the SCALE-MAMBA protocol.
+To do that, it first introduces SMPC, afterwards SPDZ, the predecessor of SCALE-MAMBA, and, after taking a detailed look at SCALE-MAMBA, compares them.
+It focuses on the algorithmic foundations of SCALE-MAMBA, especially the split into online and offline phase, how beaver triples are used to enable computational efficient multiplications and which computational complexities the algorithms posses.
+Afterwards, it tries to depict the evolution in the different versions of the protocol and to locate SCALE-MAMBA between its competitors.
+Ultimately, it will show some exemplary implementations in real world applications and will give an outlook what further evolutions might be coming in the next years.
 \end{abstract}
@@
 \section{Motivation and Introduction}\label{sec:intromoti}
-
-In todays world, and even more in the future, there are a lot of computations that have to be done in a distributed environment, because in a lot of use cases multiple different parties are involved, but also because the amount of data is getting so big that it often becomes impossible to have a single entity that processes and stores all the data.
-Furthermore, in a lot of these cases the parties participating in these computations do not want to reveal their own information, because it is sensitive information or because they do not want to share information with a lot of business value.\\
+In today's world, and even more in the future, there exist a lot of computations that have to be done in a distributed environment, because in a lot of use cases multiple parties are involved, but also because the amount of data is getting so big that it often becomes impossible to have a single entity that processes and stores all of it.
+Furthermore, in a lot of these cases the parties participating in these computations do not want to reveal their own information, because it is sensitive or it has a high business value.\\
 A basic example is an anonymous voting.
-In most elections there is a third party that has to aggregate and count the votes made by the voters.
-For instance, in Germany's elections this is done by specific people counting a small amount of votes and then forwarding the information to a central institution.
+In most elections there is a third party that has to aggregate and count the votes.
+For instance, in Germany's elections[TODO: LINK] this is done by specific people counting a small amount of votes and then forwarding the information to a central institution.
 The other problem is that it has to be guaranteed that a vote is valid, meaning that each voter holds the right to vote and is only voting one time.
-So if there would be a malicious person in this specific system, it would be easy for that person to gather information over a small subset of voters or even manipulate the local results.\\
+So in an ideal system, no person participating in the aggregation process should be able to get any information over any vote or any subset of voters.\\
 Another use case is the exchange of medical data.
 Medical data of patients is highly confidential and should never be published.
-But at the same time, a lot of different hospitals or doctors have only the data from their small data set of local patients, and for making founded diagnosis for new patients it is helpful to compare the symptoms to the diagnoses of other doctors.\\
+But at the same time, a lot of different hospitals or doctors have only the data from their small data set of local patients, and for making founded diagnosis for new patients it would be helpful to compare the symptoms to the diagnoses of other doctors.\\
 Secure Multi-Party Computation(SMPC) is one possible solution for these use cases. It is a field of research that emerged in the last 30 years, and is gathering a lot of speed in the last years. 
-The basic problem is how to evaluate a common function, with multiple parties having different inputs to this function, without revealing the different inputs of the parties.
+The basic problem is the evaluation of a common function, to which multiple parties have different inputs, without revealing these inputs.
 Although there are a lot of theoretical solutions for this problem, real-world implementations are still not performing on the desired level.
-So in the last years the research shifted more into the area of how to make the existent solutions performant enough while still guaranteeing a desired level of security.\\
-This paper's main focus will lay on SCALE-MAMBA, a system developed mainly by the Katholieke Universiteit Leuven, which is one of the forerunners in SMPC and a system that combines a lot of research by different teams with the newest ongoing research to provide a platform with highly performant SMPC.
+So in the last years the research shifted more into the area of how to improve the performance of existing solutions while still guaranteeing a desired level of security.\\
+This paper's main focus will lay on SCALE-MAMBA, a system developed mainly by the Katholieke Universiteit Leuven.
+This system is one of the forerunners in SMPC and combines a lot of research by different teams to provide a platform with highly performant SMPC.
 Nevertheless, this paper will at first give a short introduction into the history of SMPC and will try to explain its basic technologies to the reader.
-It will compare SCALE-MAMBA to its predecessors and will also try to give a short comparison to other SMPC systems which use different approaches, especially considering the usability in different use cases.
+It will compare SCALE-MAMBA to its predecessors and will also try to give a short comparison to other SMPC systems which use different approaches, especially taking a look on which different requirements are fulfilled by each.
 Ultimately, it will try to give an outlook what might be coming in the next years.\\
@@
 
-GENERATION OF BEAVER TRIPLES
+GENERATION OF BEAVER TRIPLES: Include protocol on how triples are distributively generated without revealing them
 
@@
 One of the currently ongoing improvements on SCALE-MAMBA tackles exactly this problem.
-The developer team tries to give MAMBA the possibility to compile code just in time.
-That would offere the possibility, that the offline data is calculated before and then, when the need arises, a computatation could be specified in MAMBA and then intime compiled and executed, without having to wait for the expensive offline calculations.\\
+The developer team tries to give MAMBA the possibility to compile code just in time as mentioned in [TODO: link to documentation].
+That would offer the possibility, that the offline data is calculated before and then, when the need arises, a computation could be specified in MAMBA and then in-time compiled and executed, without having to wait for the expensive offline calculations.\\
 
 \section{Implementations}
-Because SCALE-MAMBA was only released in 2018, there are not a lot documented usecases in praxis.
-Nevertheless, because of the structural and alogirthmic similarities to the earlier versions of SPDZ, taking a look on SPDZ implementations should give a general understanding of the possible usage of SCALE-MAMBA.\\
+Because SCALE-MAMBA was only released in 2018, there are not a lot documented use cases in praxis.
+Nevertheless, because of the structural and algorithmic similarities to the earlier versions of SPDZ, taking a look on SPDZ implementations should give a general understanding of the possible usage of SCALE-MAMBA.\\
 
@@
 They only implemented the online phase of SPDZ, but showed that it scaled linearly with the amount of patient records.
+In the paper the amount of parties participating in the computation was not specified, but because the computation is quite good, it can be assumed that a really simple setup is used that doesn't create a lot of overhead.
 
+TODO: compare, what is the speed and how fast are different ones? 
 TODO: CHECK HOW MANY HOSPITALS

### 2019-04-19 paper.tex
Gute nacht

@@
 Message Authentification Codes (MACs) are used to guarantee the integrity and authenticity of a message.
-Basically, a MAC is a tag that is generated for a message and that, if it is correct, verifies the identity of the sender and that the content of the message was not changed.\\
+Basically, a MAC is a key that is used to generate a tag for a message.
+This tag is used to verify the identity of the sender and that the content of the message was not changed.\\
 For using MACs we need a tupel of two algorithms \((\text{MAC},\text{VER})\):\\
@@
 
-These algorithm has the disadvantage of high storage complexity, but therefore it doesn't take a long time to calculate and it is possible to get the original message with just having the key and the tag.
-This is a property that is used for an advantage later on in the protocol.\\
+This algorithm has the disadvantage of high storage complexity, but therefore it doesn't take a long time to calculate and it is symmetric so the same key can be used for creating the tag and verifying it.
+The symmetry is used for an advantage later on in the protocol.\\
 
@@
 As we saw in the introduction, there are a lot current and much more upcoming use cases of calculating the output of a function in a distributed environment with multiple parties while at the same time keeping all the individual inputs secret.
-The first explorations in these fields started in the 1970s, with theoretical problems like the mental poker problem(how can two players play poker while not being in the same physical location and making sure no one is cheating) and Yao's millionaire problem(how can can two millionaires decide who is richer than the other while not revealing their wealth).
+The first explorations in these fields started in the 1970s, with theoretical problems like the mental poker problem(how can two players play poker while not being in the same physical location and making sure no one is cheating) and Yao's millionaire problem(how can can two millionaires decide who is richer than the other while not revealing their respective wealth).
 \\
@@
 \item Given a function \(f(x_1,...,x_m)\)
-\item Goal: Evaluating the function \(f(x_1,...,x_m)\) while guaranteeing secrecy and correctness
+\item Goal: Evaluation of the function \(f(x_1,...,x_m)\) while guaranteeing secrecy and correctness
 \end{itemize}
@@
 One is based on garbled circuits which were first introduced by Yao in 1986 \cite{4568207}.
-The basic idea is that, in a two party setting, one party, Alice, garbles the boolean circuit that has to be evaluated and sends it to the second party, Bob.
+The basic idea is that, in a two party setting, one party, Alice, garbles the Boolean circuit that has to be evaluated and sends it to the second party, Bob.
 Bob uses his own input, Alice's encrypted input and the garbled circuit to evaluate the circuit in an encrypted way.
@@
 There is a huge family of protocols build on this approach, and it is still evolving every day.
-This paper will focus on the newest of these protocols that were developed in the last 10 years, especially on the protocols which are based on or similiar to SPDZ.\\
+This paper will focus on the newest of these protocols that were developed in the last 10 years, especially on the protocols which are based on or similar to SPDZ.\\
 

### 2019-04-21 paper.tex
Sun Morn

@@
 Ultimately, it will try to give an outlook what might be coming in the next years.\\
-
-\section{Cryptographic Basics}\label{sec:basics}
-
-Before delving deeper into SMPC, this section will shortly recap on cryptography basics that might be unknown to the reader.
-These basics are not solely used by SMPC, but nevertheless, play a huge role in a lot of protocols.
-
-\subsection{Message Authentification Codes}
-Message Authentification Codes (MACs) are used to guarantee the integrity and authenticity of a message.
-Basically, a MAC is a key that is used to generate a tag for a message.
-This tag is used to verify the identity of the sender and that the content of the message was not changed.\\
-For using MACs we need a tupel of two algorithms \((\text{MAC},\text{VER})\):\\
-\begin{itemize}
-\item The MAC algorithm \(\text{MAC}(K,m)\rightarrow t\) takes a key \(K\) and a message \(m\) and outputs a tag \(t\)\\
-\item The verficiation algorithm \(\text{VER}(K,m,t)\rightarrow 0/1\) takes a key \(K\), a message \(m\) and a tag \(t\) and outputs accept(\(1\)) or reject(\(0\))\\
-\end{itemize}
-The correctness can be easily seen: \(\forall K: \text{VER}(K,m,\text{MAC}(K,m))=1\)\\
-
-In SCALE-MAMBA really simple MACs are used. We assume that we have a MAC key \( \alpha \). Then the MAC algorithm used is following:\\
-$$\text{MAC}(\alpha,m)\rightarrow \alpha * m$$ 
-$$\text{VER}(\alpha,m,t)\rightarrow (1 \Leftrightarrow t = \alpha*m) $$
-
-This algorithm has the disadvantage of high storage complexity, but therefore it doesn't take a long time to calculate and it is symmetric so the same key can be used for creating the tag and verifying it.
-The symmetry is used for an advantage later on in the protocol.\\
-
-\subsection{Homomorphic Encryption}
-Encryption as a basic concept is simple. Similiar to MACs, a tupel (E,D) of algorithms is needed.\\
-\begin{itemize}
-\item The enncryption algorithm \(\text{E}(K,m)\rightarrow c\) takes a key \(K\) and a message \(m\) and outputs a cyphertext \(t\) that is the encryption of the message \(m\)\\
-\item The decryption algorithm \(\text{D}(K,c)\rightarrow m\) takes a key \(K\) and a cyphertext \(m\) and outputs the decrypted message \(m\), assuming c was not modified\\
-\end{itemize}
-Alternatively, in a public key setting, the key for encryption is public and there is a different key for decryption that is privat.\\
-The problem is, that modifying the cypertext in any way makes the decryption useless.
-In some cases it is desirable do do computations on cypertexts, because then it is possible to make calculations without revealing the unencrypted data.
-That is why Homomorphic Encryption was developed.
-Let's assume that we have a ring \((R,+,*)\) as our message space.
-Fully Homomorphic Encryption (FHE) is preserving the ring structure.
-Therefore, it is possible to do additions and multiplications of cypertexts and when decrypting the result it is the same value as when doing the same additions and multiplications on the real values.\\
-Althought FHE has amazing properties for a lot of usecases, in praxis it is way too slow.
-Because of all the overhead generated by the encryption scheme to guarantee the preservation of the ring structure, the current algorithms are not feasible for most applications.
-
-\subsection{Somewhat Homomorphic Encryption}
-That is where Somewhat Homomorphic Encryption (SHE) becomes relevant.
-SHE only supports one operation, and in most cases that is addition.
-This takes away a lot of possibilities but, therefore, the performance gain is immense.
-So immense, that SHE is used a lot in practice, for example in SCALE-MAMBA.\\
-The SHE scheme used in SCALE-MAMBA, is based on the so called Brakerski-Gentry-Vaikuntanathan cryptosystem(BGV).\\
-
 \section{SMPC Basics and History}\label{sec:smpc}
-
 \subsection{Beginnings}
@@
 These explorations lead to a group of basic protocols which established the area of Secure Multi-Party Computation(SMPC).\\
-
 \subsection{Problem statement}
@@
 Now we will take a short look on two of the first protocols, which both are the basis for a whole family of SMPC schemes.
-
 \subsection{Yao's Garbled Circuit}
@@
 At the end they exchange the results and can compute the result this way.
-
 \subsection{Shamirs Secret Sharing Scheme}
@@
 These type of protocols were introduced by Adi Shamir at a similar time, in 1979 \cite{Shamir:1979:SS:359168.359176}.\\
-The basic idea of this approach is that multiple parties split their secrets into so called shares and then distribute them to the other parties.
-In Shamir's scheme this is done by generating random polynomials with a degree depending on the security level and then calculating the shares as points in this polynomial.
+The basic idea of this approach is the the so called share-compute-reveal paradigm, which splits the evaluation of the function into three parts.\\
+\textbf{Share:} First, the inputs secrets are split into multiple parts which then are shared shared with the other parties. Hence, every party holds a share of each input secret.\\
+\textbf{Compute:} The parties evaluate the function with the secret input shares.
+The protocol has to be defined in a way, that the final result which each parties computes with the input shares is a respective share of the complete final resut.\\
+\textbf{Reveal:} The parties agregate their shares of the final result and compute the final result together.\\
+This paradigm already shows a lot of challenges arising for a concrete protocol, for instance:
+\begin{itemize}
+\item How can the secrets be shared in a homomorphic way, so that, after computing on shared input, the complete result can be recovered from the result shares?
+\item While this can be done for addition in easy ways, how can this be done for multiplication as well?
+\item How can the correctness of the computation be verified?
+\item How can the overhead generated by a protocol can be kept in a reasonable size, so that a protocol is applicable to a real-world use case?
+\end{itemize}
+In Shamir's scheme the secret sharing is implemented by expressing a secret $s$ and its shares as a random polynomial $f(x)$, so that $f(0)=s$ and each share $s_i$ has the value $s_i=f(i)$.
 Computations are executed on these shares, and at the end it is possible to reconstruct the result from the different shares without revealing information about the initial secrets.
-This is done by using Lagrange interpolation.\\
+This is done by using Lagrange interpolation.
+.
+The degree $p$ of the polynomial can be seen as a security parameter, because $p$ shares are needed to reconstruct the secret.\\
 There is a huge family of protocols build on this approach, and it is still evolving every day.
-This paper will focus on the newest of these protocols that were developed in the last 10 years, especially on the protocols which are based on or similar to SPDZ.\\
-
+This paper will focus on the newest of these protocols that were developed in the last 10 years, especially on the protocols which are based on or similar to SPDZ.
+\\
 Before looking into exact implementations of linear secret sharing schemes, following, a short look will taken at the general adversary model and required basic definitions.
-
 \subsection{Adversary Model}
@@
 In the area of SMPC there is normally a differentiation between two types of adversaries.
-
 \subsubsection{Honest but curious}
@@
 This adversary tries to retrieve some of the secret information from the other parties, by using all the information he gets during the execution of the protocol.
-
-\subsection{Malicious adversary}
+\subsubsectionx{Malicious adversary}
 While this adversary also wants to gain all the information, he is not obeying to the rules of the protocol.
@@
 That could mean making the computation impossible, but as well modifying his own input or his own calculations, so the final result is changed and might lead to different implications
-
 It is important to decide between these two types of adversary, because it shows the different securities SMPC has to guarantee.
@@
 only protecting versus non malicious adversaries.
-
-\subsubsection{Access structures}
+\subsection{Access structures}
 There are a lot of different use cases for SMPC, so of course the adversary model changes depending on the system, the security needs, and the amount of parties involved.
@@
 
-\subsubsection{Overview}
+\subsection{Overview}
 
@@
 
-\subsubsection{MAMBA}
+\subsection{MAMBA}
 For this paper the more interesting part is the algorithmic part.
@@
 
-\subsubsection{Secret sharing}
+\subsection{Secret sharing}
 
@@
 
-\subsubsection{Revealing secrets}
+\subsection{Revealing secrets}
 
@@
 
-\subsubsection{Offline phase}
+\subsection{Offline phase}
 
 There are two main things done in the offline phase.
-The first one is the generation of a global MAC key $\alpha$.
-This MAC key is later used, to authenticate the different shares at the end of the computation.
-This key is also distributively generated, i.e.
-only if all parties pool together their shares they can recover its value.\\
+The first one is the generation of a global Message Authentification Code(MAC) key $\alpha$.
+This MAC key is later used, to authenticate the different shares at the end of the computation.\\
+The second and and most important part in the offline phase, is the generation of the beaver triples.
+These triples  are independent of the computation that is later performed in the online phase, but are used in the online phase to express multiplications as a combination of additions and multiplications with constants.
+For each multiplication, one of the triples is needed.
 
-The second and and most important part in the offline phase, is the generation of the beaver triples.
-These triples  are independent of the computation that is later performed in the online phase, but will be used in the online phase to guarantee that multiplications can be done without revealing the secret inputs of the different parties.
-For each multiplication, one of the triples is needed.
+\subsubsection{Generation of the MAC key}
+Message Authentification Codes (MACs) are used to guarantee the integrity and authenticity of a message.
+Basically, a MAC is a key that is used to generate a tag for a message.
+This tag is used to verify the identity of the sender and that the content of the message was not changed.\\
+For using MACs we need a tupel of two algorithms \((\text{MAC},\text{VER})\):\\
+\begin{itemize}
+\item The MAC algorithm \(\text{MAC}(K,m)\rightarrow t\) takes a key \(K\) and a message \(m\) and outputs a tag \(t\)\\
+\item The verficiation algorithm \(\text{VER}(K,m,t)\rightarrow 0/1\) takes a key \(K\), a message \(m\) and a tag \(t\) and outputs accept(\(1\)) if the tag is correct or reject(\(0\))\\ if it is not
+\end{itemize}
+
+In SCALE-MAMBA really simple MACs are used. We assume that we have a MAC key \( \alpha \). Then the MAC algorithm used is following:\\
+$$\text{MAC}(\alpha,m)\rightarrow \alpha * m$$ 
+$$\text{VER}(\alpha,m,t)\rightarrow (1 \Leftrightarrow t = \alpha*m) $$
+
+This algorithm has the disadvantage of high storage complexity, but therefore it doesn't take a long time to calculate and it is symmetric so the same key can be used for creating the tag and verifying it.
+The symmetry is used for an advantage later on in the protocol.\\
+
+The MAC key is distributively generated, i.e. only if all parties pool together their shares they can recover its value. 
+[TODO Expand on MAC generation
 
@@
 
-\subsubsection{Online Phase}
+\subsection{Online Phase}
 
@@
 
-\subsubsection{Upcoming changes to SCALE-MAMBA}
+\subsection{Upcoming changes to SCALE-MAMBA}
 SCALE-MAMBA is a research system, so there will be a lot more research time spend on improving the system further.
@@
 
-\subsubsection{Outlook}
+\subsection{Outlook}
 Their are a lot of future applications for Secure Multiparty Computation.
@@
 
-\subsubsection{Conclusion}
+\subsection{Conclusion}
 

### 2019-04-21 paper.tex
Progress :)

@@
 This adversary tries to retrieve some of the secret information from the other parties, by using all the information he gets during the execution of the protocol.
-\subsubsectionx{Malicious adversary}
+\subsubsection{Malicious adversary}
 While this adversary also wants to gain all the information, he is not obeying to the rules of the protocol.
@@
 An \((n,t)\)-threshold scheme with \(t<\frac{n}{l}\) means that at least \(t+1\) parties are needed to form a qualified set.
-Here, \(n\) is the number of total parties, and \(t\) is number of possible adversaries while not revealing any secrets.\\
+Here, \(n\) is the number of total parties, and \(t\) is number of possible adversaries while not revealing any secrets ~\cite{MPCBook_threshold}.\\
 A full threshold scheme means that all parties are needed to form a qualified set.
@@
 In medical healthcare, preserving privacy plays a huge role because patient records hold highly sensitive data, thus, its a natural area for the application of privacy protecting technologies.
-In [TODO: REFERENCE] a clinical decision support system was implemented with SPDZ 2.0.
+In ~\cite{PPCDSS} a clinical decision support system was implemented with SPDZ 2.0.
 The idea is that the system provides the optimal treatment for a patient with HIV based on the  records of similar patients in different hospitals.
@@
 
-https://www.researchgate.net/publication/328040632A\_New\_Approach\_to\_Privacy-Preserving\_Clinical\_Decision\_Support\_Systems\_for\_HIV\_Treatment
-https://arxiv.org/pdf/1901.00329.pdf
-https://www.researchgate.net/publication/324077995\_SEMBASEcure\_multi-biometric\_authentication
+\subsection{Other implementations}
+
+In ~\cite{SEMBA} an implementation of biometric-based authentication, i.e.
+iris and face authentication, with SPDZ was shown.
+The researchers were able to improve existing solutions with the use of SPDZ.\\
+
+~\cite{2019arXiv190100329C} implemented typical machine learning algorithms like linear and logistic regression with SPDZ and showed that it outperforms a competing protocols Obliv-C ~\cite{cryptoeprint:2015:1153} and matches SecureML ~\cite{SecureML}.
+So it can be assumed, that an implementation with SCALE-MAMBA would show even better results.\\
 
@@
 https://scapi.readthedocs.io/en/latest/intro.html
+
+While SPDZ might be the best multi party protocol, in X it was compared to tw
+
+Two party protocols
+~\cite{MPCCompression}
+In X it was shown, that SPDZ 2.- has vulnerabilities.
+It was compared to the three most cited two party protocols, and it was shown that for a specific attack, SPDZ is the most vulnerable of these.
+It was as well shown, that the secret sharing is predictable.
+The researchers didnt analyze, if this is still the case in SCALE-MAMBA.
+TinyLego ~\cite{TinyLego}
+Duplo ~\cite{Duplo}
+~\cite{Aby}
+
+Nevertheless, in a two party settings there are more secure protocols.
 

### 2019-04-21 paper.tex
Not a lot left , i just wanna finish

--- 2019-04-21 paper.tex
@@
 
-\subsection{Upcoming changes to SCALE-MAMBA}
-SCALE-MAMBA is a research system, so there will be a lot more research time spend on improving the system further.
+\subsection{Problems and upcoming changes}
+\subsubsection{Upcoming Changes}
+SCALE-MAMBA is a research system, so there will be a lot more research spend on improving the system further.
 An important aspect is the usability in the practice.
@@
 One of the currently ongoing improvements on SCALE-MAMBA tackles exactly this problem.
-The developer team tries to give MAMBA the possibility to compile code just in time as mentioned in [TODO: link to documentation].
-That would offer the possibility, that the offline data is calculated before and then, when the need arises, a computation could be specified in MAMBA and then in-time compiled and executed, without having to wait for the expensive offline calculations.\\
+The developer team tries to give MAMBA the possibility to compile code just in time as mentioned in ~\cite{ScaleMambaDocu}.
+That would offer the possibility, that the offline data is calculated before and then, when the need arises, a computation could be specified in MAMBA and in-time compiled and executed, without having to wait for the expensive offline calculations.\\
+
+\subsubsection{Problems}
+It was shown recently in ~\cite{MPCCompression} that there are security flaws in SPDZ 2.0 and that its secret sharing is predictable.
+It doesn't seem to be the case that these imperfections are already fixed in SCALE-MAMBA.\\
+Additionally, a single cheater can break the whole system because it is abort-secure.
+in ~\cite{cheaterSPDZ} an extended protocol is suggested to detect and expel cheaters if one is detected and afterwards proceed with the protocol.
+This would be an important feature for a real-world use case, so the protocol doesn't get stuck in a loop of abortion.
 
@@
 TODO: compare, what is the speed and how fast are different ones?
-TODO: CHECK HOW MANY HOSPITALS
 
@@
 
-https://www.researchgate.net/publication/328040632A\_New\_Approach\_to\_Privacy-Preserving\_Clinical\_Decision\_Support\_Systems\_for\_HIV\_Treatment
-https://arxiv.org/pdf/1901.00329.pdf
-https://www.researchgate.net/publication/324077995\_SEMBASEcure\_multi-biometric\_authentication
-
-https://www.moodle.ch/lms/pluginfile.php/6253/mod\_book/chapter/25/mpc-book.pdf
+\subsection{Other implementations}
+In ~\cite{SEMBA} an implementation of biometric-based authentication, i.e.
+iris and face authentication, with SPDZ was shown.
+The researchers were able to improve existing solutions with the use of SPDZ.\\
+~\cite{2019arXiv190100329C} implemented typical machine learning algorithms like linear and logistic regression with SPDZ and showed that it outperforms a competing protocol Obliv-C ~\cite{cryptoeprint:2015:1153} and matches SecureML ~\cite{SecureML}.
+So it can be assumed, that an implementation with SCALE-MAMBA would show even better results.\\
 
 \section{Comparison}\label{sec:comparison}
-
-a.	What else is out there?
-b.	Different use cases – when which SMPC tool is more useful; what are the strengths of SCALE-MAMBA compared to other ones?
-	- Security! High performant for such a secure system
-	- In some cases not the highest security is needed; application for a normal user, not for a company, have much less security requirements but much higher requirements on performance
-	- Active security is not always required, in non-critical systems it is often sufficient to know that noone knows the users data
-c.	Factual comparisons
-	- YAO/LSSS --> LSSS protocols right now are more suited for the praxis
-
-	- SCAPI
-	-SCAPI is an open-source Java library for implementing secure two-party and multiparty computation protocols (SCAPI stands for the “Secure Computation API”).
-It provides a reliable, 	efficient, and highly flexible cryptographic infrastructure.
-SCAPI provides only passive security
-
-https://scapi.readthedocs.io/en/latest/intro.html
+In the sector of multi party protocols with highest possible security, i.e.
+active and full-threshold security, SCALE-MAMBA, and before SPDZ, seem to be the universally accepted state of the art protocols.
+As already seen before, SCALE-MAMBA has several improvements in run-time and security compared to its predecessors.\\
+Nevertheless, the requirements are not always the same, so depending on the use-case the choice of protocol might be depend.
+No complete comparison to other protocols will be, but this section should transmit the parameters which should be considered when making this choice.\\
+\subsubsection{Two party protocols} There are a lot of protocols, which are restricted to a two party setup.
+As shown in ~\cite{MPCCompression}, frequently used protocols like TinyLEGO ~\cite{TinyLego} and  DUPLO ~\cite{Duplo}, which are both protocols based on garbled circuits, and ABY ~\cite{Aby}, which combines secret sharing and garbled circuits, all outperform SPDZ 2.0 in terms of information loss regarding compression and entropy metrics.\\
+\subsubsection{Passive Security}
+There are protocols, for instance SCAPI ~\cite{SCAPI}, which only implement passive security, i.e.
+they are not secure versus malicious adversaries.
+In a setup with looser security requirement a different protocol might be better suited because a lot of verification can be omitted.\\
+\subsubsection{Different Access structures}
+Full-threshold security is not always a desired property.
+While SCALE-MAMBA can be configured for different access structure, different protocols optimized for these access structures might outperform it.\\
 
 \section{Resume}\label{sec:outlook}
-
 \subsection{Outlook}
-Their are a lot of future applications for Secure Multiparty Computation.
-With the threoretical research of the early years much more turning to praxis oriented research, todays SMPC protocols, and especially the successors of SPDZ, have a bright future.
-With already more and more applications starting to use SMPC in systems that need to guarantee security of private data, the list will only grow more and more with the performance of the protocols and the same time the available computation power improving.
-\\
-With SCALE-MAMBA being one of the first performant systems that can be really used in a real time scenario, it will probably not take a lot of more years until there will be proper commercial solutions.
-And if big companies start to adapt and use these theoretical foundations, then, in my opinion, SMPC will be a really important tool in computer science and especially in distrbuted environments.
-There are a lot more possible applications that could be coming in the next years, for example in the financial branche SMPC could play a huge role.\\
-SMPC can be a technology that shapes the future, or it can be just a stepstone for a new technology to emerge.
-But definetly, privacy and security plays already a huge role in our society, and this will not diminish in any close future.
-
-https://www.researchgate.net/publication/309817066\_Cheater\_Detection\_in\_SPDZ\_Multiparty\_Computation
-https://www.researchgate.net/publication/330091440\_Breaking\_MPC\_implementations\_through\_compression
-
+There are a lot of future applications for Secure Multiparty Computation.
+With the theoretical research of the early years turning to more praxis-oriented research, today's SMPC protocols have a bright future.
+With already more and more applications starting to use SMPC in systems that need to guarantee security of private data, and with the protocols further improving in performance and security, the list will only continue growing.\\
+SCALE-MAMBA is one of the first systems that can is functional and performant enough to be used in real-world scenarios, and, thus, it will probably not take a lot more years until there will be proper commercial solutions.
+And if big companies start to adapt and use these theoretical foundations, then, in my opinion, SMPC will be a really important tool in computer science and especially in distributed environments.
+There are a lot more possible applications that could be coming in the next years, for example in the financial sector or in inter-governmental collaborations.\\
+SMPC can be a technology that shapes the future, or it can be a stepping stone for new technologies to emerge.
+But definitely, privacy and security already play a huge role in our society, and this will not diminish in any close future.
 \subsection{Conclusion}
-
 In this paper we could see how SMPC works in general, and, especially, how SCALE-MAMBA works and how it evolved over the years in comparison to its predecessor, SPDZ.
 It was shown, how the protocol is split into a primary offline phase in which the computational expensive calculations are executed and all the preparations are done for the later circuit evaluation, like calculating the Beaver Triples, and how afterwards in the online phase the circuit is evaluated, how the triples are used to execute a multiplication without sharing the secret inputs and how the results can be verified at the end.
+Nevertheless, SCALE-MAMBA is optimized for a setup with active and full threshold security, so if the use case varies, there might be a more suitable SMPC protocol.
 

### 2019-04-21 paper.tex
nearly finished

@@
 But at the same time, a lot of different hospitals or doctors have only the data from their small data set of local patients, and for making founded diagnosis for new patients it would be helpful to compare the symptoms to the diagnoses of other doctors.\\
-Secure Multi-Party Computation(SMPC) is one possible solution for these use cases. It is a field of research that emerged in the last 30 years, and is gathering a lot of speed in the last years. 
+Secure Multi-Party Computation(SMPC) is one possible solution for these use cases. It is a field of research that emerged within the last 30 years, and is gathering a lot of speed in the last years. 
 The basic problem is the evaluation of a common function, to which multiple parties have different inputs, without revealing these inputs.
@@
 \subsubsection{Honest but curious}
-The honest but curious adversary tries to gain as much information as possible while obeying to the rules of the protocol.
+The honest but curious adversary tries to gain as much information as possible while obeying the rules of the protocol.
 Hence, it is impossible to know that this party is controlled by a third entity.
@@
 \subsubsection{Malicious adversary}
-While this adversary also wants to gain all the information, he is not obeying to the rules of the protocol.
+While this adversary also wants to gain all the information, he is not obeying the rules of the protocol.
 This type of adversary can send different inputs to manipulate the results he receives.
@@
 That could mean making the computation impossible, but as well modifying his own input or his own calculations, so the final result is changed and might lead to different implications
-It is important to decide between these two types of adversary, because it shows the different securities SMPC has to guarantee.
+It is important to distinguish between these two types of adversary, because it shows the different securities SMPC has to guarantee.
 It has to guarantee that no information is unwillingly shared with an illegal party and at the same time verify that the final result has a correct value.
@@
 An \((n,t)\)-threshold scheme with \(t<\frac{n}{l}\) means that at least \(t+1\) parties are needed to form a qualified set.
-Here, \(n\) is the number of total parties, and \(t\) is number of possible adversaries while not revealing any secrets ~\cite{MPCBook_threshold}.\\
+Here, \(n\) is the number of total parties, and \(t\) is the number of possible adversaries while not revealing any secrets ~\cite{MPCBook_threshold}.\\
 A full threshold scheme means that all parties are needed to form a qualified set.
-This implies, that, if all other parties except one are corrupted by an adversary, this one party can still rely on the fact that it secret remains secret.
+This implies, that, if all other parties except one are corrupted by an adversary, this one party can still rely on the fact that its secret remains secret.
 Of course this has a huge impact on overhead and performance, but in some cases it can provide needed security guarantees.
@@
 In some cases there might be different type of parties, where different parties have different positions or levels of trust, so it is, for example, possible to specify that some set of parties only has to have the size of 2 to form a qualified set and another one needs the size of 4.\\
-Choosing the right access structure for the right use case is really important, because it has a realy big impact on performance and on the desired security level.
+Choosing the right access structure for the right use case is important, because it has a big impact on performance and on the desired security level.
 
@@
 
-In the last years, research in SMPC is constantly ongoing, and there are new approaches or improvements to old approaches appearing regulary.
+In the last years, research in SMPC is constantly ongoing, and there are new approaches and improvements to old approaches appearing regularly.
 But most of these publications only focus on a small area of the whole picture.
@@
 That also means for this paper, that there is not a single protocol that can be examined.
-Thus, this paper tries to give more of an overview how the process works in general, which are the main, especially which are newest and best, components and where are the problems and bottlenecks.\\
+Thus, this paper tries to give more of an overview how the process works in general, which are the main, especially newest and best performing, components and where are the problems and bottlenecks.\\
 SCALE-MAMBA consists of two parts.
@@
 It contains all the computational logic and algorithms for multi-party computations.
-The second part is called Multiparty AlgorithMs Basic Argot (MAMBA) and is a compiler for the SMPC programs.\\
+The second part is called Multiparty AlgorithMs Basic Argot (MAMBA) and is a compiler for the SMPC programs.
 
@@
 In the online phase only the final evaluation of the function is executed.
-In the offline phase, prior to the online phase, preparations are done by all the parties to prepare for the evaluation.
-This includes already calculating values for possible multiplications and MACs, so that the online phase can be executed faster.
-In this phase the expensive calculations are done, like public key cryptography or authentifications, so that the online phase only needs basic primitives.
-This phase is not completly offline, because the parties still exchange some information which are later needed for the evaluation.
-But this phasae is not that time critically, that means it can be already executed a lot earlier than the online phase, so that, when the function evaluation is really needed, the result can be calculated in a rather fast time compared to a system that still has to set up everything.
-In SCALE-MAMBA, compared to its predecessors, the offline and online phase are completly integrated.
+In the offline phase, prior to the online phase, preparations are done by all parties to prepare for the evaluation.
+This already includes calculating values for possible multiplications and MACs, so that the online phase can be executed faster.
+In this phase the expensive calculations are done, like public key cryptography or authentications, so that the online phase only needs basic primitives.
+The offline phase is not completely offline, because the parties still exchange some information which are later needed for the evaluation.
+But this phase is not that time critically, that means it can be already executed a lot earlier than the online phase, so that, when the function evaluation is really needed, the result can be calculated in a rather fast time compared to a system that still has to set up everything.
+In SCALE-MAMBA, compared to its predecessors, the offline and online phase are completely integrated.
 Therefore, it is not possible anymore to execute them separately like in SPDZ for example.
@@
 
-While SCALE-MAMBA also offers different access structures, the most commonly used security configuration is active security with abort in a full threshold scheme, that means that every except one party could be corrupted, and still no information is leaked, because the evaluation is aborted.
+While SCALE-MAMBA also offers different access structures, the most commonly used security configuration is active security with abort in a full threshold scheme, that means that every party except one could be corrupted, and still no information is leaked, because the evaluation is aborted.
 
 \subsection{MAMBA}
-For this paper the more interesting part is the algorithmic part.
-But still, the importance of the compiler shouldn't be undervalued.
-Because while the algorthmic foundations might be more important from a theoretical point of view, the compiler, written in C++, enables the user to really use these technologies in a system where he can rely on the security gurantees given by the algorithms.
-The programs can be written for MAMBA in a syntax similiar to python's.
+For this paper the most interesting part is the algorithmic.
+But still, the importance of the compiler should not be undervalued.
+Although the algorithmic foundations might be more important from a theoretical point of view, the compiler, written in C++, enables the user to use these technologies in a system where he can rely on the security guarantees given by the algorithms.
+The programs can be written for MAMBA in a syntax similar to python's.
 In these programs, the user can specify which security settings he wishes to use.
 He can specify, how he wants his SMPC to exactly be executed, and which security parameters he wants to use.
-MAMBA is the compiler which then translates the instruction to bytecode and executes them.
+MAMBA compiles the instructions to bytecode which is then executed by SCALE.
 MAMBA is the part that gives the system the possibility to be properly used and tested.
-It makes it easy, to change the parameters, to be able to execute comparable tests, and to have a abstraction level for the user.\\
+It makes it easy to change the parameters, to execute comparable tests, and to have an abstraction level for the user.\\
 
@@
 \item Choose random \((\gamma (a)_1+...+\gamma (a)_n) \in \mathcal{F}_p\) so that $\gamma (a)=((\gamma (a)_1+...+\gamma (a)_n))$
-\item Send to each party \( P_j \) \( \delta_a \) and the shares \(a_j \) and, \( \gamma (a)_j \)
+\item Send  \( \delta_a \) and the respective shares \(a_j \) and \( \gamma (a)_j \) to each party \( P_j \) 
 \item Now each party \( P_j\) holds a share of \( a\) as tuple \(\langle a_j\rangle = (\delta_a,a_j,\gamma(a)_j )\)
@@
 At the same time, no bit of information regarding \(a\) is revealed, as long as not all secret shares are known.\\
-\textbf{Correctness:} Assume that $n-1$ secret shares are known to an adversary, so all secret shares except one $a_x \in \mathcal{F}_p $.
+\textbf{Correctness:} Assume that $n-1$ secret shares are known to an adversary, i.e.
+all secret shares except one $a_x \in \mathcal{F}_p $.
 Then the adversary can calculate $\sum_{i \neq x} a_i = a-a_x$.
@@
 \textbf{Partial Reveal}
-All players send their respective shares of a value, but not the tag share, to one chosen party \(P_i\) which summs up all values and then broadcasts the result.
+All players send their respective shares of a value, but not the tag share, to one chosen party \(P_i\) which sums up all values and then broadcasts the result.
 The tag stays secretly shared between the parties.\\
@@
 There are two main things done in the offline phase.
-The first one is the generation of a global Message Authentification Code(MAC) key $\alpha$.
-This MAC key is later used, to authenticate the different shares at the end of the computation.\\
+The first one is the generation of a global Message Authentication Code(MAC) key $\alpha$.
+This MAC key is used later, to authenticate the different shares at the end of the computation.\\
 The second and and most important part in the offline phase, is the generation of the beaver triples.
@@
 
-This algorithm has the disadvantage of high storage complexity, but therefore it doesn't take a long time to calculate and it is symmetric so the same key can be used for creating the tag and verifying it.
+This algorithm has the disadvantage of high storage complexity because the tag has the same size as the message, which is a difference to other MAC algorithms.
+But it is symmetric so the same key can be used for creating the tag and verifying it.
 The symmetry is used for an advantage later on in the protocol.\\
@@
 
-In the offline phase, ZPoKP are used to guarantee that the Beaver triples are calculated correctly before the sacrificing, or atleast that they dont divert more than a specified error \(e\) from the expected result.
+In the offline phase, ZPoKP are used to guarantee that the Beaver triples are calculated correctly before the sacrificing, or at least that they dont divert more than a specified error \(e\) from the expected result.
 
-ZPoKP are a technique that was introduced  in 1985 \cite{Goldwasser:1985:KCI:22145.22178}.
+ZPoKP is a technique that was introduced  in 1985 \cite{Goldwasser:1985:KCI:22145.22178}.
 The idea is that one party can prove to another party that it knows a specific value \(x\), without revealing the value of \(x\) to the other person, nor revealing any other knowledge about the value.
@@
 Because f is homomorphic, obviously \( f(z)=f(s)+f(e*x)=f(s)+e*f(x)=a+e*y\) with \(f(s)=a\) and \(f(x)=y\).
-So Bob has to calculate \( a+e*y and f(z)\) with the provided \( z\) by Alice. If they are equal, he knows that Alice knows the \( x\) so that \( f(x)=y\).
+So Bob has to calculate \( a+e*y\) and \( f(z)\) with the provided \( z\) by Alice. If they are equal, he knows that Alice knows the \( x\) so that \( f(x)=y\).
 He can do that, because he got provided the \( a\) and \( z\) by Alice, he knew the \( y\) before the protocol and he chose the \( e\) himself.\\
@@
 \subsubsection{First improvement with MASCOT}
-In one of there earlier works, BDOZ, pairwise MACs were used to authentificate the secret sharing between the parties.
+In one of the earlier works, BDOZ, pairwise MACs were used to authentificate the secret sharing between the parties.
 It used a pairwise multiplication protocol with linear homomorpic encryption.

### 2019-04-22 paper.tex
Still not finished

@@
 In most elections there is a third party that has to aggregate and count the votes.
-For instance, in Germany's elections[TODO: LINK] this is done by specific people counting a small amount of votes and then forwarding the information to a central institution.
+For instance in Germany's elections, this is done by specific persons counting the local votes then forwarding the information to a central institution.
 The other problem is that it has to be guaranteed that a vote is valid, meaning that each voter holds the right to vote and is only voting one time.
-So in an ideal system, no person participating in the aggregation process should be able to get any information over any vote or any subset of voters.\\
+So in an ideal system, no person participating in the aggregation process should be able to get any information over any vote or any subset of votersx.\\
 Another use case is the exchange of medical data.
@@
 \subsection{Secret sharing}
+\label{sharing}
 
@@
 GENERATION OF BEAVER TRIPLES: Include protocol on how triples are distributively generated without revealing them
+BGV emcryption scheme
 
-\subsubsection{Zero Proofs of Knowledge}
+For the generation of a beaver triple $(a,b,c)$ , the Brakerski-Gentry-Vaikuntanathan encryption scheme~\cite{cryptoeprint:2011:277} is used, which has homomorphic properties for addition and multiplication.
+It is a public key crypto system, where the parties can only decrypt an encryption together.
+Be $Enc_{pk}$ the encryption with the public key, and $Dec$ the distributed decryption.
+The triples are generated as follows:
+
+\begin{itemize}
+\item Every party $P_i$ samples a random $a_i$ and $b_i$, so that $a=a_1+...+a_n$ and $b=b_1+...+b_n$
+\item Every party computes $e_{a_i}=Enc_{pk}(a_i)$ and $e_{b__i}=Enc_{pk}()$ and broadcasts the values
+\item Every party computes $e_a=e_{a_1}+...+$
+
+\end{itemize}
+
+Generation described in Overdrive
+sample ai,bi
+every pair ij:
+i->j enc(ai)
+pj: cij= bj*enc(a_i)-enc'(random e)
+ci=ao*bi+
+
+Given public key encryption, where the parties can decrypt together the result
+
+random a,b shared
+enc ai, encbi, broadcast
+compute enca, encb
+
+random fi so that f=fi
+compute encf +encm
+dec f+m
+m1=m+f-f, mi=-fi
+
+\subsubsection{Zero-Knowledge Proofs}
 In the many iterations of the SPDZ like protocols, there were different ways to guarantee that the parties are not cheating.
-An important technology for that are the so called Zero Proofs of Konwledge Protocol(ZPoKP).
-
-In the offline phase, ZPoKP are used to guarantee that the Beaver triples are calculated correctly before the sacrificing, or at least that they dont divert more than a specified error \(e\) from the expected result.
-
-ZPoKP is a technique that was introduced  in 1985 \cite{Goldwasser:1985:KCI:22145.22178}.
+An important technology for that is the so called Zero Proofs of Knowledge Protocol (ZPoKP).
+Here, they are used to guarantee that the Beaver triples are not diverting more than a specified error \(e\) from the expected value.
+ZPoKP is a technique that was introduced  in 1985 in~\cite{Goldwasser:1985:KCI:22145.22178}.
 The idea is that one party can prove to another party that it knows a specific value \(x\), without revealing the value of \(x\) to the other person, nor revealing any other knowledge about the value.
 
-Assumed Alice(in this scenario called the prover) wants to prove to Bob (the verifier) that she knows \( x\) in a Field \( \mathcal{F} \), such that \(f(x)=y\).
+Assumed Alice, who is called the prover, wants to prove to Bob, the verifier, that she knows \( x\) in a Field \( \mathcal{F} \), such that \(f(x)=y\).
 It is as well assumed, that \( f \) is homomorphic with respect to the field operations in  \( \mathcal{F} \).
@@
 We also have to make sure, that Alice did not share any information regarding her x.
-In the firt step she chose a random element s and calulated a=f(s). This information is independent on her secret x, so she didn;t share any information.
-When calculating z=s+e*x, she also didnt share any information regarding x, as long it is guaranteed that Bob didnt have any knowledge about the s.
-So under the assumption that Alice chooses a new random s each time the protcol is executed, she doesn't share any information regarding her secret x.\\
+In the first step she chose a random element s and calculated a=f(s). This information is independent on her secret x, so she did not share any information.
+When calculating z=s+e*x, she also did not share any information regarding x, as long it is guaranteed that Bob did not have any knowledge about the s.
+So under the assumption that Alice chooses a new random s each time the protocol is executed, she doesn't share any information regarding her secret x.\\
 
-It is not really important for this thesis, which ZPoKP is exactly used, but it should be said that the protocol used in SCALE-MAMBA is based on Schnorr's protocol.\\
+In SCALE-MAMBA a variant of ZPoKP similar to Schnorr's protocol~\cite{Schnorr} is used, but this paper will not examine the exact protocol.
 
-\subsubsection{First improvement with MASCOT}
-In one of the earlier works, BDOZ, pairwise MACs were used to authentificate the secret sharing between the parties.
-It used a pairwise multiplication protocol with linear homomorpic encryption.
-To guarantee the data integritym pairwise zero-knowledge proofs were used, in total O(n2) of them.
-In SPDZ instead of HE, SHE based on the BGV ecnryption scheme was used, reducing the amount of ZK proofs per triple by a factor of n and as well the overhead per triple, because they used a single ZKPoK for proving thousand of triples.
-Because the offline phase was the weak point of the existing SPDZ implementations, in regards to runtime while still managing a sufficiently secure system, in 2016 MASCOT \cite{cryptoeprint:2016:505} was introduced, which improved the speed of the offline phase by a magnitude of 2.
+\subsubsection{First improvements}
+In one of the earlier works, BDOZ~\cite{cryptoeprint:2010:514}, pairwise MACs were used to authenticate the secret sharing between the parties.
+It used a pairwise multiplication protocol with linear homomorphic encryption.
+To guarantee the data integrity, pairwise zero-knowledge proofs were used, so in total $O(n^2)$ of them, with $n$ being the numbers of parties.
+Because Beaver triples are created secretly shared, this accumulated a high overall run time because a lot of these triples are needed.
+
+In SPDZ, semi homomorphic encryption based on the Brakerski-Gentry-Vaikuntanathan cryptosystem~\cite{cryptoeprint:2011:277} was used.
+Still pairwise ZPoPKs were used, but one proof can be made to prove thousands of triples at once.
+This reduced the amount of proofs needed by a linear factor.
+
+Nevertheless, the offline phase was still the weak point of the existing SPDZ implementations regarding runtime complexity.
+In 2016 MASCOT~\cite{cryptoeprint:2016:505} was introduced, which improved the speed of the offline phase by a magnitude of 2.
 MASCOT used Oblivious Transfer instead of SHE to significantly reduce the amount of communication and computation.
@@
 
+Because the offline phase was the weak point of the existing SPDZ implementations, in regards to runtime while still managing a sufficiently secure system, in 2016 MASCOT \cite{cryptoeprint:2016:505} was introduced, which improved the speed of the offline phase by a magnitude of 2.
+MASCOT used oblivious transfer instead of SHE to significantly reduce the amount of communication and computation.
+We will not go into detail into this protocol, but rather look on its successor, which is really similiar.
+
 \subsubsection{Overdrive}
-Shortly afterwards in 2017 a new protocol for the offline phase called Overdrive was introduced \cite{cryptoeprint:2017:1230}, which once again got better results with SHE instead of OT, improving the run time at least by a factor of 6.
-Basically, it is a really similiar protocol to MASCOT, but just uses SHE in the places where MASCOT used OT.
-So explaining Overdrive already should give a good impression how MASCOT works.
-\\
- Overdrive includes two parts, a protocol called Low Gear protocol for small numbers, and a proctocol called High Gear for larger numbers.
+Shortly afterwards in 2017 a new protocol for the offline phase called Overdrive was introduced in~\cite{cryptoeprint:2017:1230}, which once again used semi homomorphic encryption instead of oblivious transfer.
+It was able to further improve the run time of the offline phase of MASCOT by a factor of 6.
+Overdrive uses the same ideas that were introduced in MASCOT, but applies them to semi homomorphic encryption.
+It includes two parts, a protocol called Low Gear for small numbers, and a protocol called High Gear for larger numbers.
 In version 1.2 of Scale-mamba only the HighGear protocol was used, even for smaller numbers.\\
+The biggest difference between Overdrive and SPDZ is, that the parties not anymore prove multiple smaller statements between each other with ZKPoK, but instead make a joint statement and prove this one toegether.
+Because ZKPoK are computational expensive, this makes a big difference in the overall performance.\\
+This way, Overdrive did not improve the amount of communication channels used, but instead it decreased the computational costs by a linear factor.\\
+The zero proofs of knowledge allow a so called soundness slack~\cite{10.1007/978-3-319-56620-7_17}.
+The idea is that the proof allows a small error in the value, because it is cheaper to prove than to prove the exact value.
+This error is later eliminated by a technique called sacrificing the beaver triples.
 
-The biggest difference between the ZKPoK usage in Overdrive and in SPDZ is, that the parties not anymore prove multiple smaller statements between each other with ZKPoK, but instead make a joint statement and prove thi one toegether.
-Because ZKPoK are really expensive, this makes a big difference in the overall performance.\\
-This way, Overdrive did not improve the amount of communication channels used, but it decreased the computational costs by a factor n.\\
-A second big strategy used by SPDZ and Overdrive is the so called Soundness Slack.
-Because in practice it is to expensive to prove that a result is exact, so it is sufficient to prove that the value, here the product of the beaver triple, exact but with the legalization of a small error.
-Later on, this error, if existing and introduced by an adversary, will be detected by a different technique called sascrificing.\\
-Overdrive uses as well drowning, which is the technique of adding noise to a secret sharing.
-Because otherwise, the noise of the SHE function would reveal information over the secret.
-So the noise is drowned by introducing an additional random noise that is bigger than the maximum possible noise.\\
+Additionally, drowning is used in Overdrive, which is the process of adding random noise to a secret sharing so the encryption does not reveal any information over the secret.
 
 \subsubsection{TopGear}
-Recently, an even newer improvement for the offline phase was introduced, called TopGear \cite{cryptoeprint:2019:035}.
+Recently in 2019, a further improvement for the offline phase was introduced, called TopGear \cite{cryptoeprint:2019:035}.
 It is implemented in SCALE-MAMBA since version 1.3\cite{ScaleMambaDocu}, which was released in January 2019.
-It further improved the HighGear protocol of the Overdrive implementation, by targeting specifically the ZKPoKs.
+It further improved the HighGear protocol of the Overdrive implementation, by targeting specifically the ZKPoKs and increasing the security guarantees provided while maintaining roughly the same performance.
 
@@
 
-Remark: As shown later in chapter [X], addition of two secretly shared values and multiplication of a secretly shared value with a public constant can be done individually by every party, so that, when later revealing the results of each party, the correct result can be computed by the parties together.
+Remark: As shown later in chapter [TODO: X], addition of two secretly shared values and multiplication of a secretly shared value with a public constant can be done individually by every party, so that, when later revealing the results of each party, the correct result can be computed by the parties together.
 Accordingly, beforehand operations can be assumed as working as seen later.
@@
 
-\textbf{Comparison of Sacrificing}
-This small improvement in the protocol nearly halved the mathematical operations necessary for the scarification process.
+\subsubsection{Comparison}
+This small improvement in the protocol of sacrificing nearly halved the mathematical operations necessary for the scarification process.
 This is huge, because the triple generation is one of the most time consuming task, so it sped up the offline phase.
@@
 
-In SPDZ: Sacrifice in online phase???
+Alone the changes in the usage of the zero knowledge proofs already improved the run time of the offline phase, in comparison between SPDZ and SCALE-MAMBA, by a factor of 1000.
+Together with the changes in the the sacrificing and with a lot more smaller changes that will remain undocumented in this paper, the run time of the offline phase improved by a huge margin in the last years.
 
@@
 
-Now it is assumed that the offline phase was succesfully executed, so all preliminaries are given to execute the online phase.
-It's main goal is to evaluate a specified circuit with secret inputs by the different parties.
-So this is the heart of the system, the part where the real computatiuon is executed.
-\\
-It is important to note, that the online phase is nearly perfectly optimized when assuming the offline phase was carried out succesfully.
-The online phase has a runtime complexity in O(n), and therefore there is not a really big performance impact.
+After the offline phase is executed successfully, the online phase takes place, which contains the evaluation of the given function with the provided input secrets.
+It is important to note, that the online phase is nearly perfectly optimized given an earlier execution of the offline phase.
+The online phase has a linear run time complexity depending on the amount of operations needed for the evaluation of the function, and therefore there is not a really big performance impact.
 It is also noteworthy, that in the online phase only computational and no information theoretical problems are solved.
-All this leads to the fact, that there is not really any change in the online phase in the SPDZ like protocols in the last years, therefore the online phase is the same like in the earlier version of SPDZ  ~\cite{cryptoeprint:2011:535}.\\
+All this leads to the fact, that there is not really any change in the online phase in the SPDZ like protocols in the last years, therefore the online phase is the same like in the earlier version of SPDZ  ~\cite{cryptoeprint:2011:535}.
 
@@
 
-First, the secret inputs  $a_1,...,a_m$ are additvely shared with the other parties as described before in [Link to secret sharing in SM XXX].
+First, the secret inputs  $a_1,...,a_m$ are additvely shared with the other parties as described before in chapter \ref{sharing}.
 So now, each party $P_j$ holds a share ${(a_i)}_j$ of every secret $a_i$.\\
@@
 Following, it is going to be demonstrated, how addition and multiplication are done in this environment.
-These two operations are sufficient, because a group with addition and multiplication is already touring complete, so there is no need for any other operation[TODO: ADD reference].
-Furthermore, addition and multiplications with constants are also discussed, because they can be done in a simple way that saves computation time.\\
+These two operations are sufficient, because a group with addition and multiplication is already touring complete, i.e.
+it is possible to express every function with a combination of additions and multiplications.
+Furthermore, addition and multiplications with constants are also discussed, because they can be done in a optimized way that saves computation time.
 

### 2019-04-22 paper.tex
Car commit: nearly finished finally

@@
 
-GENERATION OF BEAVER TRIPLES: Include protocol on how triples are distributively generated without revealing them
-BGV emcryption scheme
+For the generation of a beaver triple $(a,b,c)$, homomorphic public key cryptography based on the Brakerski-Gentry-Vaikuntanathan encryption scheme~\cite{cryptoeprint:2011:277} is used.
+Every party $P_i$ generates for the communication with each party $P_j$ a set of keys $(pk_{ij},sk_{ij})$.
+Then, one triple is generated as follows:
 
-For the generation of a beaver triple $(a,b,c)$ , the Brakerski-Gentry-Vaikuntanathan encryption scheme~\cite{cryptoeprint:2011:277} is used, which has homomorphic properties for addition and multiplication.
-It is a public key crypto system, where the parties can only decrypt an encryption together.
-Be $Enc_{pk}$ the encryption with the public key, and $Dec$ the distributed decryption.
-The triples are generated as follows:
-
-\begin{itemize}
-\item Every party $P_i$ samples a random $a_i$ and $b_i$, so that $a=a_1+...+a_n$ and $b=b_1+...+b_n$
-\item Every party computes $e_{a_i}=Enc_{pk}(a_i)$ and $e_{b__i}=Enc_{pk}()$ and broadcasts the values
-\item Every party computes $e_a=e_{a_1}+...+$
-
-\end{itemize}
-
-Generation described in Overdrive
-sample ai,bi
-every pair ij:
-i->j enc(ai)
-pj: cij= bj*enc(a_i)-enc'(random e)
-ci=ao*bi+
-
-Given public key encryption, where the parties can decrypt together the result
-
-random a,b shared
-enc ai, encbi, broadcast
-compute enca, encb
-
-random fi so that f=fi
-compute encf +encm
-dec f+m
-m1=m+f-f, mi=-fi
+\begin{enumerate}
+\item Every party $P_i$ samples a random $a_i$ and $b_i$; then $a=\sum_{j=1}^n a_i$ and $b=\sum_{j=1}^n b_i$
+\item Every party $P_i$ executes following protocol with each other party $P_j$:
+\begin{enumerate}
+\item $P_i$ sends $P_j$ $Enc_{pk_{ij}}(a_i)$
+\item $P_j$ samples random $e_{ij}$, calculates $g_{ij}=b_j*Enc_{pk_{ij}}(a_i)-Enc_{pk_{ij}}(e_{ij})$ and sends $g_{ij}$ it to $P_i$
+\item $P_i$ decrypts $d_{ij}=Dec_{sk_{ij}}(g_{ij})$
+\end{enumerate}
+\item Every party $P_i$ computes $c_i=a_i*b_i+\sum_{i \neq j} (e_{ij}+d_{ij})$
+\end{enumerate}
+Now, $c$ is correctly shared so that $c=c_1+...+c_n$.\\
+\textbf{Correctness:}
+$$d_{ij}=b_j*a_i-e_{ij}$$
+$$c_i=a_i*b_i+\sum_{i \neq j} (e_{ij}+d_{ij})= a_i*b_i+ \sum_{i \neq j} (b_j*a_i)$$
+$$c_i= a_i*b_i+ a_i*\sum_{i \neq j} (b_j)=a_i*(b_i+ \sum_{i \neq j} b_j)=a_i*b$$
+$$c=c_1+...+c_n=a_1*b+...+a_n*b=(a_1+...+a_n)*b=a*b$$
 

### 2019-04-22 paper.tex
Fertig

@@
 \begin{abstract}
-This paper examines Secure Multi-Party Computation(SMPC), especially the SCALE-MAMBA protocol.
-To do that, it first introduces SMPC, afterwards SPDZ, the predecessor of SCALE-MAMBA, and, after taking a detailed look at SCALE-MAMBA, compares them.
-It focuses on the algorithmic foundations of SCALE-MAMBA, especially the split into online and offline phase, how beaver triples are used to enable computational efficient multiplications and which computational complexities the algorithms posses.
-Afterwards, it tries to depict the evolution in the different versions of the protocol and to locate SCALE-MAMBA between its competitors.
+This paper examines Secure Multi-Party Computation (SMPC), especially the SCALE-MAMBA protocol.
+First it introduces Secure Multi-Party Computation and fundamental primitives, and then takes a detailed look at SCALE-MAMBA, especially in comparison to its predecessor SPDZ.
+It focuses on the algorithmic foundations, especially the split into online and offline phase, how beaver triples are used to enable computational efficient multiplications and which computational complexities the algorithms posses.
+Afterwards, it tries to depict the evolution in the different versions of the protocol and tries to locate SCALE-MAMBA between its competitors.
 Ultimately, it will show some exemplary implementations in real world applications and will give an outlook what further evolutions might be coming in the next years.
@@
 In today's world, and even more in the future, there exist a lot of computations that have to be done in a distributed environment, because in a lot of use cases multiple parties are involved, but also because the amount of data is getting so big that it often becomes impossible to have a single entity that processes and stores all of it.
-Furthermore, in a lot of these cases the parties participating in these computations do not want to reveal their own information, because it is sensitive or it has a high business value.\\
+Furthermore, in a lot of these cases the parties participating in these computations do not want to reveal their own information, because it is sensitive or it has a high business value.
+
 A basic example is an anonymous voting.
@@
 The other problem is that it has to be guaranteed that a vote is valid, meaning that each voter holds the right to vote and is only voting one time.
-So in an ideal system, no person participating in the aggregation process should be able to get any information over any vote or any subset of votersx.\\
+So in an ideal system, no person participating in the aggregation process should be able to get any information over any vote or any subset of voters.
+
 Another use case is the exchange of medical data.
 Medical data of patients is highly confidential and should never be published.
-But at the same time, a lot of different hospitals or doctors have only the data from their small data set of local patients, and for making founded diagnosis for new patients it would be helpful to compare the symptoms to the diagnoses of other doctors.\\
-Secure Multi-Party Computation(SMPC) is one possible solution for these use cases. It is a field of research that emerged within the last 30 years, and is gathering a lot of speed in the last years. 
+But at the same time, a lot of different hospitals or doctors have only the data from their small data set of local patients, and for making founded diagnosis for new patients it would be helpful to compare the symptoms to the diagnoses of other doctors.
+
+Secure Multi-Party Computation (SMPC) is one possible solution for these use cases, because we can model above mentioned and a lot similar problems as functions.
+SMPC is a field of research that emerged within the last 30 years, and is gathering a lot of speed in the last years.
 The basic problem is the evaluation of a common function, to which multiple parties have different inputs, without revealing these inputs.
 Although there are a lot of theoretical solutions for this problem, real-world implementations are still not performing on the desired level.
-So in the last years the research shifted more into the area of how to improve the performance of existing solutions while still guaranteeing a desired level of security.\\
-This paper's main focus will lay on SCALE-MAMBA, a system developed mainly by the Katholieke Universiteit Leuven.
-This system is one of the forerunners in SMPC and combines a lot of research by different teams to provide a platform with highly performant SMPC.
+So in the last years the research shifted more into the area of how to improve the performance of existing solutions while still guaranteeing a desired level of security.
+
+This paper's main focus will lay on SCALE-MAMBA~\cite{ScaleMambaDocu}, a system developed mainly by the Katholieke Universiteit Leuven.
+This system is one of the forerunners in this area and combines a lot of research by different teams to provide a platform with highly performant SMPC.
 Nevertheless, this paper will at first give a short introduction into the history of SMPC and will try to explain its basic technologies to the reader.
 It will compare SCALE-MAMBA to its predecessors and will also try to give a short comparison to other SMPC systems which use different approaches, especially taking a look on which different requirements are fulfilled by each.
-Ultimately, it will try to give an outlook what might be coming in the next years.\\
+Ultimately, it will try to give an outlook what might be coming in the next years.
 \section{SMPC Basics and History}\label{sec:smpc}
 \subsection{Beginnings}
-As we saw in the introduction, there are a lot current and much more upcoming use cases of calculating the output of a function in a distributed environment with multiple parties while at the same time keeping all the individual inputs secret.
-The first explorations in these fields started in the 1970s, with theoretical problems like the mental poker problem(how can two players play poker while not being in the same physical location and making sure no one is cheating) and Yao's millionaire problem(how can can two millionaires decide who is richer than the other while not revealing their respective wealth).
-\\
-These explorations lead to a group of basic protocols which established the area of Secure Multi-Party Computation(SMPC).\\
+As we saw in the introduction, there are a lot current, and much more upcoming, use cases of calculating the output of a function in a distributed environment with multiple parties while at the same time keeping all the individual inputs secret.
+The first explorations in these fields started in the 1970s, with theoretical problems like the mental poker problem~\cite{mentalPoker}, that asks how two players can play poker while not being in the same physical location and making sure no one is cheating, and Yao's millionaire problem~\cite{Yao1982ProtocolsFS}, which tries to answer the question which of two millionaires is richer than the other while not revealing their respective wealth.
+
+These explorations led to a group of basic protocols which established the area of Secure Multi-Party Computation(SMPC).\\
 \subsection{Problem statement}
@@
 \textbf{Correctness:} At the end of the computation, the result is correct.\\\\
-In a lot of cases each party has exactly one corresponding secret input, but this is not necessarily required. There can also be parties without input and parties with multiple inputs.\\
+In a lot of cases each party has exactly one corresponding secret input, but this is not necessarily required. There can also be parties without input and parties with multiple inputs.
+
 Now we will take a short look on two of the first protocols, which both are the basis for a whole family of SMPC schemes.
@@
 There are typically two families, into which most of the SMPC protocols can be categorized.
-One is based on garbled circuits which were first introduced by Yao in 1986 \cite{4568207}.
-The basic idea is that, in a two party setting, one party, Alice, garbles the Boolean circuit that has to be evaluated and sends it to the second party, Bob.
+One is based on garbled circuits which were first introduced by Yao in 1986~\cite{4568207}.
+The basic idea is that, in a two party setting, one party, Alice, encrypts the Boolean circuit that has to be evaluated and sends it to the second party, Bob.
 Bob uses his own input, Alice's encrypted input and the garbled circuit to evaluate the circuit in an encrypted way.
@@
 \subsection{Shamirs Secret Sharing Scheme}
-This paper will not further discuss the family of garbled circuit protocols, but rather the second family which are protocols based on so called Linear Secret Sharing. 
-These type of protocols were introduced by Adi Shamir at a similar time, in 1979 \cite{Shamir:1979:SS:359168.359176}.\\
-The basic idea of this approach is the the so called share-compute-reveal paradigm, which splits the evaluation of the function into three parts.\\
-\textbf{Share:} First, the inputs secrets are split into multiple parts which then are shared shared with the other parties. Hence, every party holds a share of each input secret.\\
+This paper will not further discuss the family of garbled circuit protocols, but rather the second family, which are protocols based on so called Linear Secret Sharing. 
+These type of protocols were introduced by Adi Shamir at a similar time, in 1979~\cite{Shamir:1979:SS:359168.359176}.
+
+The basic idea of this approach is the the so called share-compute-reveal paradigm, which splits the evaluation of the function into three parts.\\\\
+\textbf{Share:} First, the inputs secrets are split into multiple parts which then are shared with the other parties. Hence, every party holds a share of each input secret.\\
 \textbf{Compute:} The parties evaluate the function with the secret input shares.
-The protocol has to be defined in a way, that the final result which each parties computes with the input shares is a respective share of the complete final resut.\\
-\textbf{Reveal:} The parties agregate their shares of the final result and compute the final result together.\\
+The protocol has to be defined in a way, that the final result which each parties computes with the input shares is a respective share of the complete final result.\\
+\textbf{Reveal:} The parties aggregate their shares of the final result and compute the final result together.
+
 This paradigm already shows a lot of challenges arising for a concrete protocol, for instance:
 \begin{itemize}
-\item How can the secrets be shared in a homomorphic way, so that, after computing on shared input, the complete result can be recovered from the result shares?
+\item How can the secrets be shared in a homomorphic way, so that, after computing on shared inputs, the complete result can be recovered from the result shares?
 \item While this can be done for addition in easy ways, how can this be done for multiplication as well?
 \item How can the correctness of the computation be verified?
-\item How can the overhead generated by a protocol can be kept in a reasonable size, so that a protocol is applicable to a real-world use case?
+\item How can the overhead generated by a protocol can be kept on a reasonable scale, so that a protocol is applicable to a real-world use case?
 \end{itemize}
@@
 This is done by using Lagrange interpolation.
-.
-The degree $p$ of the polynomial can be seen as a security parameter, because $p$ shares are needed to reconstruct the secret.\\
+The degree $p$ of the polynomial can be seen as a security parameter, because $p+1$ shares are needed to reconstruct the secret.
+
 There is a huge family of protocols build on this approach, and it is still evolving every day.
-This paper will focus on the newest of these protocols that were developed in the last 10 years, especially on the protocols which are based on or similar to SPDZ.
-\\
+This paper will focus on the newest of these protocols that were developed in the last 10 years, especially on the protocols which are based on or similar to SPDZ~\cite{cryptoeprint:2011:535} which was published in 2011.
+
 Before looking into exact implementations of linear secret sharing schemes, following, a short look will taken at the general adversary model and required basic definitions.
 \subsection{Adversary Model}
-In cryptography, the adversary which is considered is normally a third party that wants to get information for his own gain or want to sabotage something because of his own interest.
-In SMPC it is expected that the adversary is not only participating as one party, but could could control multiple parties.
+In cryptography, the adversary which is considered is normally a third party that wants to get information for his own gain or want to sabotage a protocol because of his own interest.
+In SMPC it is expected that the adversary is not only participating as one party, but could control multiple parties.
 That means a potential adversary can control between one and all but one of the parties which are participating.
-The difference to the normal cryptographic adversary model is, that the adversary is controlling parties that are participating in the scheme and not a third party that tries to intercept communication.\\
+The difference to the normal cryptographic adversary model is, that the adversary is controlling parties that are participating in the scheme and not a third party that tries to intercept communication.
+
 In the area of SMPC there is normally a differentiation between two types of adversaries.
-\subsubsection{Honest but curious}
-The honest but curious adversary tries to gain as much information as possible while obeying the rules of the protocol.
+\subsubsection{Honest-but-curious}
+The honest-but-curious adversary tries to gain as much information as possible while obeying the rules of the protocol.
 Hence, it is impossible to know that this party is controlled by a third entity.
@@
 \subsubsection{Malicious adversary}
-While this adversary also wants to gain all the information, he is not obeying the rules of the protocol.
+While this adversary also wants to gain all possible information, he is not obeying the rules of the protocol.
 This type of adversary can send different inputs to manipulate the results he receives.
 Furthermore, he wants to interrupt the whole protocol.
-That could mean making the computation impossible, but as well modifying his own input or his own calculations, so the final result is changed and might lead to different implications
-It is important to distinguish between these two types of adversary, because it shows the different securities SMPC has to guarantee.
-It has to guarantee that no information is unwillingly shared with an illegal party and at the same time verify that the final result has a correct value.
-Most SMPC protocols try to guarantee safety versus both type of adveraries, bvut there are also approaches that only use passive security, i.e.
-only protecting versus non malicious adversaries.
+That could mean making the computation impossible, but as well modifying his own input or his own calculations, so the final result is changed and might lead to different implications.
+
+It is important to distinguish between these two types of adversary, because different measures are needed for each adversary.
+Additionally, some protocols only provide passive security, which means they protect against a honest-but-curious adversary but not against a malicious one.
 \subsection{Access structures}
 There are a lot of different use cases for SMPC, so of course the adversary model changes depending on the system, the security needs, and the amount of parties involved.
-it is important to differentiate between these different use cases, because the different needs have big implications for security measurements that have to be taken and, thus, for runtime complexity.
-SMPC should run as fast as possible, while still providing all the necessary security guarantees.\\
+It is important to differentiate between these different use cases, because the different needs have big implications for security measures that have to be taken and, thus, for run-time complexity.
+SMPC should run as fast as possible, while still providing all the necessary security guarantees.
+
 To describe which party has which rights, and how many parties can be corrupted, so that the security can still be guaranteed, access structures were introduced.
@@
 An access structure is called \(Q_l \)-access structure if at least \(l+1\) parties are needed to  form a qualified set.
-An \((n,t)\)-threshold scheme with \(t<\frac{n}{l}\) means that at least \(t+1\) parties are needed to form a qualified set.
-Here, \(n\) is the number of total parties, and \(t\) is the number of possible adversaries while not revealing any secrets ~\cite{MPCBook_threshold}.\\
+An \((n,t)\)-threshold scheme with \(t<\frac{n}{l}\) means that at least \(t+1\) parties are needed to form a qualified set~\cite{MPCBook_threshold}.
+Here, \(n\) is the number of total parties, and \(t\) is the number of possible adversaries while not revealing any secrets.
+
 A full threshold scheme means that all parties are needed to form a qualified set.
@@
 Of course this has a huge impact on overhead and performance, but in some cases it can provide needed security guarantees.
+
 An access structure also does not have to be symmetric.
-In some cases there might be different type of parties, where different parties have different positions or levels of trust, so it is, for example, possible to specify that some set of parties only has to have the size of 2 to form a qualified set and another one needs the size of 4.\\
+In some cases there might be different type of parties, where different parties have different positions or levels of trust, so it is, for example, possible to specify that some set of parties only has to have the size of 2 to form a qualified set and another one needs the size of 4.
+
 Choosing the right access structure for the right use case is important, because it has a big impact on performance and on the desired security level.
-
-\section{SCALE-MAMBA}\label{sec:scalemamba}
-
+\section{SCALE-MAMBA}
+\label{sec:scalemamba}
 \subsection{Overview}
-
 In the last years, research in SMPC is constantly ongoing, and there are new approaches and improvements to old approaches appearing regularly.
 But most of these publications only focus on a small area of the whole picture.
-SCALE-MAMBA\cite{ScaleMambaDocu} is a project that tries to combine the state-of-the-art research into one system.
-That means that it is offering an end-to-end system for SMPC.\\
+SPDZ~\cite{cryptoeprint:2011:535} is a system that appeared in 2011 and, with a lot of evolutions, was one of the most refined SMPC schemes.
+
+SCALE-MAMBA\cite{ScaleMambaDocu} is the newest successor of SPDZ and is a project that tries to combine most of the state-of-the-art research into one system.
+This implies that it is offering an end-to-end system for SMPC.
+
 It is a research system that is continuously in development, therefore, there is not a single final version that can be examined.
@@
 That also means for this paper, that there is not a single protocol that can be examined.
-Thus, this paper tries to give more of an overview how the process works in general, which are the main, especially newest and best performing, components and where are the problems and bottlenecks.\\
+Thus, this paper tries to give more of a overview how the process works in general, which are the main, especially newest and best performing, components and where are the problems and bottlenecks.
+This overview is only selective and by no means complete.
+
 SCALE-MAMBA consists of two parts.
@@
 In the offline phase, prior to the online phase, preparations are done by all parties to prepare for the evaluation.
-This already includes calculating values for possible multiplications and MACs, so that the online phase can be executed faster.
-In this phase the expensive calculations are done, like public key cryptography or authentications, so that the online phase only needs basic primitives.
+This already includes calculating values for possible multiplications and message authentication codes, so that the online phase can be executed faster.
+In the offline phase the expensive calculations are done, like public key cryptography or authentications, so that the online phase only needs basic primitives.
 The offline phase is not completely offline, because the parties still exchange some information which are later needed for the evaluation.
@@
 In SCALE-MAMBA, compared to its predecessors, the offline and online phase are completely integrated.
-Therefore, it is not possible anymore to execute them separately like in SPDZ for example.
-But it is a much closer approach to a real world application.\\
+Therefore, it is not possible anymore to execute them separately like it was in SPDZ for example.
+This is done to be closer to a real-world implementation.
 
 While SCALE-MAMBA also offers different access structures, the most commonly used security configuration is active security with abort in a full threshold scheme, that means that every party except one could be corrupted, and still no information is leaked, because the evaluation is aborted.
-
 \subsection{MAMBA}
@@
 MAMBA compiles the instructions to bytecode which is then executed by SCALE.
-MAMBA is the part that gives the system the possibility to be properly used and tested.
-It makes it easy to change the parameters, to execute comparable tests, and to have an abstraction level for the user.\\
-
+Because of MAMBA, it is much more easy to test and compare different configurations.
+This abstraction layer makes the system more usable and analyzable.
 \subsection{Secret sharing}
 \label{sharing}
+As mentioned before, SCALE-MAMBA and similar protocols implement the share-compute-reveal paradigm. The sharing is implemented as follows:
 
-As mentioned before, SCALE-MAMBA, and in general similiar protocols like SPDZ, use linear secret sharing.\\
-Given party \(P_i\) which holds a secret value \( a \in \mathcal{F}_p\) and a share of the global MAC key \( \alpha\).
+Given a party \(P_i\) which holds a secret value \( a \in \mathcal{F}_p\) and a share of the global MAC key \( \alpha\).
 Then \( P_i\) can additvely share \( a\) with the other \(n-1\) parties as follows:
@@
 \end{enumerate}
-
 Now the value of \(a\) and the corresponding tag \(\gamma(a)\) can only be reconstructed if, and only if, all \(n\) secret shares are known.
@@
 Nevertheless, because $a_x$ was drawn from an independent distribution and the adversary has no information over the value of \(a_x\), knowing $a-a_x$ reveals not a single bit of information about $a$.\\
-
 \subsection{Revealing secrets}
-
-For obvious reasons, the secrets also have to be revealed again. SCALE-MAMBA uses two different types of revealing them.\\
-
+SCALE-MAMBA uses two different types of revealing a secretly shared value.\\\\
 \textbf{Partial Reveal}
@@
 \textbf{Full Reveal}
-In this case all players broadcast their respective shares and the respective MACs to all players, so all players can calculate the secret and verify the result with the corresponding tag.\\
-
+In this case all players broadcast their respective shares and the respective MACs to all players, so all players can calculate the secret and verify the result with the corresponding tag.
 \subsection{Offline phase}
-
+The offline phase prepares everything, so the function can be evaluated as efficient as possible in the online phase.
 There are two main things done in the offline phase.
 The first one is the generation of a global Message Authentication Code(MAC) key $\alpha$.
-This MAC key is used later, to authenticate the different shares at the end of the computation.\\
-The second and and most important part in the offline phase, is the generation of the beaver triples.
-These triples  are independent of the computation that is later performed in the online phase, but are used in the online phase to express multiplications as a combination of additions and multiplications with constants.
+This MAC key is used later, to authenticate the different shares at the end of the computation.
+
+The second and and most important part in the offline phase, is the generation of the beaver triples~\cite{10.1007/3-540-46766-1_34}.
+These triples are needed in the online phase to express multiplications as a combination of additions and multiplications with constants.
 For each multiplication, one of the triples is needed.
-
 \subsubsection{Generation of the MAC key}
@@
 For using MACs we need a tupel of two algorithms \((\text{MAC},\text{VER})\):\\
-\begin{itemize}
+\begin{enumerate}
 \item The MAC algorithm \(\text{MAC}(K,m)\rightarrow t\) takes a key \(K\) and a message \(m\) and outputs a tag \(t\)\\
 \item The verficiation algorithm \(\text{VER}(K,m,t)\rightarrow 0/1\) takes a key \(K\), a message \(m\) and a tag \(t\) and outputs accept(\(1\)) if the tag is correct or reject(\(0\))\\ if it is not
-\end{itemize}
-
+\end{enumerate}
 In SCALE-MAMBA really simple MACs are used. We assume that we have a MAC key \( \alpha \). Then the MAC algorithm used is following:\\
@@
 $$\text{VER}(\alpha,m,t)\rightarrow (1 \Leftrightarrow t = \alpha*m) $$ 
+Compared to other MAC algorithms, this algorithm has high storage complexity because the tag has the same size as the message.
+But it has other advantages, which are that it is not computation complex with only one multiplication, it is symmetric and it is homomorphic.
 
-This algorithm has the disadvantage of high storage complexity because the tag has the same size as the message, which is a difference to other MAC algorithms.
-But it is symmetric so the same key can be used for creating the tag and verifying it.
-The symmetry is used for an advantage later on in the protocol.\\
-
-The MAC key is distributively generated, i.e. only if all parties pool together their shares they can recover its value. 
-[TODO Expand on MAC generation
-
+The MAC key is distributively generated, so that every party has a additive share $\alpha_i$. Only at the end of the computation, the parties aggregate their shares to verify the result.
 \subsubsection{Beaver Triples}
-
-Beaver triples are triples of three numbers \((a,b,c)\), with \( a,b,c\in \mathcal{F}_p \), with \(a\) and \(b\) being chosen uniformly randomly, and \(a*b=c\).
+Beaver triples, introduced in~\cite{10.1007/3-540-46766-1_34}, are triples of three numbers \((a,b,c)\), with \( a,b,c\in \mathcal{F}_p \), with \(a\) and \(b\) being chosen uniformly randomly, and \(a*b=c\).
 In the offline or pre-processing phase, the most important part is the generation of the beaver triples.
-These triples  are independent of the computation that is later performed in the online phase, but will be used in the online phase to guarantee that multiplications can be done without revealing the secret inputs of the different parties.\\
+These triples  are independent of the computation that is later performed in the online phase, but will be used in it to guarantee that multiplications can be done without revealing the secret inputs of the different parties.
 
@@
 Both triples are additively shared between all parties like the input values.
-In SPDZ both triples are completely unrelated to each other, and the sacrificing is done as follows:\\
+In SPDZ both triples are completely unrelated to each other, and the sacrificing is done as follows:
 
@@
 Then, one triple is generated as follows:
-
 \begin{enumerate}
@@
 $$c=c_1+...+c_n=a_1*b+...+a_n*b=(a_1+...+a_n)*b=a*b$$
-
 \subsubsection{Zero-Knowledge Proofs}
 In the many iterations of the SPDZ like protocols, there were different ways to guarantee that the parties are not cheating.
-An important technology for that is the so called Zero Proofs of Knowledge Protocol (ZPoKP).
+An important technology for that are the so called Zero-Knowledge Proofs(ZKP).
 Here, they are used to guarantee that the Beaver triples are not diverting more than a specified error \(e\) from the expected value.
-ZPoKP is a technique that was introduced  in 1985 in~\cite{Goldwasser:1985:KCI:22145.22178}.
+ZKPs are a technique that was introduced  in 1985 in~\cite{Goldwasser:1985:KCI:22145.22178}.
 The idea is that one party can prove to another party that it knows a specific value \(x\), without revealing the value of \(x\) to the other person, nor revealing any other knowledge about the value.
@@
 \end{enumerate}
-
 \textbf{Correctness:}
@@
 He can do that, because he got provided the \( a\) and \( z\) by Alice, he knew the \( y\) before the protocol and he chose the \( e\) himself.\\
-
 \textbf{Confidentiality:}
@@
 When calculating z=s+e*x, she also did not share any information regarding x, as long it is guaranteed that Bob did not have any knowledge about the s.
-So under the assumption that Alice chooses a new random s each time the protocol is executed, she doesn't share any information regarding her secret x.\\
+So under the assumption that Alice chooses a new random s each time the protocol is executed, she doesn't share any information regarding her secret x.
 
-In SCALE-MAMBA a variant of ZPoKP similar to Schnorr's protocol~\cite{Schnorr} is used, but this paper will not examine the exact protocol.
-
+In SCALE-MAMBA a variant of ZKPs similar to Schnorr's protocol~\cite{Schnorr} is used, but this paper will not examine the exact protocol.
 \subsubsection{First improvements}
@@
 In SPDZ, semi homomorphic encryption based on the Brakerski-Gentry-Vaikuntanathan cryptosystem~\cite{cryptoeprint:2011:277} was used.
-Still pairwise ZPoPKs were used, but one proof can be made to prove thousands of triples at once.
+Still pairwise ZKPs were used, but one proof can be made to prove thousands of triples at once.
 This reduced the amount of proofs needed by a linear factor.
 
-Nevertheless, the offline phase was still the weak point of the existing SPDZ implementations regarding runtime complexity.
+Nevertheless, the offline phase was still the weak point of the existing SPDZ implementations regarding run time complexity.
 In 2016 MASCOT~\cite{cryptoeprint:2016:505} was introduced, which improved the speed of the offline phase by a magnitude of 2.
@@
 
-Because the offline phase was the weak point of the existing SPDZ implementations, in regards to runtime while still managing a sufficiently secure system, in 2016 MASCOT \cite{cryptoeprint:2016:505} was introduced, which improved the speed of the offline phase by a magnitude of 2.
+Because the offline phase was the weak point of the existing SPDZ implementations, in regards to run time while still managing a sufficiently secure system, in 2016 MASCOT \cite{cryptoeprint:2016:505} was introduced, which improved the speed of the offline phase by a magnitude of 2.
 MASCOT used oblivious transfer instead of SHE to significantly reduce the amount of communication and computation.
-We will not go into detail into this protocol, but rather look on its successor, which is really similiar.
-
+We will not go into detail into this protocol, but rather look on its successor, which is really similar.
 \subsubsection{Overdrive}
@@
 It includes two parts, a protocol called Low Gear for small numbers, and a protocol called High Gear for larger numbers.
-In version 1.2 of Scale-mamba only the HighGear protocol was used, even for smaller numbers.\\
+In version 1.2 of Scale-mamba only the HighGear protocol was used, even for smaller numbers.
+
 The biggest difference between Overdrive and SPDZ is, that the parties not anymore prove multiple smaller statements between each other with ZKPoK, but instead make a joint statement and prove this one toegether.
-Because ZKPoK are computational expensive, this makes a big difference in the overall performance.\\
-This way, Overdrive did not improve the amount of communication channels used, but instead it decreased the computational costs by a linear factor.\\
+Because ZKPs are computational expensive, this makes a big difference in the overall performance.
+
+This way, Overdrive did not improve the amount of communication channels used, but instead it decreased the computational costs by a linear factor.
+
 The zero proofs of knowledge allow a so called soundness slack~\cite{10.1007/978-3-319-56620-7_17}.
@@
 Additionally, drowning is used in Overdrive, which is the process of adding random noise to a secret sharing so the encryption does not reveal any information over the secret. 
-
 \subsubsection{TopGear}
@@
 It is implemented in SCALE-MAMBA since version 1.3\cite{ScaleMambaDocu}, which was released in January 2019.
-It further improved the HighGear protocol of the Overdrive implementation, by targeting specifically the ZKPoKs and increasing the security guarantees provided while maintaining roughly the same performance.
-
+It further improved the HighGear protocol of the Overdrive implementation, by targeting specifically the ZKPs and increasing the security guarantees provided while maintaining roughly the same performance.
 \subsubsection{Sacrificing of triples}
@@
 This is done with a second triple \( (\langle x_2 \rangle ,\langle y_2\rangle ,\langle z_2\rangle )\) in a procedure called sacrificing, which uses the second triple to guarantee the soundness of the first.
-In SPDZ that is done as follows:\\
+In SPDZ that is done as follows:
 \begin{enumerate}
@@
 \item If \( \phi \) is 0 then the triples are correct and it is proceeded with \( (\langle x_1 \rangle ,\langle y_1\rangle , \langle z_1\rangle )\) as a correct, secretly shared and not revealed triple.
-Otherwise the system detected a malicious adversary and it aborts.\\
+Otherwise the system detected a malicious adversary and it aborts.
 \end{enumerate}
-
-Remark: As shown later in chapter [TODO: X], addition of two secretly shared values and multiplication of a secretly shared value with a public constant can be done individually by every party, so that, when later revealing the results of each party, the correct result can be computed by the parties together.
+Remark: As shown later in chapter \ref{online}, addition of two secretly shared values and multiplication of a secretly shared value with a public constant can be done individually by every party, so that, when later revealing the results of each party, the correct result can be computed by the parties together.
 Accordingly, beforehand operations can be assumed as working as seen later.
@@
 $$= t* (z_1 - x_1*y_1) + x_2*y_2 - z_2 $$
-Therefore, \( \phi = 0 \Leftrightarrow x_1 * y_1 = z_1 \land x_2*y_2=z_2 \), so this only holds if both triples are correct. \\
-In this process the triple \( (\langle x_1 \rangle ,\langle y_1\rangle , \langle z_1\rangle ) \) wasn't revealed, so now it can be used for the multiplication knowing that one the one hand it is still secretly shared and unknown and on the other hand that it was not modified by an adversary.\\
+Therefore, \( \phi = 0 \Leftrightarrow x_1 * y_1 = z_1 \land x_2*y_2=z_2 \), so this only holds if both triples are correct.
 
+In this process the triple \( (\langle x_1 \rangle ,\langle y_1\rangle , \langle z_1\rangle ) \) wasn't revealed, so now it can be used for the multiplication knowing that one the one hand it is still secretly shared and unknown and on the other hand that it was not modified by an adversary.
 \textbf{Sacrificing of triples in SCALE-MAMBA}
 
-In later versions of SPDZ, since ~\cite{cryptoeprint:2017:123}, and in the current implementations of SCALE-MAMBA, a slightly modified version of aforementioned protocol is used.
+In later versions of SPDZ, since ~\cite{cryptoeprint:2017:1230}, and in the current implementations of SCALE-MAMBA, a slightly modified version of aforementioned protocol is used.
 The idea is to not sacrifice a random second triple, to guarantee the correctness of the first one, but to sacrifice a similar triple.
@@
 \item If \( \phi \) is 0 then the triples are correct and it is proceeded with \( (\langle x \rangle ,\langle y\rangle , \langle z\rangle )\) as a correct, secretly shared and not revealed triple.
-Otherwise the system detected a malicious adversary and it aborts.\\
+Otherwise the system detected a malicious adversary and it aborts.
 \end{enumerate}
-
-\textbf{Correctness}: Parallel to before, simplifying \( \phi \) leads to the same implications:\\
+\textbf{Correctness}: Parallel to before, simplifying \( \phi \) leads to the same implications:
 $$\phi =r*z-z'-p*y=r*z-z'-(r*x-x')*y=r*(z-x*y)+(x'*y-z')$$
 Therefore, \( \phi = 0 \Leftrightarrow x * y = z \land x'*y=z' \)
-
 \subsubsection{Comparison}
@@
 This is huge, because the triple generation is one of the most time consuming task, so it sped up the offline phase.
-In following table, the differences between both sacrificing protocols can be seen:
+In following table, the differences in needed operations between both sacrificing protocols can be seen:
 
@@
 \subsection{Online Phase}
-
+\label{online}
 \subsubsection{Preliminaries}
-
 After the offline phase is executed successfully, the online phase takes place, which contains the evaluation of the given function with the provided input secrets.
@@
 It is also noteworthy, that in the online phase only computational and no information theoretical problems are solved.
-All this leads to the fact, that there is not really any change in the online phase in the SPDZ like protocols in the last years, therefore the online phase is the same like in the earlier version of SPDZ  ~\cite{cryptoeprint:2011:535}.
+All this leads to the fact, that there is not really any change in the online phase in the SPDZ like protocols in the last years, therefore the online phase is the same like in the earlier version of SPDZ~\cite{cryptoeprint:2011:535}.
 
-We assume following preliminaries are given after the execution of the offline phase:\\
+We assume following preliminaries are given after the execution of the offline phase:
 \begin{enumerate}
@@
 Furthermore, addition and multiplications with constants are also discussed, because they can be done in a optimized way that saves computation time.
-
 \subsubsection{Addition}
@@
 Naturally, both of these values are additively shared between all the parties, i.e.
-every party has a share \( \langle a_i\rangle\) and a share \( \langle b_i\rangle\).\\
+every party has a share \( \langle a_i\rangle\) and a share \( \langle b_i\rangle\).
 
@@
 \end{enumerate}
-
+\textbf{Runtime:} The runtime is constant, because every party has to do 3 additions per execution of the protocol.
 \subsubsection{Addition with constant}
@@
 \textbf{Correctness:} It can be seen, that the constant is only added to one share $a_1$ of the value and subtracted from \( \delta_a\). \( \langle a+c \rangle \) is properly secretly shared because:
-\begin{itemize}
+\begin{enumerate}
 \item $a+c=(a_1+...+a_n)+c=(a+1+c)+a_2+...+a_n$
 \item \( \gamma(a)=\alpha*(a+\delta_a)=\alpha*((a+c)+(\delta_a-c))=\gamma(a+c)\)
-\end{itemize}
-
+\end{enumerate}
+\textbf{Runtime:} The runtime is constant, because every party has to do 1 or 2 additions per execution of the protocol.
 \subsubsection{Multiplication with constant}
@@
 \end{enumerate}
+\textbf{Runtime:} The runtime is constant, because every party has to do 3 multiplications per execution of the protocol.
 
 \subsubsection{Multiplication}
+Multiplication is the only operation, except for distributing the shares and calculating and verifying the end result, where the parties have to interact in the online phase.
 
-Multiplication is the only operation, except for distributing the shares and calculating and verifying the end result, where the parties have to interact in the online phase.
 Assumed there are two secretly shared values \( \langle a \rangle\) and \( \langle b \rangle\) which should be multiplied with each other.
-Naturally, each party has the shares \(\langle a_i\rangle \) and \(\langle b_i\rangle \).\\
-For each multiplication operation, we need one of the Beaver triples \( (\langle x_1 \rangle ,\langle y_1\rangle , \langle z_1\rangle )\) generated in the offline phase.
-In the offline phase we checked the quality of our triples, but we still allowed the possibility that each triple has an error \(e\) so that \(c=a*b+e\).
-To check the quality of our triple, we use a second triple \( (\langle x_2 \rangle ,\langle y_2\rangle ,\langle z_2\rangle )\) in a procedure called sacrificing it.
-Both triples are additively shared between all parties like the input values.
-In SPDZ both triples are completly unrelated to each other, and the sacrificing is done as follows:\\
+Naturally, each party $P_i$ holds the shares \(\langle a_i\rangle \) and \(\langle b_i\rangle \).
 
-\begin{enumerate}
-\item Choose a random variable \( r\) and open it to the other parties. This is done by each party generating a random share \(r_i\) and then calculating \(r=r_1+...+r_n)\).
-\item The parties calculate  \( p=t*\langle x_1 \rangle - \langle x_2 \rangle \) and \( \sigma=\langle y_1 \rangle -  \langle y_2 \rangle \) and partially open the results afterwards
-\item The parties calculate \(\phi = t * \langle z_1 \rangle\ - \langle z_2 \rangle - \sigma * \langle x_2 \rangle - p * \langle y_2 \rangle - \sigma * p)\) and then partially open the result
-\item If \( \phi \) is 0 then the triples are ok and it is proceeded with \( (\langle x_1 \rangle ,\langle y_1\rangle , \langle z_1\rangle )\).
-Otherwise the system detected a malicious adversary and it aborts.\\
-\end{enumerate}
+For each multiplication operation, one of the secretly shared Beaver triples \( (\langle x \rangle ,\langle y\rangle , \langle z\rangle )\) generated in the offline phase is needed, which is taken out of the queue.
 
-\textbf{Correctness}: If we simplify \( \phi \), we can see that every term disappears (given that the triples are proper triples):
-$$\phi = t*z_1 - z_2 - y_1 + y_1*x_2 + y_2 * x_2 - t * x_1 * y_2 + x_2*y_2 - t*x_1*y_1 + t*x_1*y_2 + x_2*y_1 - x_2*y_2$$
-$$ = t*(z_1 - x_1*y_2+x_1*y_2 -x_1*y_1) + y_1*x_2 - x_2*y_1+ x_2*y_2 - z_2 + x_2*y_2 -  x_2*y_2$$
-$$= t* (z_1 - x_1*y_1) + x_2*y_2 - z_2 $$
-So we can see that \( \phi = 0 \Leftrightarrow x_1 * y_1 = z_1 \land x_2*y_2=z_2 \), so we know the triple is correct. \\
-In this process we didn't reveal the triple \( (\langle x_1 \rangle ,\langle y_1\rangle , \langle z_1\rangle ) \), so now we can use it for the multiplication knowing that one the one hand it is still secretly shared and unknown and on the other hand that it was not modified by an adversary.\\
-
-There is a small improvement in SCALE-MAMBA compared to SPDZ regarding the triple generation that was added later in ~\cite{cryptoeprint:2017:123}.
-The idea is to not sacrifice a random second triple, to guarantee the correctness of the first one, but to sacrifice a similiar triple.
-For this, the triples in the online phase are already generated in pairs, so that there are are two triples \( (\langle x \rangle ,\langle y \rangle , \langle x*y \rangle ) \) and \( (\langle x' \rangle ,\langle y\rangle , \langle x'*y\rangle ) \).
-This way the triples can be calculated faster in the offline phase, because one half of the multiplication is the same, and still the triples keep the property to be sacrificed for each other to check the correctness of the other triple.\\\\
-
-In SCALE-MAMBA, this triple sacrifice is also already executed in the offline phase.
-But for consistency the explanation remains here, to better be able to see the difference between SPDZ and SCALE-MAMBA.\\
-
-Multiplication is the only operation, except for distributing the shares and calculating and verifying the end result, where the parties have to interact in the online phase.
-Assumed there are two secretly shared values \( \langle a \rangle\) and \( \langle b \rangle\) which should be multiplied with each other.
-Naturally, each party $P_i$ holds the shares \(\langle a_i\rangle \) and \(\langle b_i\rangle \).\\
-For each multiplication operation, one of the secretly shared Beaver triples \( (\langle x \rangle ,\langle y\rangle , \langle z\rangle )\) generated in the offline phase is needed, which is taken out of the queue.\\ The multiplication is done with following protocol:
+The multiplication is done with following protocol:
 \begin{enumerate}
@@
 As it can be seen, the computations are correct and calculate the expected result, and at the same time no information about \(\langle a \rangle \) or \(\langle b \rangle \) is revealed.\\
-
 \textbf{Runtime:} With the help of the beaver triple, it was possible to express a multiplication as a combination of 5 additions, 2 multiplications with constants and 1 multiplication of two constants by each party.
 Additionally two values were partially opened, but this can be done in constant time too.
-So a multiplication can be done in \(O(1)\).\\
-
+So a multiplication can be done in \(O(1)\).
 \subsubsection{Determining the Result and Verifying Correctness}
@@
 \end{enumerate}
-Correctness: 
+\textbf{Correctness:} 
 $$ \sum_{i=1}^n \sigma_i =  \sum_{i=1}^n  \gamma(r)_i - \alpha_i * r $$
@@
 If the calculations were correct, the MAC was shared properly and not modified, and the tag values were not modified, then the difference between both sums will be 0.
-If it is not 0, the protocol detects that there is a malicious adversary and aborts.\\
-
+If it is not 0, the protocol detects that there is a malicious adversary and aborts.
 \subsubsection{Runtime Complexity}
@@
 Additionally, each party can share a secret in O(n), so the total complexity class of the online phase is linear.
-Therefore, the online phase is really well scalable.\\
-
+Therefore, the online phase is really well scalable.
 \subsection{Problems and upcoming changes}
@@
 The developer team tries to give MAMBA the possibility to compile code just in time as mentioned in ~\cite{ScaleMambaDocu}.
-That would offer the possibility, that the offline data is calculated before and then, when the need arises, a computation could be specified in MAMBA and in-time compiled and executed, without having to wait for the expensive offline calculations.\\
-
+That would offer the possibility, that the offline data is calculated before and then, when the need arises, a computation could be specified in MAMBA and in-time compiled and executed, without having to wait for the expensive offline calculations.
 \subsubsection{Problems}
 It was shown recently in ~\cite{MPCCompression} that there are security flaws in SPDZ 2.0 and that its secret sharing is predictable.
-It doesn't seem to be the case that these imperfections are already fixed in SCALE-MAMBA.\\
+It doesn't seem to be the case that these imperfections are already fixed in SCALE-MAMBA.
+
 Additionally, a single cheater can break the whole system because it is abort-secure.
@@
 This would be an important feature for a real-world use case, so the protocol doesn't get stuck in a loop of abortion.
-
 \section{Implementations}
@@
 The idea is that the system provides the optimal treatment for a patient with HIV based on the  records of similar patients in different hospitals.
-To preserve privacy, the hospitals compute the efficiency of one treatment for one patient together with SMPC.\\
+To preserve privacy, the hospitals compute the efficiency of one treatment for one patient together with SMPC.
+
 The researchers used a setup of 20.000 patient records split between the hospitals, and were able to compute the effectiveness of 100 treatments for one patient in 24 minutes in the online phase.
@@
 They only implemented the online phase of SPDZ, but showed that it scaled linearly with the amount of patient records.
-In the paper the amount of parties participating in the computation was not specified, but because the computation is quite good, it can be assumed that a really simple setup is used that doesn't create a lot of overhead.
-
-TODO: compare, what is the speed and how fast are different ones?
+In the paper the amount of parties participating in the computation was not specified, but because the speed of the computation appears to be quite good, it can be assumed that a really simple setup is used that does not create a lot of overhead.
 
 Supposedly, the offline phase took them around 22 minutes.
-
 \subsection{Other implementations}
@@
 iris and face authentication, with SPDZ was shown.
-The researchers were able to improve existing solutions with the use of SPDZ.\\
+The researchers were able to improve existing solutions with the use of SPDZ.
+
 ~\cite{2019arXiv190100329C} implemented typical machine learning algorithms like linear and logistic regression with SPDZ and showed that it outperforms a competing protocol Obliv-C ~\cite{cryptoeprint:2015:1153} and matches SecureML ~\cite{SecureML}.
-So it can be assumed, that an implementation with SCALE-MAMBA would show even better results.\\
-
+So it can be assumed, that an implementation with SCALE-MAMBA would show even better results.
 \section{Comparison}\label{sec:comparison}
@@
 active and full-threshold security, SCALE-MAMBA, and before SPDZ, seem to be the universally accepted state of the art protocols.
-As already seen before, SCALE-MAMBA has several improvements in run-time and security compared to its predecessors.\\
+As already seen before, SCALE-MAMBA has several improvements in run-time and security compared to its predecessors.
+
 Nevertheless, the requirements are not always the same, so depending on the use-case the choice of protocol might be depend.
-No complete comparison to other protocols will be, but this section should transmit the parameters which should be considered when making this choice.\\
+No complete comparison to other protocols will be, but this section should transmit the parameters which should be considered when making this choice.
 \subsubsection{Two party protocols} There are a lot of protocols, which are restricted to a two party setup.
-As shown in ~\cite{MPCCompression}, frequently used protocols like TinyLEGO ~\cite{TinyLego} and  DUPLO ~\cite{Duplo}, which are both protocols based on garbled circuits, and ABY ~\cite{Aby}, which combines secret sharing and garbled circuits, all outperform SPDZ 2.0 in terms of information loss regarding compression and entropy metrics.\\
+As shown in ~\cite{MPCCompression}, frequently used protocols like TinyLEGO ~\cite{TinyLego} and  DUPLO ~\cite{Duplo}, which are both protocols based on garbled circuits, and ABY ~\cite{Aby}, which combines secret sharing and garbled circuits, all outperform SPDZ 2.0 in terms of information loss regarding compression and entropy metrics.
 \subsubsection{Passive Security}
@@
 they are not secure versus malicious adversaries.
-In a setup with looser security requirement a different protocol might be better suited because a lot of verification can be omitted.\\
+In a setup with looser security requirement a different protocol might be better suited because a lot of verification can be omitted.
 \subsubsection{Different Access structures}
 Full-threshold security is not always a desired property.
-While SCALE-MAMBA can be configured for different access structure, different protocols optimized for these access structures might outperform it.\\
-
+While SCALE-MAMBA can be configured for different access structure, different protocols optimized for these access structures might outperform it.
 \section{Resume}\label{sec:outlook}
@@
 With the theoretical research of the early years turning to more praxis-oriented research, today's SMPC protocols have a bright future.
-With already more and more applications starting to use SMPC in systems that need to guarantee security of private data, and with the protocols further improving in performance and security, the list will only continue growing.\\
+With already more and more applications starting to use SMPC in systems that need to guarantee security of private data, and with the protocols further improving in performance and security, the list will only continue growing.
+
 SCALE-MAMBA is one of the first systems that can is functional and performant enough to be used in real-world scenarios, and, thus, it will probably not take a lot more years until there will be proper commercial solutions.
 And if big companies start to adapt and use these theoretical foundations, then, in my opinion, SMPC will be a really important tool in computer science and especially in distributed environments.
-There are a lot more possible applications that could be coming in the next years, for example in the financial sector or in inter-governmental collaborations.\\
+There are a lot more possible applications that could be coming in the next years, for example in the financial sector or in inter-governmental collaborations.
+
 SMPC can be a technology that shapes the future, or it can be a stepping stone for new technologies to emerge.
@@
 Nevertheless, SCALE-MAMBA is optimized for a setup with active and full threshold security, so if the use case varies, there might be a more suitable SMPC protocol.
-
 \renewcommand{\bibsection}{\section*{References}} % requried for natbib to have "References" printed and as section*, not chapter*
